---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

knitr::opts_chunk$set(
  echo = FALSE
)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

It's interesting how they can have R code being run within the text by using backticks. So, I can say that the standard deviation of the carats all diamonds is `r diamonds |> summarize(sd(carat))`. I also found it interesting how they hid the code so we can't see their _dirty_ secrets. Also, this is in **bold**.

```{r}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01) +
    facet_wrap(~ cut)

smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01) +
    facet_wrap(~ color)

smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01) +
    facet_wrap(~ clarity)
```

