---
title: "Diamond_Exercise"
author: "Shana Mandelbaum"
date: "2024-06-07"
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

The most striking features of this frequency plot are that certain ring sizes are much more popular, for example, ~.25, .75, and 1.0 carat.

```{r, echo = FALSE}
diamonds |> 
  ggplot(aes(x = carat)) +
  geom_histogram(aes(group = cut, fill = cut)) +
  facet_wrap(~ cut)

diamonds |> 
  ggplot(aes(x = carat)) +
  geom_histogram(aes(group = color, fill = color)) +
  facet_wrap(~ color)

diamonds |> 
  ggplot(aes(x = carat)) +
  geom_histogram(aes(group = clarity, fill = clarity)) +
  facet_wrap(~ clarity)
```
This shows that people are more likely to buy small diamonds that have better cut, clarity, and color 
