---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```
```
Practice 27.3.1
Answers for question 3: The striking feature of this paragraph is that it spikes frequently between the carat size of 0.5 and 1. It shows that smaller diamonds are more common. 
```
Practice 27.4.7```
Q1: Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting echo = FALSE on each chunk, set a global option.
```{r}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE
)

smaller %>%
  group_by(cut, carat) %>%
  ggplot(aes(x = carat, y = )) +
  geom_point(color = "red", size = 3, alpha = 0.7) +  
  labs(
    title = "Size ~ Cut",
    x = "Cut",
    y = "Size"
  ) + 
  theme_minimal()
```
```{r}
smaller %>%
  ggplot(aes(x = carat)) + 
  geom_histogram(color = "red", size = 15, alpha = 0.7) +  
  labs(
    title = "Diamond Carat Size Variations based on Cut",
    x = "Carat Size",
    y = "Count"
  ) + facet_wrap(~ cut) + 
  theme_minimal()
  
```
```{r}
smaller %>%
  ggplot(aes(x = carat)) + 
  geom_histogram(color = "skyblue", size = 15, alpha = 0.7) +  
  labs(
    title = "Diamond Carat Size Variations based on Cut",
    x = "Carat Size",
    y = "Count"
  ) + facet_wrap(~ color) + 
  theme_minimal()
```
```{r}
smaller %>%
  ggplot(aes(x = carat)) + 
  geom_histogram(color = "skyblue", size = 15, alpha = 0.7) +  
  labs(
    title = "Diamond Carat Size Variations based on Cut",
    x = "Carat Size",
    y = "Count"
  ) + facet_wrap(~ clarity) + 
  theme_minimal()
```

```


