---
title: "Day 3 Week 2 assignment"
author: "Alexandra Roffe"
date: 2022-06-08
output: html_document
---

day 3, week 2
chapter 7, exercise 7.1:

1: find the population average of the variable "bmi"
```{r}
pop.2 <- read.csv("pop2.csv", stringsAsFactors = TRUE)
mean(pop.2$bmi) # answer: 24.98446
```

2:
```{r}
sd(pop.2$bmi)
```

3: expectation of the sampling distribution for sample average n = 150
```{r}
X.bar <- rep(0, 10^5)
for(i in 1:10^5)
{
  X.samp <- sample(pop.2$bmi, 150)
  X.bar[i] <- mean(X.samp)
}

mean(X.bar)
```
4: stand deviation
```{r}
sd(X.bar)
```
5: identify the central region contains 80%
```{r}
quantile(X.bar,c(0.1,0.9))
```

6: using Central Limit Theorem
```{r}
qnorm(c(0.1,0.9),mean(X.bar),sd(X.bar))
```

Chapter 9, exercise 9.1:
1. average is 3.5
```{r }
magnets <- read.csv("magnets.csv", stringsAsFactor = TRUE)
summary(magnets)
```


2. Is the variable “active” a factor or a numeric variable?
a factor

3. 
```{r}
mean(magnets$change[1:29])
mean(magnets$change[30:50])
```

4. 
```{r}
sd(magnets$change[1:29])
sd(magnets$change[30:50])
```

5:

```{r}
boxplot(magnets$change[1:29])
boxplot(magnets$change[30:50]) # There are 3 outliers

```


Chapter 10, exercises 10.1 and 10.2:
10.1:
1. Simulate the sampling distribution of average and the median of a sample
of size n = 100 from the Normal(3, 2) distribution. Compute the expectation and the variance of the sample average and of the sample median.
Which of the two estimators has a smaller mean square error?

```{r}
mu <- 3
sig <- sqrt(2)
n <- 100

X.bar <- rep(0,10^5)
X.median <- rep(0,10^5)

for(i in 1:10^5)
{
  X <- rnorm(n, mu, sig) # rnorm does random deviates
  X.bar[i] <- mean(X)
  X.median[i] <- median(X)
}
# variance:
var(X.bar) # sample average
var(X.median) # sample median. 

# expectation: 
mean(X.bar)
mean(X.median)



```
The variance of the sample average of the mean is 0.020 roughly. 
The variance of the sample median is .03 roughly. 
The mean error of the sample average is less than the mean square error of the
sample median.

2: Simulate the sampling distribution of average and the median of a sample
of size n = 100 from the Uniform(0.5, 5.5) distribution. Compute the
expectation and the variance of the sample average and of the sample
median. Which of the two estimators has a smaller mean square error?
```{r}
# need for the uniform distribution, alpha = 0.5, b = 5.5. 
a <- 0.5
b <- 5.5
X.bar <- rep(0, 10^5)
X.median <- rep(0, 10^5)
for(i in 1:10^5)
{
  x <- runif(100, a, b)
  X.bar[i] <- mean(x)
  X.median[i] <- median(x)
}

mean(X.bar)
mean(X.median)

var(X.bar)
var(X.median)

```
In this case, the estimators are unbiased. 
The variance of each of the estimators is equal to it's mean square error. 
Like in the problem right above, the square error of the sample average is less than that of its sample median. 


**Exercise 10.2**
1:
```{r}
ex2.csv <- read.csv("ex2.csv", stringsAsFactors = TRUE) # this is a sample of 150
mean(ex2.csv$group == "HIGH")

```


2:
```{r}

mean(pop.2$group == "HIGH")

```

3:
```{r}
X.bar <- rep(0, 10^5)
for(i in 1:10^5)
{
  X.samp <- sample(pop.2$group, 150)
  X.bar[i] <- mean(X.samp == "HIGH")
}

mean(X.bar)
```
4:
```{r}
var(X.bar)
```

```{r}
p <- mean(pop.2$group == "HIGH")
p*(1-p)/150


```