---
title: "day4_exercises"
output: html_document
date: '2022-06-09'
---
Chapter 2 (ISRS)
2.2

a) 30/34 in control and 45/69 in treatment
b) 
  i) HO: null hypothesis is that the treatment does not work (group and outcome are independent)
     HA: alternative hypothesis is that the treatment is effective (group and outcome aren't independent)
     
  ii)
    We write “alive” on 28 cards representing patients who were alive at the end of study, and “dead” on 75 cards representing patients who were not. Then, we shuffle these cards and split them into two groups: one group of size 69 representing treatment, and another group of size 34 representing control. We calculate the difference between the proportion of “dead” cards in the treatment and control groups (treatment - control) and record this value. We repeat this 100 times to build a distribution centered at 0. Lastly, we calculate the fraction of simulations where the simulated differences in proportions are less than the alpha. If this fraction is low, we conclude that it is unlikely to have observed an outcome by chance and that the null hypothesis should be rejected in favor of the alternative.


  iii) 45/69 - 30/34 = -0.23; From the graph we can see that it never occurred in any of our random trials that the outcome was -0.23, therefore we reject the null (p-value is 0% which is smaller than alpha). It is very unlikely that treatment and death are independent of each other. In other words, null is rejected because our real experiment falls in the alpha section of the null distribution.
    
2.5 

a) Null Hypothesis: provocative clothing has no effect on diners (they are independent of each other)
  Alternative Hypothesis: provocative clothing has effect on diners (they are dependent of each other)
  
b) 5/20 - 15/25 = -0.35

c) from the graph we can estimate that the probability of getting -0.35 is around 0.025 (2.5%) which is less than the alpha (5%) so we reject the null. Therefore it is very unlikely that clothing and effect on diners are independent of each other.

Chapter 9

9.2.1
```{r}
expect <- 3.5 
sd_a <- 3 #active magnet patients 
sd_i <- 1.5 #inactive placebo patients
# create a sampling distribution of the statistic 
statistic_outcomes <- rep(0,10^5)
for (i in 1:10^5){
  norm_a <- rnorm(29,expect,sd_a)
  norm_i <- rnorm(21,expect,sd_i)
  N <- mean(norm_a) - mean(norm_i)
  D <- sqrt(var(norm_a)/29 + var(norm_i)/21)
  statistic_outcomes[i] <- N/D
}

# ? 95% interval of the sampling distribution of the statistic
quantile(statistic_outcomes, c(0.025, 0.975))
#     2.5%     97.5% 
# -2.017883  2.009704 
```

2 
Did the magnets have any effect? 
```{r}
mg <- read.csv(file="magnets.csv")
active_mg <- mg$change[1:29]
inactive_placebo <- mg$change[30:50]
N <- mean(active_mg) - mean(inactive_placebo)
D <- sqrt(var(active_mg)/29 + var(inactive_placebo)/21)
N/D #5.985601

```
The magnets treatment have an effect on patients because 5.985601 is greater than 
2.009704 which is 97.5% (so since it is less than 5% chance we reject the null)


  