
# Setup
```{r set-up}
library(tidyverse)
```

# Section 12.2.1
## Exercise 2

Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

1. Extract the number of TB cases per country per year.
2. Extract the matching population per country per year.
3. Divide cases by population, and multiply by 10000.
4. Store back in the appropriate place.

```{r table2-rates}

table2 |>
    pivot_wider(names_from = type, values_from = count) |>
    mutate(cases_by_population = (cases/population)*10000) |>
    pivot_longer(names_to = "type", values_to = "count", -c(country, year, cases_by_population))


```



```{r table4a-table4b-rates}

table4a_longer <- table4a |>
    pivot_longer(names_to = "year", values_to = "cases", -country)
table4a_longer

table4b_longer <- table4b |>
    pivot_longer(names_to = "year", values_to = "population", -country)
table4b_longer


inner_join(table4a_longer, table4b_longer) |>
    mutate(cases_by_population = (cases/population)*10000)


```


Which representation is easiest to work with? Which is hardest? Why?
- `table2` representation is easiest to work with and `table4a + table4b` is the hardest because in `table2`, we don't need to join two tables which makes easiest to seee that data while 
in `table4a + table4b`, before we join, we need to change to long format and then join.


# Section 12.3.3
## Exercise 1

Why are pivot_longer() and pivot_wider() not perfectly symmetrical?
Carefully consider the following example:

```{r example-code-ex1}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")
```
(Hint: look at the variable types and think about column names.)

pivot_longer() has a names_ptypes argument, e.g.  names_ptypes = list(year = double()). What does it do?

```{r sec-12-3-3-ex-1}

stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return", names_ptypes = list(year = character()))

```

Answer:
- pivot_longer() and pivot_wider() not perfectly symmetrical because when we do longer the data type of year becomes from double to character because the column names are character.
- We can use names_ptypes in piviot_longer() to check if the data type is expected


## Exercise 3

What would happen if you widen this table? Why? How could you add a new column to uniquely identify each value?

```{r}
people <- tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
```

```{r sec-12-3-3-ex-3}
people |>
    pivot_wider(names_from = names, values_from = values)

people |>
    summarise(n = n(), .by = c(name, names)) |>
    filter(n > 1L)
```

Answer:
- If I widen this table, I get an error because in name, there are not unique name. We have same name for age, which 1 person can't have 2 differnet age.
- We could add an id column to identify each value. 


# Section 27.2.1
## exercises 1


## exercises 2