
---
title: "Day 4 Week 2 assignment"
author: "Alexandra Roffe"
date: 2022-06-10
output: html_document
---
Chapter 5: 

5.20 If a linear model has a very strong negative relationship
with a correlation of -0.97, how much of the variation in the response is explained by the explanatory variable?

  R = -0.97, R^2 = 0.94. Then, 94% of the variation is from the explanatory variable. 

  R square described the proportion of the variability in the response variable explained by explanatory variable. 
  
  
5.29: Body measurements:
a) a linear positive relationship. 
  
b) y = mx + b

  m = 1.0176 is slope
  b = -105.0113
  
  With an increase of 1 cm of height, the width increase by 1.017. Slope is positive because there is a positive correlation between height and width, according to the data given. 
  
  height of 0 will have a weight of -105. 
  
  
c) 
  Null: Height does not effect weight; there is no relationship between the two. The slope of the line is zero.
  Alternative: Height does effect weight. The slope of the graph is not zero.
  
  p-value = 0.0000
  We can reject the null and conclude that height and weight do affect one another. There is positive relation between height and wight. 
  
d) 
R = 0.72
```{r}
.72 * .72
```
  This means that proportion of variation in height and weight is .52. 
  52% of the variation is due to explanatory data. 
  R squ
  
  

Section 3.1 of Intro to Statistical Learning, do Lab 3.6.2
```{r}

library(MASS)
library(ISLR2)
#head(Boston)

lm.fit <- lm(medv ~ lstat, data = Boston)
#lm.fit
#summary(lm.fit)

names(lm.fit)
coef(lm.fit)
confint(lm.fit)


predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "confidence")
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "prediction")


plot(Boston$lstat, Boston$medv, col = "green")
abline(lm.fit)

par(mfrow = c(2, 2))
plot(lm.fit)


```
  
```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```
  