---
title: "Day 7 Exercises"
author: "Shoshana Farber"
---

# Chapter 7

```{r}
pop.1 <- read.csv("pop1.csv")
X.samp <- sample(pop.1$height, 100)
X.samp
```

```{r}
mean(X.samp)
mean(pop.1$height)
```

```{r}
X.samp <- sample(pop.1$height,100)
X.bar <- mean(X.samp)
X.bar
```

```{r}
X.samp <- sample(pop.1$height,100)
X.bar <- mean(X.samp)
X.bar
```

```{r}
X.bar <- rep(0, 10^5)
for(i in 1:10^5)
{
  X.samp <- sample(pop.1$height, 100)
  X.bar[i] <- mean(X.samp)
}
hist(X.bar)
```

```{r}
mean(pop.1$height)
sd(pop.1$height)
mean(X.bar)
sd(X.bar)
```

```{r}
mean(abs(X.bar - mean(pop.1$height)) <= 1)
```

```{r}
X.bar <- rep(0, 10^5)
for(i in 1:10^5)
{
  X.samp <- rbinom(64, 10, 0.5)
  X.bar[i] <- mean(X.samp)
}
hist(X.bar)
```

```{r}
mean(X.bar)
sd(X.bar)
```

```{r}
quantile(X.bar, c(0.025, 0.975))
```

```{r}
qnorm(c(0.025, 0.975), mean(X.bar), sd(X.bar))
```

variability decreases as sample size increases
```{r}
unif.10 <- rep(0, 10^5)
unif.100 <- rep(0, 10^5)
unif.1000 <- rep(0, 10^5)
for(i in 1:10^5)
{
  X.samp.10 <- runif(10, 3, 7)
  unif.10[i] <- mean(X.samp.10)
  X.samp.100 <- runif(100, 3, 7)
  unif.100[i] <- mean(X.samp.100)
  X.samp.1000 <- runif(1000, 3, 7)
  unif.1000[i] <- mean(X.samp.1000)
}

var(unif.10)
var(unif.100)
var(unif.1000)
```

```{r}
exp.10 <- rep(0, 10^5)
exp.100 <- rep(0, 10^5)
exp.1000 <- rep(0, 10^5)
for(i in 1:10^5)
{
  X.samp.10 <- rexp(10, 0.5)
  exp.10[i] <- mean(X.samp.10)
  X.samp.100 <- rexp(100, 0.5)
  exp.100[i] <- mean(X.samp.100)
  X.samp.1000 <- rexp(1000, 0.5)
  exp.1000[i] <- mean(X.samp.1000)
}

mean(exp.10)
mean(exp.100)
mean(exp.1000)

var(exp.10)
var(exp.100)
var(exp.1000)
```

## Question 7.1

1. Compute the population mean of the variable "bmi"
2. Compute the population standard deviation of the variable "bmi"
```{r}
pop.2 <- read.csv("pop2.csv")
mean(pop.2$bmi)
sd(pop.2$bmi)
```
pop mean = 24.98
pop sd = 4.19

3. Compute the expectation of the sampling distribution for the sample average of the variable
4. Compute the standard deviation of the sampling distribution for the sample average of the variable
*assume sample size of n = 150
```{r}
X.bar <- rep(0, 10^5)
for(i in 1:10^5)
{
  X.samp <- sample(pop.2$bmi, 150)
  X.bar[i] = mean(X.samp)
}
mean(X.bar)
sd(X.bar)
```
sample avg = 24.98
sample sd = 0.34

5. Identify, using simulations, the central region that contains 80% of the sampling distribution of the sample average
```{r}
quantile(X.bar, c(0.1, 0.9))
```
[24.55, 25.42]

6. Identify, using CLT, an approx of the central region that contains 80% of the sampling distribution of the sample average
```{r}
qnorm(c(0.1, 0.9), mean(X.bar), sd(X.bar))
```
[24.55, 25.42]

# Chapter 9

```{r}
cars <- read.csv("cars.csv")
summary(cars)
```

```{r}
X.bar <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- rexp(201, 1/12000)
  X.bar[i] <- mean(X)
}
mean(abs(X.bar-12000) <= 1.96 * 0.0705 * 12000)
```

```{r}
mid.range <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- runif(100, 3, 7)
  mid.range[i] <- (max(X) + min(X))/2
}
quantile(mid.range, c(0.025, 0.975))

mean(mid.range)
sd(mid.range)
```

## Question 9.1

n = 50

1. What is the sample average of the change in score between the patient's rating before the application and after the application

2. Is the variable "active" a factor or a numeric variable?

```{r}
magnets <- read.csv("magnets.csv")
summary(magnets)
```
sample average = mean = 3.5
"active" is a factor

3. Compute the average value of the variable "change: for the patients that received an active magnet and for those that received an inactive placebo

```{r}
mean(magnets$change[1:29])
mean(magnets$change[30:50])
```
mean active = 5.24
mean inactive = 1.10

Compute the sample sd of the variable "change" for those that received active and inactive

```{r}
sd(magnets$change[1:29])
sd(magnets$change[30:50])
```
sd active = 3.24
sd inactive = 1.58

5. Produce a boxplot of the variable "change" for the patients that received an active magnet and for those that received an inactive placebo. War is the number of outliers in each subsequence?

```{r}
boxplot(magnets$change[1:29])
boxplot(magnets$change[30:50])
```
0 outliers for active
3 outliers for inactive

# Chapter 10

```{r}
mean(cars$price)
mean(cars$price, na.rm = TRUE)
```

```{r}
lam <- 1/13000
X.bar <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- rexp(201, lam)
  X.bar[i] <- mean(X)
}
mean(abs(X.bar - 1/lam) <= 1000)
```

```{r}
mu <- 13000
sig <- 13000/sqrt(201)
pnorm(14000, mu, sig) - pnorm(12000, mu, sig)
```

```{r}
mu <- 12000
sig <- 12000/sqrt(201)
pnorm(13000, mu, sig) - pnorm(11000, mu, sig)
```

```{r}
mu <- 3
sig <- sqrt(2)
X.bar <- rep(0, 10^5)
mid.range <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- rnorm(100, mu, sig)
  X.bar[i] <- mean(X)
  mid.range[i] <- (max(X) + min(X))/2
}
var(X.bar)
var(mid.range)
```

```{r}
a <- 0.5
b <- 5.5
X.bar <- rep(0, 10^5)
mid.range <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- runif(100, a, b)
  X.bar[i] <- mean(X)
  mid.range[i] <- (max(X) + min(X))/2
}
var(X.bar)
var(mid.range)
```

```{r}
mu <- 5
std <- sqrt(3)
X.var <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- rnorm(20, mu, std)
  X.var[i] <- var(X)
}
mean(X.var)
mean((19/20) * X.var)

var(X.var)
var((19/20) * X.var)
```

## Question 10.1

1. Simulate the sampling dist of avg and the median of a sample of size n = 100 from the Normal(3,2) dist. Compute the expectation and the varaince of the sample avg and of the sample median. Which of the 2 estimators has a smaller mean square error?
```{r}
mu <- 3
sig <- sqrt(2)
X.bar <- rep(0, 10^5)
X.med <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- rnorm(100, mu, sig)
  X.bar[i] <- mean(X)
  X.med[i] <- median(X)
}
mean(X.bar)
mean(X.med)

var(X.bar)
var(X.med)
```
MSE of X.bar is smaller 

2. Simulate the sampling dist of average and the median of a sample of size n = 100 from the Uniform(0.5, 5.5) dist. Compute the expectation and the variance of the sample average and of the sample median. Which of the 2 estimators had a smaller mean square error?
```{r}
a <- 0.5
b <- 5.5
X.bar <- rep(0, 10^5)
X.med <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- runif(100, a, b)
  X.bar[i] <- mean(X)
  X.med[i] <- median(X)
}
mean(X.bar)
mean(X.med)

var(X.bar)
var(X.med)
```
MSE of X.bar is smaller

## Question 10.2

1. Compute the proportion in the sample of those with a high level of blood pressure
```{r}
ex.2 <- read.csv("ex2.csv", stringsAsFactors = TRUE)
summary(ex.2)
```
proportion = 37/150 = 0.247
```{r}
37/150
```

```{r}
mean(ex.2$group == "HIGH")
```

2. Compute the proportion in the population of those with a high level of blood pressure
```{r}
pop.2 <- read.csv("pop2.csv", stringsAsFactors = TRUE)
summary(pop.2)
```
proportion = 28126/100000 = 0.281
```{r}
28126/100000
```

```{r}
mean(pop.2$group == "HIGH")
```

3. Simulate the sampling dist of the sample proportion and compute its expectation
```{r}
P.hat <- rep(0, 10^5)
for(i in 1:10^5)
{
  X <- sample(pop.2$group, 150)
  P.hat[i] <- mean(X == "HIGH") 
}
mean(P.hat)
```
4. Compute the variance of the sample proportion
```{r}
var(P.hat)
```

5. It is proposed in Section 10.5 that the variance of the sample proportion is Var(P.hat) = p(1-p)/n. Examine this proposal in the current setting. 
p = prob of event (having HIGH)
n = sample size (150)
```{r}
p <- mean(pop.2$group == "HIGH")
p*(1-p)/150
```

