```{r}
library(tidyverse)
library(dplyr)
library(readr)

ex.1 <- read.csv("ex1.csv")
summary(ex.1)

pop.1 <- read.csv("pop1.csv")
summary(pop.1)
plot(table(pop.1$height))

sample(pop.1$height, 1) # returns a singular random value of height 

X <- pop.1$height
mean(abs(X-170) <= 10)

Y <- c(6.3, 6.9, 6.6, 3.4, 5.5, 4.3, 6.5, 4.7, 6.1, 5.3)
abs(Y-5)
abs(Y-5) <= 1
mean(abs(Y-5) <= 1)
```

Question 4.1

1. what is the value of p?
  p + 2p + 3p + 4p + 5p + 6p = 1 
  21p = 1
  p = 1/21

2. P(Y < 3) = ?
  P(Y < 3) = P(Y = 0) + P(Y = 1) + P(Y = 2)
  (1/21) + (2/21) + (3/21) = 0.2857

3. P(Y = odd) = ?
  P(Y = odd) = P(Y = 1) + P(Y = 3) + P(Y = 5)
  (2/21) + (4/21) + (6/21) = 0.5714

4. P(1 <= Y <= 4) = ?
  P(1 <= Y <= 4) = P(Y = 1) + P(Y = 2) + P(Y = 3) + P(Y = 4)
  (2/21) + (3/21) + (4/21) + (5/21) = 0.6667

```{r}
Y_val = c(0,1,2,3,4,5)
P_val = c(1,2,3,4,5,6)/21
```

5.  P(|Y - 3| < 1.5) = ?
```{r}
abs(Y_val - 3)
```
  P(|Y - 3| < 1.5) = P(Y = 2) + P(Y = 3) + P(Y = 4)
  (3/21) + (4/21) + (5/21) = 0.5714

```{r}
Y_val = c(0,1,2,3,4,5)
P_val = c(1,2,3,4,5,6)/21

# 6. E(Y) = ?
  (E <- sum(Y_val * P_val))
  # 3.333

# 7. Var(Y) = ?
  (Var <- sum(((Y_val - E)^2) * P_val))
  # 2.222

# 8. what is SD of Y?
  (sd <- sqrt(Var))
  # 1.491
```

Question 4.2

1. What is the prob of winning the game?
P{H} = (1/2)
P{win} = P{H,H,H} = (1/2)(1/2)(1/2) = (1/8)

2. What is the prob of losing the game?
P{lose} = 1 - P{win} = 1 - (1/8) = (7/8)

3. What is the expected gain for the player that plays?
```{r}
gain <- c(8, -2)
prob <- c(1,7)/8
(E <- sum(gain * prob))
```
Expected gain = -0.75

Question 6.1

mu = 560
sigma = 57
n = 9
sample mean = 630
sample SD = 61

1. What is the prob that the total weight of 8 people exceeds 650?
```{r}
1 - pnorm(650, 560, 57)
```
p = 0.05717

2. What is the prob that the total weight of 9 people exceeds 650?
```{r}
1 - pnorm(650,630,61)
```
p = 0.3715

3. What is the central region that contains 80% of the distribution of the total weight of 8 people?
80% = 10% - 90%
```{r}
qnorm(0.1, 560, 57)
qnorm(0.9, 560, 57)
```
[486.95, 633.05]

4. What is the central region that contains 80% of the distribution of the total weight of 9 people?
```{r}
qnorm(0.1, 630, 61)
qnorm(0.9, 630, 61)
```
[551.83, 708.175]
