# Exercie 1

Compute the rate for table2, and table4a + table4b. You will need to perform four operations:

1. Extract the number of TB cases per country per year.
2. Extract the matching population per country per year.
3. Divide cases by population, and multiply by 10000.
4. Store back in the appropriate place.


```{r table2-rates}
table2 |>
    pivot_wider(names_from = type, values_from = count) |>
    mutate(cases_by_pop = (cases/population)*10000) |>
    pivot_longer(names_to = "type", values_to = "count", -c(country, year, cases_by_pop))
```

```{r table4a-table4b-rates}
table4a
table4b

table4a_longer <- table4a |>
    pivot_longer(names_to = "year", values_to = "cases", -country)

table4a_longer

table4b_longer <- table4b |>
    pivot_longer(names_to = "year", values_to = "population", -country)
table4b_longer

inner_join(table4a_longer, table4b_longer) |>
    mutate(cases_by_pop = (cases/population)*10000)


```

Which representation is easiest to work with? Which is hardest? Why
-  `table2` is easier than `table4a + table4b` to work with because we are working with a single dataframe and it requires less lines of code. However, I do think that `table4a + table4b` makes more logical sense.

# Exercie 12.3.3

## Exercise 1


Why are pivot_longer() and pivot_wider() not perfectly symmetrical?
Carefully consider the following example:

```{r}

stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  pivot_wider(names_from = year, values_from = return) %>% 
  pivot_longer(`2015`:`2016`, names_to = "year", values_to = "return")

```

(Hint: look at the variable types and think about column names.)

- pivot_longer() and pivot_wider() are not symmetrical because when pivot_longer() is executed, the column names are in
different format than the original value. But for pivot_wider(), the new columns remain in the same format.

pivot_longer() has a names_ptypes argument, e.g.  names_ptypes = list(year = double()). What does it do?

- The argument checks if the conversion maintains the original data type of a column such as year (which was originally a double)


## Exercie 3

What would happen if you widen this table? Why? How could you add a new column to uniquely identify each value?
```{r}

people <- tribble(
  ~name,             ~names,  ~values,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)

```

 - The table depicts that there are two people with the same name which can cause issues with pivot_wider(). An new column with the specific ID must be
 established to differenciate each person.

 
