---
title: "Diamond sizes"
date: 2026-06-01
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01) +
  scale_x_continuous(breaks = seq(0, 2.5, by = 0.2)) +
  labs(x = "Carat", y = "Diamond count", title = "Frequency of carat size")
```

---
27.3.1 Exercise 3
---

#### Most striking observations
1. The diamond count peaks between 0.3 and 0.4 carats.
2. The majority of diamonds are below 1.5 carats.

---

---
27.4.7 Exercise 1
---

### Diamond Size By Cut, Colour, and Clarity

#### Diamond Size By Cut

```{r, echo = FALSE}
smaller %>%
  ggplot(aes(x = carat)) +
  geom_freqpoly(binwidth = 0.01) +
  facet_wrap(~ cut) +
  labs(x = "Carat", y = "Count", title = "Carat Frequencies By Cut")
```

```{r, echo = FALSE}
smaller %>%
  ggplot(aes(x = cut, y = carat, fill = cut)) +
  geom_boxplot() +
  labs(x = "Cut", y = "Carat", title = "Carat Variations By Cut")
```

#### Diamond Size By Colour

```{r, echo = FALSE}
smaller %>%
  ggplot(aes(x = carat)) +
  geom_freqpoly(binwidth = 0.01) +
  facet_wrap(~ color) +
  labs(x = "Carat", y = "Count", title = "Carat Frequencies By Color")
```

```{r, echo = FALSE}
smaller %>%
  ggplot(aes(x = color, y = carat, fill = color)) +
  geom_boxplot() +
  labs(x = "Color", y = "Carat", title = "Carat Variations By Color")
```

#### Diamond Size By Clarity
```{r, echo = FALSE}
smaller %>%
  ggplot(aes(x = carat)) +
  geom_freqpoly(binwidth = 0.01) +
  facet_wrap(~ clarity) +
  labs(x = "Carat", y = "Count", title = "Carat Frequencies By Clarity")
```

```{r, echo = FALSE}
smaller %>%
  ggplot(aes(x = clarity, y = carat, fill = clarity)) +
  geom_boxplot() +
  labs(x = "Clarity", y = "Carat", title = "Carat Variations By Clarity")
```