---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)
# knitr::opts_chunk$set(echo = FALSE)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

---
27.3.1 Exercise 3
---
---
In this frequency polygon we can see that a larger number of diamonds are below 1.5carats and a smaller number are above larger than 1.5. Also, noticeabley there is a smaller production of diamond that are 0.5,1,1.5 , 2 and 2.5 carats 
---

---
27.4.7 Exercise 1
---
---
```{r, echo = FALSE}
head(diamonds)

diamonds%>% ggplot(aes (x= carat , color = cut) )+
geom_freqpoly(binwidth = 0.07) + facet_wrap(~cut)
# or geomhistogram() wiht fill = cut
diamonds%>%group_by(cut)%>% summarize(count = n())

diamonds %>% ggplot(aes(carat, color = color))+
geom_freqpoly(binwidth = 0.07) + facet_wrap(~color)
diamonds%>%group_by(color)%>% summarize(count = n())

diamonds %>% ggplot(aes(carat, color = clarity))+
geom_freqpoly(binwidth = 0.07)+facet_wrap(~clarity)
diamonds%>%group_by(clarity)%>% summarize(count = n())

```