---
title: "Diamond sizes"
author: "Alexandra Roffe"
date: 2022-06-07
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds |> 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller |> 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

The striking features of the graph is that there are stiff peaks at 0.3 and 1.0 carats. 

Diamonds sizes vary by cut, color, and clarity as shown in the charts below. 

```{r}
ggplot(diamonds, aes(y = carat, x = cut)) +
  geom_boxplot()
```

Diamonds with lower cuts tend to be be larger.

```{r}
ggplot(diamonds, aes(y = carat, x = color)) +
  geom_boxplot()
```

Diamonds with worse color are larger than diamonds with color. J is the worst and D is the best. 


```{r}
ggplot(diamonds, aes(y = carat, x = clarity)) +
  geom_boxplot()
```

Lastly, diamonds with worse clarity (I1) are larger than diamonds with better clarity (IF). 