---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

---
27.3.1 Exercise 3
---
---
TODO
---
# Frequency Polygon of Diamond Carat Weights
> The frequency polygon is noticably left skewed, with lighter carat weights occuring most frequently.
> The graph has a trend of immediately spiking up and down, eventually subsiding at a carat size of 2.5.

---
27.4.7 Exercise 1
---
---
TODO
---
Add a section that explores how diamond sizes vary by cut, colour, and clarity. 
Assume you’re writing a report for someone who doesn’t know R, and 
instead of setting echo = FALSE on each chunk, set a global option.
knitr::opts_chunk$set(
  echo = FALSE
)
```{r}
smaller %>% 
  ggplot(mapping = aes(x = cut)) + 
  geom_bar(mapping = aes(fill = clarity)) +
  facet_wrap(~ color)
```