---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

---
27.3.1 Exercise 3
---
---

- Maximum number of diamonds are roughly around 0.3 carat then around 1 carat.
- Overall there are more of the smaller carat diamonds then larger ones.
```{r}
summary(smaller['carat'])%>%
knitr::kable(col.names = "KEY STATS", caption="Tab 1")
```

---
---
27.4.7 Exercise 1
---
---


## Distribution by color
```{r}
diamonds%>%
ggplot(aes(x=carat,color = color))+
geom_freqpoly(binwidth=0.1)+
facet_wrap(~color, scales = "free_y")
```

### Summary table
```{r}
 select(diamonds,color)%>%
 group_by(color)%>%
 summarise(Counts = n())%>%
 knitr::kable(caption = "Tab 2:color_summary",col.names = c("Colors","Counts"))
```

## Distribution by cut

```{r}
diamonds%>%
ggplot(aes(x = carat, color = cut))+
geom_freqpoly(binwidth = 0.1)+
facet_wrap(~cut,scales = "free_y")
```

### Summary table

```{r}
 select(diamonds,cut)%>%
 group_by(cut)%>%
 summarise(Counts = n())%>%
 knitr::kable(caption = "Tab 3:cut_summary",col.names = c("Type of Cut", "Counts"))
```
## Distribution by clarity

```{r}
diamonds%>%
ggplot(aes(x = carat, color = clarity))+
geom_freqpoly(binwidth = 0.1)+
facet_wrap(~clarity,scales = "free_y")
```

### Summary table
```{r}
 select(diamonds,clarity)%>%
 group_by(clarity)%>%
 summarise(Counts = n())%>%
 knitr::kable(caption = "Tab 4:clarity_summary",col.names=c("Clarity","Counts"))
```
---

