---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds |> 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller |> 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

Striking features of the above graph: it presents three main peaks at around 0.3, 0.7 and 1.0 

Add a section that explores how diamond sizes vary by cut, colour, and clarity. Assume you’re writing a report for someone who doesn’t know R, and instead of setting echo = FALSE on each chunk, set a global option:

```{r}
knitr::opts_chunk$set(
  echo = FALSE
)
```

Diamond size by cut:
```{r}
ggplot(diamonds, aes(x = carat)) +
  geom_freqpoly(binwidth = 0.01) + 
  facet_wrap(~ cut)
```



Diamond size by colour:
```{r}
ggplot(diamonds, aes(x = carat)) +
  geom_freqpoly(binwidth = 0.01) + 
  facet_wrap(~ color)
```

Diamond size by clarity:
```{r}
ggplot(diamonds, aes(x = carat)) +
  geom_freqpoly(binwidth = 0.01) + 
  facet_wrap(~ clarity)
```

