---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

 The striking features include:
 - Most diamonds are around 2.8 carats
 - Majority of diamonds are less than 1.25 carats

<!-- this is comment -->


# Section 27.4.7
## Exercise 1

Add a section that explores how diamond sizes vary by cut, colour, and clarity. 
Assume you’re writing a report for someone who doesn’t know R, 
and instead of setting echo = FALSE on each chunk, set a global option.

```{r, echo = FALSE}
knitr::opts_chunk$set(
  echo = FALSE
)
```


```{r}
diamonds |>
  ggplot(aes(x = carat, color = cut, fill = cut)) +
  geom_histogram(bins = 30)
```

```{r}
diamonds |>
  ggplot(aes(x = carat, color = color, fill = color)) +
  geom_histogram(bins = 30)
```

```{r}
diamonds |>
  ggplot(aes(x = carat, color = clarity, fill = clarity)) +
  geom_histogram(bins = 30)
```