---
title: "Diamond sizes"
date: 2016-08-25
output: html_document
---

```{r setup, include = FALSE}
library(ggplot2)
library(dplyr)

smaller <- diamonds %>% 
  filter(carat <= 2.5)
```

We have data about `r nrow(diamonds)` diamonds. Only 
`r nrow(diamonds) - nrow(smaller)` are larger than
2.5 carats. The distribution of the remainder is shown
below:

```{r, echo = FALSE}
smaller %>% 
  ggplot(aes(carat)) + 
  geom_freqpoly(binwidth = 0.01)
```

---
27.3.1 Exercise 3
---
---
The polygon's extremas are concentrated to where the carat value is lower. This indicates that the higher the carat value is, the more rare it is; in addition, the lower the carat value, the more common it is. 
---

---
27.4.7 Exercise 1
---

```{r}
knitr::opts_chunk$set(
  echo = FALSE
)
smaller %>%
  ggplot(aes(x=cut, y=carat, fill=cut)) +
  geom_boxplot() +
  labs(
    title="How Diamond Sizes Vary For Different Cuts"
  )
```

---
The different cuts all have similar minimums and maximums, however their mean and IQR all vary. The 'Fair' cut has the largest mean of 1.0 carat, while 'Ideal' has a lower mean near 0.5 carat. 
---

```{r}
knitr::opts_chunk$set(
  echo = FALSE
)
smaller %>%
  ggplot(aes(x=color, y=carat, fill=color)) +
  geom_boxplot() +
  labs(
    title="How Diamond Sizes Vary For Different Colors"
  )
```

---
Certain consecutive colors have similar size means, however overall the average diamond size and IQR of different colors also vary. 
---

```{r}
knitr::opts_chunk$set(
  echo = FALSE
)
smaller %>%
  ggplot(aes(x=clarity, y=carat, fill=clarity)) +
  geom_boxplot() +
  labs(
    title="How Diamond Sizes Vary For Different Clarities"
  )
```

---
There is a clear correlation between the average diamond sizes and the clarities. F1 clarity has the lowest diamond sizes, meanwhile I1 has the largest diamond sizes.
---