---
title: "Statistical Inference"
author: "Jake Hofman"
date: "April 20, 2017"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r}
library(tidyverse)
#library(ggplot2)
#library(reshape)
#library(dplyr)

theme_set(theme_bw())

set.seed(42)
```

# Estimating a proportion
## Point estimate and sampling distribution
Repeatedly flip a biased coin 100 times and estimate its bias.
Adapted from Yakir 11.2.3.
```{r}
estimate_coin_bias <- function(n, p) {
  mean(rbinom(n,1,p))
}

n <- 100
p <- 0.3
p_hat <- replicate(1e5, estimate_coin_bias(n, p))

# plot the sampling distribution
qplot(x=p_hat, geom="histogram", binwidth=0.01) +
  geom_vline(xintercept=p) +
  geom_vline(xintercept=mean(p_hat), linetype=2, color="red")

# repeat this for different sample sizes
plot_data <- data.frame()
for (n in c(100, 200, 400, 800)) {
  tmp <- data.frame(n=n, p_hat=replicate(1e5, estimate_coin_bias(n, p)))
  plot_data <- rbind(plot_data, tmp)
}

qplot(data=plot_data, x=p_hat, geom="histogram", binwidth=0.01, facets = . ~ n)

se <- plot_data %>%
  group_by(n) %>%
  summarize(se=sd(p_hat))
qplot(data=se, x=n, y=se) +
  stat_function(fun=function(n) {sqrt(p * (1 - p) / n)}, linetype=2)

```

## Confidence intervals
```{r}
set.seed(42)
n <- 100
p <- 0.3
p_hat <- replicate(1e5, estimate_coin_bias(n, p))

# compute upper and lower confidence intervals
LCL <- p_hat - 1.96*sqrt(p_hat*(1-p_hat)/n)
UCL <- p_hat + 1.96*sqrt(p_hat*(1-p_hat)/n)

# check how often the true proportion is contained in the estimated confidence interval
mean(p >= LCL & p <= UCL)

# plot 100 confidence intervals and the true value
plot_data <- data.frame(p_hat, LCL, UCL)[1:100,]
plot_data <- transform(plot_data, contains_p=(p >= LCL & p <= UCL))
ggplot(data=plot_data, aes(x=1:nrow(plot_data), y=p_hat, color=contains_p)) +
  geom_pointrange(aes(ymin=LCL, ymax=UCL)) +
  geom_hline(yintercept=p, linetype=2) +
  xlab('') +
  scale_color_manual(values=c("red","darkgreen")) +
  theme(legend.position="none")
```


## Hypothesis testing
```{r}
# construct a null distribution: what would happen if the coin were fair?
set.seed(42)
n <- 100
p0_hat <- replicate(1e5, estimate_coin_bias(n, p=0.5))

# now conduct one experiment with 100 rolls from a biased coin
p_hat <- estimate_coin_bias(n, p=0.3)

# plot the null distribution and see where the observed estimate lies in it
qplot(x=p0_hat, geom="histogram", binwidth=0.01) +
  geom_vline(xintercept=p_hat, linetype=2, color="red")

# compare this to our experiment
# how likely is it that we would see an estimate this extreme if the coin really were fair?
num_as_extreme <- sum(p0_hat <= p_hat)
p_value <- num_as_extreme / length(p0_hat)
p_value
```
Only `r num_as_extreme` out of `r length(p0_hat)` estimates from a fair coin with p=0.5 would result in an estimate of p_hat=`r p_hat` or smaller, corresponding to a p-value of `r p_value`.


# Comparing two proportions
## Point estimates and sampling distributions
Repeatedly flip two coins, each 500 times and estimate their bias.
```{r}
estimate_coin_bias <- function(n, p) {
  mean(rbinom(n,1,p))
}

pa <- 0.12
pb <- 0.08
n <- 500
pa_hat <- replicate(1e5, estimate_coin_bias(n, pa))
pb_hat <- replicate(1e5, estimate_coin_bias(n, pb))

# wrangle the results into one data frame
plot_data <- rbind(data.frame(split='A', trial=1:length(pa_hat), p_hat=pa_hat),
                   data.frame(split='B', trial=1:length(pb_hat), p_hat=pb_hat))

# plot the sampling distributions for each split
ggplot(data=plot_data, aes(x=p_hat, fill=split)) +
  geom_histogram(position="identity", binwidth=0.002, alpha=0.5) +
  scale_alpha(guide=F)

# plot the sampling distribution of the difference
qplot(x=pa_hat-pb_hat, geom="histogram", binwidth=0.002) +
  geom_vline(xintercept=pa-pb) +
  geom_vline(xintercept=mean(pa_hat-pb_hat), linetype=2, color="red")

# note that variances add for independent random variables
variance_of_difference <- var(pa_hat - pb_hat)
sum_of_variances <- var(pa_hat) + var(pb_hat)
```

## Confidence intervals
```{r}
# plot 100 confidence intervals by split
plot_data <- transform(plot_data, 
                       LCL = p_hat - 1.96*sqrt(p_hat*(1-p_hat)/n),
                       UCL = p_hat + 1.96*sqrt(p_hat*(1-p_hat)/n))
plot_data <- subset(plot_data, trial <= 100)
ggplot(data=plot_data, aes(x=trial, y=p_hat, linetype=split, position="dodge")) +
  geom_pointrange(aes(ymin=LCL, ymax=UCL)) +
  xlab('') +
  theme(legend.title=element_blank())
```

## Hypothesis testing
```{r}
# construct a null distribution: what would happen if both coins had the same bias (e.g., A and B are the same)?
p0a <- 0.08
p0b <- 0.08
n <- 500
dp0_hat <- replicate(1e5, estimate_coin_bias(n, p0a)) -
           replicate(1e5, estimate_coin_bias(n, p0b))

# run one experiment where there is an underlying difference
pa <- 0.12
pb <- 0.08
dp_hat <- estimate_coin_bias(n, pa) - estimate_coin_bias(n, pb)

# plot the null distribution and see where the observed estimate lies in it
qplot(x=dp0_hat, geom="histogram", binwidth=0.01) +
  geom_vline(xintercept=dp_hat, linetype=2, color="red")

# compare this to our experiment
# how likely is it that we would see an estimate this extreme both coins were identical?
num_as_extreme <- sum(dp0_hat >= dp_hat)
p_value <- num_as_extreme / length(dp0_hat)
p_value
```
Only `r num_as_extreme` out of `r length(dp0_hat)` estimates from two identical coins with p=0.08 would result in an estimate of dp_hat=`r dp_hat` or smaller, corresponding to a p-value of `r p_value`.

# Power calculations
## Computing sample size using built-in R functions
```{r}
# use power.prop.test to compute the sample size you need
power.prop.test(p1=0.08, p2=0.12, sig.level=0.05, power=0.80, alternative="one.sided")
```

## Computing power by direct simulation
```{r}
run_experiment <- function(pa, pb, n, alpha) {
  na <- sum(rbinom(n, 1, pa))
  nb <- sum(rbinom(n, 1, pb))
  test <- prop.test(x = c(na, nb), n = c(n, n), alternative="greater", conf.level = alpha)
  test$p.value < alpha
}

compute_power <- function(pa, pb, n, alpha, r = 1000) {
  mean(replicate(r, run_experiment(pa, pb, n, alpha)))
}

pa <- 0.12
pb <- 0.08
alpha <- 0.05

N <- seq(100,1000, by = 100)
pow <- c()
for (n in N) {
  pow <- c(pow, compute_power(pa, pb, n, alpha))
}

qplot(N, pow)
```
