---
title: "week2_day5_book_exercises"
author: "Christopher Esquivel"
date: '2022-06-10'
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##
Chapter 5 exercises
5.20
a) The equation is y = 0.607975x + 105.2963
```{r}
avg_shoulder_girth <- 108.20
avg_height <- 171.14
sd_shoulder_girth <- 10.37
sd_height <- 9.41
R <- 0.67

b1 <- (sd_height / sd_shoulder_girth) * R
b0 <- 105.2963

```
b) For the slope, for every increase of 10 cm in shoulder girth, we expect an increase in height of
6.07675 cm on average. The intercept describes the average height (105.2963 cm) when the shoulder girth is a small value.

c) R-squared is 0.4489 or 45%. R-squared is the strength of a linear fit.
```{r}
r_sqr <- R * R
r_sqr
```
d) When the shoulder girth is 100 cm, the height predicted by the model is 166.0938
```{r}
y <- 0.607975 * (100) + 105.2963
y
```
e) (100, 160) The residual is -6.1. This means that the linear model is overestimating.
```{r}
e <- 160 - 166.1
e
```
f) When the shoulder girth is 56cm, the predicted height is 139.3429. The average height of a
1 year old is 77-70 cm. The model is not appropiate to predict the height of this child.
```{r}
y <- 0.607975 * (56) + 105.2963
y
```
5.29
a) As height increases, weight increases.
b) weight = -105.0113 + 1.0176 * height. The intercept is the estimated value when height is a small value. The slope indicates that in average, weight increases by 1.0176 as height increases.
c) The null hypothesis is that there is no connection between height and weight. The alternative hypothesis states that as height increases, weight increases too. The p-value is 0.0000. By looking at the graph, I think that the alternative hypothesis is more likely.
d) R-squared is 0.5184 or 52%. It represents the strength of the linear fit.

Lab 3.6.2
```{r}
#library(MASS)
#install.packages("ISLR2")
#library(ISLR2)
head(Boston)
lm.fit <- lm(medv ~ lstat, data = Boston)
summary(lm.fit)
names(lm.fit)
coef(lm.fit)
confint(lm.fit)
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "confidence")
predict(lm.fit, data.frame(lstat = (c(5, 10, 15))), interval = "prediction")
plot(Boston$lstat, Boston$medv)
abline(lm.fit)
abline(lm.fit, lwd = 3)
abline(lm.fit, lwd = 3, col = "red")
plot(Boston$lstat, Boston$medv, col = "red")
plot(Boston$lstat, Boston$medv, pch = 20)
plot(Boston$lstat, Boston$medv, pch = "+")
plot(1:20, 1:20, pch = 1:20)
par(mfrow = c(2, 2))
plot(lm.fit)
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```

