#Week 2 _ Day 2 Exercises
```{r}
library(tidyverse)

theme_set(theme_bw())

set.seed(42)
```
#CHAPTER 4
#code in book
```{r}
library(readr)

ex.1 <- read.csv("ex1.csv")
summary(ex.1)

pop.1 <- read.csv(file = "pop1.csv")
summary(pop.1)

sample(pop.1$height,1)
sample(pop.1$height,1)
sample(pop.1$height,1)

X <- pop.1$height
mean(abs(X - 170) <= 10)

Y <- c(6.3, 6.9, 6.6, 3.4, 5.5, 4.3, 6.5, 4.7, 6.1, 5.3)
abs(Y - 5)
abs(Y - 5) <= 1
mean(abs(Y - 5) <= 1)
```

#Question 4.1.1 - What is the value of p?
Answer - The probabilities of the random variables of Y should add up to 1, so:
p + 2p + 3p + 4p + 5p + 6p = 21p 
21p = 1
p = 1/21

#Question 4.1.2 - P(Y < 3) = ?
Answer - P(Y < 3) = P(Y = 0) + P(Y = 1) + y(Y = 2)
                  = 1(1/21) + 2(1/21) + 3(1/21)
                  = 6/21

#Question 4.1.3 - P(Y = odd) = ?
Answer - P(Y = odd) = P(Y = 1) + P(Y = 3) + P(Y = 5)
                    = 2/21 + 4/21 + 6/21
                    = 12/21
                    
#Question 4.1.4 - P(1 <= Y < 4) = ?
Answer - P(1 <= Y < 4) = P(Y = 1) + P(Y = 2) + P(Y = 3)
                       = 2/21 + 3/21 + 4/21
                       = 9/21
                       
#Question 4.1.5 - P(|Y - 3| < 1.5) = ?
Answer - P(|Y - 3| < 1.5) = P(Y = 2) + P(Y = 3) + P(Y = 4)
                          = 3/21 + 4/21 + 5/21
                          = 12/21
                          
#Question 4.1.6 - E(Y) = ?
Answer - E(Y) = 0 * 1/21 + 1 * 2/21 + 2 * 3/21 + 3 * 4/21 + 4 * 5/21 + 5 * 6/21
              = 0 + 2/21 + 6/21 + 12/21 + 20/21 + 30/21
              = 70/21
```{r}
Y_values <- c(0,1,2,3,4,5)
P_values <- c(1/21, 2/21, 3/21, 4/21, 5/21, 6/21)
E_Y <- sum(Y_values * P_values)
E_Y
```

#Question 4.1.7 - Var(Y) = ?
Answer - Var(Y) = 
```{r}
Var_Y = sum((Y_values - E_Y)^2 * P_values)
Var_Y
```

#Question 4.1.8 - What is the standard deviation of Y.
```{r}
SD_X = sqrt(Var_Y)
SD_X
```

#Question 4.2 - One invests $2 to participate in a game of chance. In this game a 
#coin is tossed three times. If all tosses end up "Head" then the player wins $10.
#Otherwise, the player loses the investment. 

#Question 4.2.1 - What is the probability of winning the game?
Answer - (1/2)^3 = 1/8

#Question 4.2.2 - What is the probability of losing the game?
Answer - 1 - 1/8 = 7/8

#Question 4.2.3 - What is the expected gain for the player that plays this game?
#(Expectation can obtain a negative value.)
Answer - X (= gain) = {8 (win),-2 (loss)}
        P(X) = {1/8, 7/8}
        E(X) = 8 * (1/8) + (-2) * (7/8) 
             = 1 + -14/8
             = -6/8 = -3/4
```{r}
X_values <- c(8,-2)
P_values <- c(1/8, 7/8)
E_X <- sum(X_values * P_values)
E_X
```
             
#CHAPTER 6 
#code in book
```{r}
pnorm(5,2,3) - pnorm(0,2,3)

pnorm((5 - 2)/3) - pnorm((0 - 2)/3)

qnorm(0.975)
qnorm(0.025)

qnorm(0.975,2,3)
qnorm(0.025,2,3)

2 + qnorm(0.975)*3
2 + qnorm(0.025)*3

#third and first quartiles
qnorm(0.75)
qnorm(0.25)

#upper and lower thresholds
qnorm(0.75) + 1.5*(qnorm(0.75)-qnorm(0.25))
qnorm(0.25) - 1.5*(qnorm(0.75)-qnorm(0.25))

2*(1-pnorm(2.697959))

#binomial probability
pbinom(2060,4000,0.5) - pbinom(1939,4000,0.5)

mu <- 4000*0.5
sig <- sqrt(4000*0.5*0.5)
pnorm(2060,mu,sig) - pnorm(1939,mu,sig)

qnorm(0.975,mu,sig)
qnorm(0.025,mu,sig)

qbinom(0.975,4000,0.5)
qbinom(0.025,4000,0.5)

pbinom(6,30,0.3)
pnorm(6,30*0.3,sqrt(30*0.3*0.7))

pnorm(6.5,30*0.3,sqrt(30*0.30*0.7))

ppois(3,2)

pbinom(3,20,0.1)
pnorm(3.5,2,sqrt(20*0.1*0.9))

pbinom(3,200,0.01)
pnorm(3.5,2,sqrt(200*0.01*0.99))

pbinom(3,2000,0.001)
pnorm(3.5,2,sqrt(2000*0.001*0.999))
```

#Question 6.1 - Consider the problem of establishing regulations concerning
#the maximum number of people who can occupy a lift. In particular, we would like
#to assess the probability of exceeding maximal weight when 8 people are allowed
#to use the lift simultaneously and compare that to the probability of allowing
#9 people into the lift.
#Assume that the total weight of 8 people chosen at random follows a normal
#distribution with a mean of 560 kg and a standard deviation of 57 kg. Assume that 
#the total weight of 9 people chosen at random follows a normal distribution with
#a mean of 630 kg and a standard deviation of 61 kg.

#Question 6.1.1 - What is the probability that the total weight of 8 people exceeds
#650 kg?
Answer - P(X > 650) = 1 - P(X <= 650)
X ~ N(560,57^2)
```{r}
1 - pnorm(650,560,57)
```

#Question 6.1.2 - What is the probability thaTthe total weight of 9 people exceeds
#650 kg?
Answer - P(X > 650) = 1 - P(X <= 650)
X ~ N(630,61^2)
```{r}
1 - pnorm(650,630,61)
```

#Question 6.1.3 - What is the central region that contains 80% of distribution of 
#total weight of 8 people?
Answer - P(a <= X <= b) = 0.80
X ~ N(560,57^2) want to find a and b
```{r}
qnorm(0.1,560,57)
qnorm(0.9,560,57)
```

#Question 6.1.4 - What is the central region that contains 80% of distribution of 
#total weight of 9 people?
Answer - P(a <= X <= b) = 0.80
X ~ N(630,61^2) want to find a and b
```{r}
qnorm(0.1,630,61)
qnorm(0.9,630,61)
```