---
title: "Untitled"
author: "Sushobhan Parajuli"
date: '2022-06-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(ISLR2)
```

5.20 If a linear model has a very strong negative relationship with a correlation of -0.97, how much of the variation in the response is explained by
the explanatory variable?
~~> R^2 = (-0.97)^2 = 0.94 percentage of the variation in the response is explained by the explanatory variable.

5.29 
a) positively correlated, linear relationship
b) weight = 1.0176 * height - 105.0113
c) Ho : there is no relationship between height and weight 
   Ha : there is a relationship between height and weight
   p-value is 0.00000
d) 52 percentage of variation in weight is explained variation in height.

3.6 Lab Exercise

```{r}
head(Boston)
```
```{r}
lm.fit <- lm(medv ~ lstat, Boston)
lm.fit
```
```{r}
summary(lm.fit)
```
```{r}
names(lm.fit)
```
```{r}
coef(lm.fit)
```
```{r}
confint(lm.fit)
```
```{r}
predict(lm.fit, data.frame(lstat = (c(5,10,15))),
        interval = "confidence")
```
```{r}
predict(lm.fit, data.frame(lstat = (c(5,10,15))),
        interval = "prediction")
```
```{r}
plot(Boston$lstat, Boston$medv)
abline(lm.fit)
```
```{r}
plot(Boston$lstat, Boston$medv)
abline(lm.fit)
abline(lm.fit, lwd = 3, col = "red")
plot(Boston$lstat, Boston$medv, col = "red")
plot(Boston$lstat, Boston$medv, pch = 20)
plot(Boston$lstat, Boston$medv, pch = "+")
plot(1:20, 1:20, pch = 1:20)
```
```{r}
par(mfrow = c(2,2))
plot(lm.fit)
```
```{r}
plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), residuals(lm.fit))
```
```{r}
plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))
```