---
title: "Week 2 Day 5 Classwork"
output: html_document
date: '2022-06-10'
---

```{r}
library(tidyverse)
```


5.20 Body measurements, Part III. Exercise 5.13 introduces data on shoulder girth and height of a group of individuals. The mean shoulder girth is 108.20 cm with a standard deviation of 10.37 cm. The mean height is 171.14 cm with a standard deviation of 9.41 cm. The correlation between height and shoulder girth is 0.67.

(a) Write the equation of the regression line for predicting height.

  (9.41/10.37) * 0.67 = 0.6079749
  
```{r}
(9.41/10.37) * 0.67
```
  Ans: y - 171.14 = 0.6079749(x - 108.20)

(b) Interpret the slope and the intercept in this context.
  Slope: 0.608
    As the shoulder girth increases by 1 cm, the height of the 
    individuals increases by about 6 cm.
    
  Intercept: 105.36
    The estimated intercept 105.36 cm describes the average height if 
    an individual had 0 cm shoulder girth. This intercept is irrelevant
    to this application because an individual cannot have shoulder 
    girth of 0 cm.

(c) Calculate R2 of the regression line for predicting height from shoulder girth, and interpret it in the context of the application.

```{r}
0.67^2
```
  R^2 = 0.4489
  Approximately 45% of the variability in height can be explained by 
  the shoulder girth of individuals.

(d) A randomly selected student from your class has a shoulder girth of 100 cm. Predict the height of this student using the model.
  y - 171.14 = 0.6079749(100 - 108.20)
```{r}
0.6079749 * (100 - 108.20) + 171.14
```
  height = 166.1546 cm

(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means.
```{r}
160 - 166.1546
```
  The residual of -6.1546 shows that the actual data lies 6.1546cm 
  below the predicted value of the linear model.

(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model to predict the height of this child?

  It is not appropriate to use this linear model to predict the height
  of this child because the data used to create the linear model does 
  not include any data points less than 80 cm for shoulder girth and 
  less than 140cm for height. Thus, it wouldn't be appropriate to use 
  this linear model
  
5.29 Body measurements, Part IV. The scatterplot and least squares summary below show the relationship between weight measured in kilograms and height measured in centimeters of 507 physically active individuals.

          Estimate Std. Error t value Pr(>|t|)
(Intercept) -105.0113 7.5394 -13.93 0.0000
height        1.0176 0.0440 23.13 0.0000

(a) Describe the relationship between height and weight.

  As the height of an individual increases, the weight of an 
  individual also increases.

(b) Write the equation of the regression line. Interpret the slope and intercept in context.
  weight = 1.0176*height - 105.0113
  
  Slope: 1.0176
    As the height of an individual increases by 1 cm, the weight of an 
    individual increases by about 1.0176 kg.
  Intercept: -105.0113
    The estimated intercept -105.0113 kg describes the average weight 
    if an individual was 0 cm tall. This intercept is irrelevant
    to this application because an individual cannot have a height of 
    0 cm or a negative weight.

(c) Do the data provide strong evidence that an increase in height is associated with an increase in weight? State the null and alternative hypotheses, report the p-value, and state your conclusion.

  H0: The true coefficient for an individual's height is 0.
  HA: The true coefficient for an individual's height is not 0.
  
  The p-value = 0, thus we reject the null hypothesis. There is 
  convincing evidence that the true coefficient for an individual's
  height is not 0. That the height of an individual influences their
  weight.

(d) The correlation coefficient for height and weight is 0.72. Calculate R2 and interpret it in context.

```{r}
0.72^2
```
  Approximately 52% of the variability in weight can be explained by 
  the height of individuals.
  
  
Lab 3.6.2
  
```{r}
head(Boston)

lmfit <- lm(medv ~ lstat, data = Boston)
attach(Boston)
lm.fit <- lm(medv ~ lstat)

summary(lm.fit)


names(lm.fit)
coef(lm.fit)

confint(lm.fit)


predict(lm.fit, data_frame(lstat = c(5,10,15)), interval = "confidence")
predict(lm.fit, data.frame(lstat = c(5,10,15)), interval = "prediction")


plot(lstat, medv)
abline(lm.fit)
abline (lm.fit , lwd = 3)
abline (lm.fit , lwd = 3, col = " red ")
plot (lstat , medv , col = " red ")
plot (lstat , medv , pch = 20)
plot (lstat , medv , pch = "+")
plot (1:20, 1:20, pch = 1:20)

par(mfrow = c(2,2))
plot(lm.fit)

plot(predict(lm.fit), residuals(lm.fit))
plot(predict(lm.fit), rstudent(lm.fit))

plot(hatvalues(lm.fit))
which.max(hatvalues(lm.fit))

```

