---
title: "week_2_day_3_textbook_exercises"
output: html_document
date: '2022-06-09'
---

###############Chapter 2 (ISRS)

####2.2 

(a)
```{r}
control_patients_died <- 30/34
control_patients_died

treated_patients_died <-45/69
treated_patients_died

```

(b)
          (i)
          The claims being tested are:
          Null hypothesis: Survival rate is not effected by treatment. Observations of higher survival rate                            occured due to chance. 
          Alternate hypothesis: Survival rate is effected by treatment. Observations of higher survival                                    rate didn't occur due to chance. 
    
          (ii)
          Blank1: 28
          Blank2: 75
          Blank3: 69
          Blank4: 34
          Blant5: 0
          Blank6: -0.23
```{r}
Blank6 <- (1 - control_patients_died) - (1 - treated_patients_died)
Blank6
```
          
          (iii)
          Having a a differnece in proprtion of 23% while due to chance is highly unlikely because                   simulation results express the liklihood of a difference in 23% due to chance being less than 5%.           Therefore we should reject the null hypothesis and accept that treated patients have a higher              chance of survival.
          

          
####2.5

(a)
Null hypothesis: No difference in number of interventions based on dressing "provocatively" or                              "conservatively"
Alternate hypothesis: There is a difference in number of interventions based on dressing "provocatively"                         or conservatively"

(b)

```{r}
provocative_interventions <- 5/20
conservative_interventions <- 15/25

Difference_in_interventions <- provocative_interventions - conservative_interventions
Difference_in_interventions
```
(c)
left tail using left two bins
```{r}
bin1 <- 0.02
bin2 <- 0.005
left_tail <- bin1 + bin2
p_value <- left_tail * 2
p_value
```
Reject null hypothesis since p-value is 0.05. Data suggests strong evidence that interventions occur differently under the two different scenarios



############ Chapter 9 (IST)

####9.2

#Part 1

```{r}
?rnorm

Interval <- rep(0, 10^5)
for(i in 1:10^5){
  group1 <- rnorm(29, 3.5, 3)
  group2 <- rnorm(21, 3.5, 1.5)
  group1_mean <- mean(group1)
  group2_mean <- mean(group2)
  group1_var <- var(group1)
  group2_var <- var(group2)
  
  Interval[i] <- (group1_mean-group2_mean)/sqrt(group1_var/29 + group2_var/21)
  
}
quantile(Interval, c(0.025, 0.975))



```

#Part 2

```{r}
  magnets <- read.csv("magnets.csv")
  group1_mean2 <- mean(magnets$change[1:29])
  group2_mean2 <- mean(magnets$change[30:50])
  group1_var2 <- var(magnets$change[1:29])
  group2_var2 <- var(magnets$change[30:50])
  
  (group1_mean2-group2_mean2)/sqrt(group1_var2/29 + group2_var2/21)

```

