#Week 2 _ Day 2 Exercises
```{r}
library(tidyverse)

theme_set(theme_bw())

set.seed(42)
```

#code in book
```{r}
library(readr)

ex.1 <- read.csv("ex1.csv")
summary(ex.1)

pop.1 <- read.csv(file = "pop1.csv")
summary(pop.1)

sample(pop.1$height,1)
sample(pop.1$height,1)
sample(pop.1$height,1)

X <- pop.1$height
mean(abs(X - 170) <= 10)

Y <- c(6.3, 6.9, 6.6, 3.4, 5.5, 4.3, 6.5, 4.7, 6.1, 5.3)
abs(Y - 5)
abs(Y - 5) <= 1
mean(abs(Y - 5) <= 1)
```

#Question 4.1.1 - What is the value of p?
Answer - The probabilities of the random variables of Y should add up to 1, so:
p + 2p + 3p + 4p + 5p + 6p = 21p 
21p = 1
p = 1/21

#Question 4.1.2 - P(Y < 3) = ?
Answer - P(Y < 3) = P(Y = 0) + P(Y = 1) + y(Y = 2)
                  = 1(1/21) + 2(1/21) + 3(1/21)
                  = 6/21

#Question 4.1.3 - P(Y = odd) = ?
Answer - P(Y = odd) = P(Y = 1) + P(Y = 3) + P(Y = 5)
                    = 2/21 + 4/21 + 6/21
                    = 12/21
                    
#Question 4.1.4 - P(1 <= Y < 4) = ?
Answer - P(1 <= Y < 4) = P(Y = 1) + P(Y = 2) + P(Y = 3)
                       = 2/21 + 3/21 + 4/21
                       = 9/21
                       
#Question 4.1.5 - P(|Y - 3| < 1.5) = ?
Answer - P(|Y - 3| < 1.5) = P(Y = 2) + P(Y = 3) + P(Y = 4)
                          = 3/21 + 4/21 + 5/21
                          = 12/21
                          
#Question 4.1.6 - E(Y) = ?
Answer - E(Y) = 0 * 1/21 + 1 * 2/21 + 2 * 3/21 + 3 * 4/21 + 4 * 5/21 + 5 * 6/21
              = 0 + 2/21 + 6/21 + 12/21 + 20/21 + 30/21
              = 70/21
```{r}
Y_values <- c(0,1,2,3,4,5)
P_values <- c(1/21, 2/21, 3/21, 4/21, 5/21, 6/21)
E_Y <- sum(Y_values * P_values)
E_Y
```

#Question 4.1.7 - Var(Y) = ?
Answer - Var(Y) = 
```{r}
Var_Y = sum((Y_values - E_Y)^2 * P_values)
Var_Y
```

#Question 4.1.8 - What is the standard deviation of Y.
```{r}
SD_X = sqrt(Var_Y)
SD_X
```

#Question 4.2 - One invests $2 to participate in a game of chance. In this game a 
#coin is tossed three times. If all tosses end up "Head" then the player wins $10.
#Otherwise, the player loses the investment. 

#Question 4.2.1 - What is the probability of winning the game?
Answer - (1/2)^3 = 1/8

#Question 4.2.2 - What is the probability of losing the game?
Answer - 1 - 1/8 = 7/8

#Question 4.2.3 - What is the expected gain for the player that plays this game?
#(Expectation can obtain a negative value.)
Answer - X (= gain) = {8,-2}
        P(X) = {1/8, 7/8}
        E(X) = 8 * (1/8) + (-2) * (7/8) 
             = 1 + -14/8
             = -6/8 = -3/4