#Week2 Day4 Exercises
```{r}
library(tidyverse)

theme_set(theme_bw())

set.seed(42)
```
#Exercise 2.2 - Part (a)
Answer - Proportion of patients in treatment group that died: $\frac{30}{34}$
         Proportion of patients in control group that died: $\frac{45}{69}$

#Exercise 2.2 - Part (b) (i)
Answer - 
$H_{0}$ = The amount of patients dying does not depend on whether the patients are receiving a transplant or not.
$H_{A}$ = The amount of patients dying does depend on whether or not the patients are receiving a transplant.

#Exercise 2.2 - Part (b) (ii)
Answer- 
We write alive on 28 cards representing patients who were alive at the end of the study, and dead on 75 cards representing patients who were not. Then, we shuffle these cards and split them into two grpups: one group of size 69 representing treatment, and another group of size 34 representing control. We calculate the difference between the proportion of dead cards in the treatment and control groups (treatment - control) and record this value. We repeat this many times to build a distribution centered at 0%. Lastly, we calculate the fraction of simulations where the simulated differences in proportions are  30/34 - 45/69 = 23%. If this fraction is low, we conclude that it is unlikely to have observed such an outcome by chance and that the null hypothesis should be rejected in favor of the alternative. 

#Exercise 2.2 - Part(c) - 
Answer- 6/103 = 0.06 
- If we take that the threshold value is 0.05, then we fail to reject the null hypothesis since the p-value is greater than alpha. And so there is not enough evidence to support the idea that the amount of patients dying does not depend on whether the patients are receiving a transplant or not. 

#Exercise 2.5 Part(a) - 
Null Hypothesis: The way in which these women were dressed did not affect the way in which observed people decided to intervene. 
Alternative Hypothesis: The way in which these women dressed (provocatively or conservatively) did affect the way the observed people decided to intervene.

#Exercise 2.5 Part(b) - 
Answer - 
5/20 - 15/25 = -0.35

#Exercise 2.5 Part(c) - 
Answer - 
0.005 + 0.02 = 0.025 
p_value = 2*0.025 = 0.05 - Thus, we can reject the null hypothesis. 

#Exercise 9.2 - 

#Exercise 9.2.1 - Assume that the expectation of the measurement is equal to 3.5, regardless of what the type of treatment that the patient received. We take the standard deviation of the measurement for patients the receives an active magnet to be equal to 3 and for those that received the inactive placebo we take it to be equal to 1.5. Assume that the distribution of the measurements is Normal and there are 29 patients in the first group and 21 in the second. Find the interval that contains 95% of the sampling distribution of the statistic.
Answer - 
inactive group ~N(3.5,1.5)
active group ~N(3.5,3)
```{r}
mu1 <- 3.5
sigma1 <- 3
mu2 <- 3.5
sigma2 <- 1.5
test.stat <- rep(0,10^5)
for (i in 1:10^5)
{
  X1 <- rnorm(29, mu1, sigma1)
  X2 <- rnorm(21, mu2, sigma2)
  
  X1.bar <- mean(X1)
  X2.bar <- mean(X2)
  
  X1.var <- var(X1)
  X2.var <- var(X2)
  
  test.stat[i] <- (X1.bar - X2.bar)/sqrt(X1.var/29 + X2.var/21)
}
quantile(test.stat, c(0.025,0.975))
```

#Exercise 9.2.2 - Does the observed value of the statistic, computed for the data frame "magnets", fall inside or outside of the interval that is computed in 1.
Answer - Active = "1", Inactive = "2"
```{r}
library(readr)
magnets <- read.csv(file = "magnets.csv",stringsAsFactors = TRUE)
  
magnets %>% group_by(active)

X1.bar <- mean(magnets$change[1:29])
X2.bar <- mean(magnets$change[30:50])

X1.var <- var(magnets$change[1:29])
X2.var <- var(magnets$change[30:50])

test.stat <- (X1.bar - X2.bar)/sqrt(X1.var/29 + X2.var/21)
test.stat
```