---
title: "week2_day4_book_exercises"
author: "Christopher Esquivel"
date: '2022-06-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###
Chapter 2 exercises
2.2
a)
The proportion of the patients in the treatment group who died = 45/69
The proportion of the patients in the control group who died = 30/34
b)
Null hypothesis = The treatment has no effect on the patients
Alternative hypothesis = The treatment has some effect on the patients
{28, 75, 69, 34, 0, -0.23017}
The treatment had some effect on the patients. It rejects the null hypothesis

2.5
a)
Null Hypothesis = Dressing provocatively or conservatively has no effect on the number
of interventions.
Alternative Hypothesis = Dressing provocatively or conservatively has some effect on the number of
interventions.
b)
5/20 - 15/25 = -7/20 = -0.35
c)
p-value = 0.05
Reject the null hypothesis. Dressing provocatively or conservatively has some effect on the number of interventions.

9.2
1. The interval that contains 95% of the sampling distribution of the statistic = [−2.014838, 2.018435]
```{r}
mu1 <- 3.5
sig1 <- 3
mu2 <- 3.5
sig2 <- 1.5
test.stat <- rep(0, 10^5)
for(i in 1:10^5)
{
  X1 <- rnorm(29, mu1, sig1)
  X2 <- rnorm(21, mu2, sig2)
  X1.bar <- mean(X1)
  X2.bar <- mean(X2)
  X1.var <- var(X1)
  X2.var <- var(X2)
  test.stat[i] <- (X1.bar-X2.bar)/sqrt(X1.var/29 + X2.var/21)
}
quantile(test.stat,c(0.025,0.975))

```
2. 5.985601 does not belong to the interval [−2.014838, 2.018435]
```{r}
magnets <- read.csv("magnets.csv")
x1.bar <- mean(magnets$change[1:29])
x2.bar <- mean(magnets$change[30:50])
x1.var <- var(magnets$change[1:29])
x2.var <- var(magnets$change[30:50])
(x1.bar-x2.bar)/sqrt(x1.var/29 + x2.var/21)

```
