---
title: "test_citibike_predictions"
author: "Shana Mandelbaum"
date: "2024-06-24"
output: html_document
---

```{r}
library(tidyverse)

load("trips.RData")
num_trips <- trips |> 
  group_by(ymd) |> 
  summarise(num_trips = n())

trips_2015 <- inner_join(num_trips, weather, by="ymd")
```

```{r}
data <- "1,2012-01-02,New Year Day
2,2012-01-16,Martin Luther King Jr. Day
3,2012-02-20,Presidents Day (Washingtons Birthday)
4,2012-05-28,Memorial Day
5,2012-07-04,Independence Day
6,2012-09-03,Labor Day
7,2012-10-08,Columbus Day
8,2012-11-12,Veterans Day
9,2012-11-22,Thanksgiving Day
10,2012-12-25,Christmas Day
11,2013-01-01,New Year Day
12,2013-01-21,Martin Luther King Jr. Day
13,2013-02-18,Presidents Day (Washingtons Birthday)
14,2013-05-27,Memorial Day
15,2013-07-04,Independence Day
16,2013-09-02,Labor Day
17,2013-10-14,Columbus Day
18,2013-11-11,Veterans Day
19,2013-11-28,Thanksgiving Day
20,2013-12-25,Christmas Day
21,2014-01-01,New Year Day
22,2014-01-20,Martin Luther King Jr. Day
23,2014-02-17,Presidents Day (Washingtons Birthday)
24,2014-05-26,Memorial Day
25,2014-07-04,Independence Day
26,2014-09-01,Labor Day
27,2014-10-13,Columbus Day
28,2014-11-11,Veterans Day
29,2014-11-27,Thanksgiving Day
30,2014-12-25,Christmas Day
31,2015-01-01,New Year Day
32,2015-01-19,Martin Luther King Jr. Day
33,2015-02-16,Presidents Day (Washingtons Birthday)
34,2015-05-25,Memorial Day
35,2015-07-03,Independence Day
36,2015-09-07,Labor Day
37,2015-10-12,Columbus Day
38,2015-11-11,Veterans Day
39,2015-11-26,Thanksgiving Day
40,2015-12-25,Christmas Day
41,2016-01-01,New Year Day
42,2016-01-18,Martin Luther King Jr. Day
43,2016-02-15,Presidents Day (Washingtons Birthday)
44,2016-05-30,Memorial Day
45,2016-07-04,Independence Day
46,2016-09-05,Labor Day
47,2016-10-10,Columbus Day
48,2016-11-11,Veterans Day
49,2016-11-24,Thanksgiving Day
50,2016-12-25,Christmas Day
51,2017-01-02,New Year Day
52,2017-01-16,Martin Luther King Jr. Day
53,2017-02-20,Presidents Day (Washingtons Birthday)
54,2017-05-29,Memorial Day
55,2017-07-04,Independence Day
56,2017-09-04,Labor Day
57,2017-10-09,Columbus Day
58,2017-11-10,Veterans Day
59,2017-11-23,Thanksgiving Day
60,2017-12-25,Christmas Day
61,2018-01-01,New Year Day
62,2018-01-15,Martin Luther King Jr. Day
63,2018-02-19,Presidents Day (Washingtons Birthday)
64,2018-05-28,Memorial Day
65,2018-07-04,Independence Day
66,2018-09-03,Labor Day
67,2018-10-08,Columbus Day
68,2018-11-12,Veterans Day
69,2018-11-22,Thanksgiving Day
70,2018-12-25,Christmas Day
71,2019-01-01,New Year Day
72,2019-01-21,Martin Luther King Jr. Day
73,2019-02-18,Presidents Day (Washingtons Birthday)
74,2019-05-27,Memorial Day
75,2019-07-04,Independence Day
76,2019-09-02,Labor Day
77,2019-10-14,Columbus Day
78,2019-11-11,Veterans Day
79,2019-11-28,Thanksgiving Day
80,2019-12-25,Christmas Day
81,2020-01-01,New Year Day
82,2020-01-20,Martin Luther King Jr. Day
83,2020-02-17,Presidents Day (Washingtons Birthday)
84,2020-05-25,Memorial Day
85,2020-07-03,Independence Day
86,2020-09-07,Labor Day
87,2020-10-12,Columbus Day
88,2020-11-11,Veterans Day
89,2020-11-26,Thanksgiving Day
90,2020-12-25,Christmas Day"

holidays <- read.csv(text = data, header = FALSE, 
                     col.names = c("id", "date", "holiday"))

trips_2015 <- trips_2015 |> 
  mutate(day_of_week = wday(ymd, label = TRUE)) |> 
  mutate(is_weekend = (day_of_week == "Sat" | 
                         day_of_week == "Sun")) |> 
  mutate(is_holiday = ifelse(ymd %in% holidays$date, TRUE, FALSE)) |> 
  mutate(avg_temp = (tmin + tmax)/2) |> 
  mutate(avg_temp = avg_temp/10)
```

```{r}
load("final_model.RData")
```


```{r}

#model4 <- lm(num_trips ~ poly(avg_temp, 5, raw = T)*prcp + is_weekend + is_holiday,data = trips_2015)
model4
Test_RMSE<- sqrt(mean((predict(model4, trips_2015) -
                         trips_2015$num_trips)^2))
Test_RMSE
```
```{r}
load("BestModel.RData")

trips_2015_test <- trips_2015 |> 
  rename(tavg = avg_temp) |> 
  rename(IsWeekday = is_weekend) |> 
  rename(IsHoliday = is_holiday) |> 
  mutate(IsWeekday = ifelse(wday(ymd, week_start = 1) >=6, FALSE, TRUE))

Test_RMSE<- sqrt(mean((predict(model, trips_2015_test) -
                         trips_2015_test$num_trips)^2))
Test_RMSE
```


```{r}
trips_2015_preds <- trips_2015 |> 
  mutate(preds = predict(model4, trips_2015))

trips_2015_preds |> 
  ggplot() +
  geom_point(aes(x = ymd, y = num_trips), color = "red") +
  geom_line(aes(x = ymd, y = preds))

trips_2015_preds |> 
  ggplot(aes(x = preds, y = num_trips)) +
  geom_point() +
  geom_abline(linetype = "dashed") +
  xlab('Predicted') +
  ylab('Actual')
```
Originally, the errors that Salma and I were getting were huge. We found that the way that the temp was being calculated was different than in our original model. We divided the temp by 10 to obtain the values we were working with on Friday.
Compared to my model on the test data, the model on the 2015 worked worse. The error on the test data was 3724 and the error on the 2015 data was 8308. This seems to be because of a big increase in rides during the second half of the year that did not occur in 2014. 
Additionally, Salma and I had a few variables named differently so we needed to change our data to run the model.
