---
title: "Testing model"
author: "Nabin Oli"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(here)
library(lubridate)
library(modelr)
```

## Load Data

```{r, data loading and transformation}
trips = read_tsv(here("week4","trips_per_day_2015.tsv"))
weather = read_csv(here("week4","weather_2015.csv"))
weather <- weather%>%mutate(TMAX=TMAX/10,TMIN = TMIN/10)
weather <- rename(weather,ymd=DATE,tmax = TMAX, tmin = TMIN, snow = SNOW, snwd = SNWD, prcp = PRCP)
test_data <- inner_join(trips, weather, by="ymd")
```

## Feature extraction
```{r, feature extraction from test data}
test_data<-test_data%>%
select(num_trips,ymd,tmax,tmin,snow,snwd,prcp)%>%
mutate(day = wday(ymd,week_start = 1))%>%
mutate(day_type = ifelse(day %in% c(6,7),"weekend","weekdays" ))%>%
mutate(sigprcp = (prcp>=0.10))%>%
mutate(avgtemp =log((tmax+tmin)/2))%>%
mutate(month = month(ymd))%>%
mutate(seasonality = ifelse(month %in%c(1,2,3,11,12),"winter","summer"))%>%
mutate(fold = 1)%>%
mutate(date =1 )
```

## Loading model and testing
```{r, test stats}
model = load(here("week3","model.Rdata"))
summary(final_model)
rmse<- rmse(final_model,test_data)
Rsquared <-rsquare(final_model,test_data)
test_changed <- test_data%>%filter(month != 10 & month != 11 & month != 9)
rmse_exlcude <- rmse(final_model,test_changed)
rsquared_exclude <- rsquare(final_model,test_changed)
```

## Ploting agianst test data points
```{r, ploting}
add_predictions(test_data,final_model)%>%
ggplot(aes(x=ymd,y=pred,color = 'Predicted'))+
geom_line()+
geom_point(aes(x=ymd,y=num_trips,color = 'Actual'))+
labs(
    color = "Label",
    x= "Days",
    y="Number of rides"
)

add_predictions(test_changed,final_model)%>%
ggplot(aes(x=ymd,y=pred,color = 'Predicted'))+
geom_point()+
geom_point(aes(x=ymd,y=num_trips,color = 'Actual'))+
labs(
    color = "Label",
    x= "Days",
    y="Number of rides"
)
```

## Conclusion
`r rmse` This was my final RMSE score which is far from my initial guess of 3500-4000.
R^2^ value is `r Rsquared`. Looking at predicted value and actual value graph, there are a lot of outliers in September, October and November. 
Just to see how well model fits if I exclude data from these months I my new RMSE is `r rmse_exlcude` and R^2^ is `r rsquared_exclude`, which is good but still worse than 
train-test split scroes. 