---
title: "Test Citibike Predictions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(lubridate)  # For handling dates

```

```{r}
# Load the 2015 trips and weather data
load("trips_2015.RData")

# Ensure trips and weather data are correctly loaded
head(trips)
head(weather)

# Create the trips_per_day_2015 data frame by merging trips and weather data
trips_per_day_2015 <- trips %>%
  mutate(ymd = as.Date(starttime)) %>%
  group_by(ymd) %>%
  summarise(num_trips = n()) %>%
  left_join(weather, by = "ymd")

# Add holiday indicator
trips_per_day_2015 <- trips_per_day_2015 %>%
  mutate(holiday = ifelse(ymd %in% c(
    '2015-01-01', '2015-01-19', '2015-02-16', '2015-05-25',
    '2015-07-03', '2015-09-07', '2015-10-12', '2015-11-11',
    '2015-11-26', '2015-12-25'), 1, 0))

# Add isweekend indicator

trips_per_day_2015$isweekend <- as.numeric(wday(trips_per_day_2015$ymd) %in% c(1,7))

# Inspect the merged data
head(trips_per_day_2015)


```

```{r}
load("../week3/best_model.RData")

k <- 5

# Inspect the loaded model
summary(model)

```

```{r}
# Make predictions using the loaded model

model <- lm(num_trips ~ holiday * isweekend * poly(tmin, k, raw = TRUE) + 
                           poly(tmax, k, raw = TRUE) + prcp + snwd + snow, 
            data = trips_per_day_2015)

trips_per_day_2015 <- trips_per_day_2015 %>%
  add_predictions(model, var = "pred")

# Check for missing values
missing_values <- trips_per_day_2015 %>%
  summarize(missing_num_trips = sum(is.na(num_trips)),
            missing_pred = sum(is.na(pred)))

print(missing_values)

# Inspect the first few rows with predictions
head(trips_per_day_2015)

# Compute the RMSE, ignoring any rows with missing values
rmse <- trips_per_day_2015 %>%
  filter(!is.na(pred) & !is.na(num_trips)) %>%
  summarize(rmse = sqrt(mean((pred - num_trips)^2)))

print(rmse)
summary

```
```{r}

ggplot(trips_per_day_2015, aes(x = num_trips, y = pred)) +
  geom_point(alpha = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Predicted vs Actual Number of Trips", 
       x = "Actual Number of Trips", 
       y = "Predicted Number of Trips") +
  theme_minimal()
```

## Summary and Thoughts

### Model Performance in 2014 vs. 2015
- **Data Size**: The small dataset size led to higher RMSE due to limited generalization capability.
- **Overfitting**: The model overfitted the training data, capturing noise instead of the actual signal, resulting in poor performance on the 2015 data.

### Approach to Finding the Best Model_2014
- **Polynomial Terms**: Explored different polynomial degrees for temperature variables using `regsubsets` from the `leaps` library. Visualized the results to determine the optimal polynomial degree.
