---
title: "Replication of \"Anti-Immigrant Rhetoric and ICE Reporting Interest: Evidence from a Large-Scale Study of Web Search Data\""
date: '`r Sys.time()`'
output:
  html_document:
    #code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(tidyverse)

# set plot theme
theme_set(theme_bw())
```

Import the immigration reporting trends from Google Trends. 
Cleaned and formatted the data frame and plotted
```{r Immigration Reporting Trends}
# Import immigrant reporting trend data
reporting <- read_csv("immigrant_reporting.csv")
reporting$Month <- as.Date(paste0(reporting$Month, "-01"), format = "%Y-%m-%d")
reporting <- reporting |> 
  filter(Trend != 0) # Maybe drop 0s
reporting <- reporting |> 
  mutate(President = case_when(
    between(Month, as.Date("2001-01-01"), as.Date("2009-12-31")) ~ "Bush",
    between(Month, as.Date("2009-1-1"), as.Date("2017-1-19")) ~ "Obama",
    TRUE ~ "Trump"
  ))

# Reporting Trends Graph by Presidential Term
reporting |> 
  ggplot(aes(x = Month, y = Trend, color = President)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(
    x = "",
    y = "Google Trends"
  ) +
  ggtitle("Reporting Trends")
```


Import the immigration crime trends from Google Trends. 
Cleaned and formatted the data frame and plotted
```{r Immigration Crime Trends}
# Import immigrant crime trend data
crime <- read_csv("immigrant_crime.csv")
crime$Month <- as.Date(paste0(crime$Month, "-01"), format = "%Y-%m-%d")
crime <- crime |> 
  filter(Trend != 0) # Maybe drop 0s
crime <- crime |> 
  mutate(President = case_when(
    between(Month, as.Date("2001-01-01"), as.Date("2009-12-31")) ~ "Bush",
    between(Month, as.Date("2009-1-1"), as.Date("2017-1-19")) ~ "Obama",
    TRUE ~ "Trump"
  ))

# Crime Trends Graph by Presidential Term
crime |> 
  ggplot(aes(x = Month, y = Trend, color = President)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(
    x = "",
    y = "Google Trends"
  ) +
  ggtitle("Crime Trends")
```


Import the immigration welfare trends from Google Trends. 
Cleaned and formatted the data frame and plotted
```{r Immigration Welfare Trends}
# Import immigrant welfare trend data
welfare <- read_csv("immigrant_welfare.csv")
welfare$Month <- as.Date(paste0(welfare$Month, "-01"), format = "%Y-%m-%d")
welfare <- welfare |> 
  filter(Trend != 0) # Maybe drop 0s
welfare <- welfare |> 
  mutate(President = case_when(
    between(Month, as.Date("2001-01-01"), as.Date("2009-12-31")) ~ "Bush",
    between(Month, as.Date("2009-1-1"), as.Date("2017-1-19")) ~ "Obama",
    TRUE ~ "Trump"
  ))

# Welfare Trends Graph by Presidential Term
welfare |> 
  ggplot(aes(x = Month, y = Trend, color = President)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(
    x = "",
    y = "Google Trends"
  ) +
  ggtitle("Welfare Trends")
```

Compare to Snapshotted Google Trends data
```{r compare to snapshotted data}

# Reporting Graph
reporting_snap <- read_csv("project/from_replication_files/google_trends_report.csv")

reporting_snap <- reporting_snap |> 
  unite("date", year, month, sep = "-")

reporting_snap$date <- as.Date(paste0(reporting_snap$date, "-01"), format = "%Y-%m-%d")

reporting_snap <- reporting_snap |> 
  mutate(President = case_when(
    between(date, as.Date("2001-01-20"), as.Date("2009-01-20")) ~ "Bush",
    between(date, as.Date("2009-1-20"), as.Date("2017-1-20")) ~ "Obama",
    TRUE ~ "Trump"
  ))

reporting_snap |> 
  ggplot(aes(x = date, y = search, color = President)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(
    x = "",
    y = "Google Trends"
  ) +
  ggtitle("Reporting Trends")



# Crime Graph
crime_snap <- read_csv("project/from_replication_files/google_trends_crime.csv")

crime_snap <- crime_snap |> 
  unite("date", year, month, sep = "-")

crime_snap$date <- as.Date(paste0(crime_snap$date, "-01"), format = "%Y-%m-%d")

crime_snap <- crime_snap |> 
  mutate(President = case_when(
    between(date, as.Date("2001-01-20"), as.Date("2009-01-20")) ~ "Bush",
    between(date, as.Date("2009-1-20"), as.Date("2017-1-20")) ~ "Obama",
    TRUE ~ "Trump"
  ))

crime_snap |> 
  ggplot(aes(x = date, y = search, color = President)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(
    x = "",
    y = "Google Trends"
  ) +
  ggtitle("Crime Trends")



# Welfare Graph
welfare_snap <- read_csv("project/from_replication_files/google_trends_welfare.csv")

welfare_snap <- welfare_snap |> 
  unite("date", year, month, sep = "-")

welfare_snap$date <- as.Date(paste0(welfare_snap$date, "-01"), format = "%Y-%m-%d")

welfare_snap <- welfare_snap |> 
  mutate(President = case_when(
    between(date, as.Date("2001-01-20"), as.Date("2009-01-20")) ~ "Bush",
    between(date, as.Date("2009-1-20"), as.Date("2017-1-20")) ~ "Obama",
    TRUE ~ "Trump"
  ))

welfare_snap |> 
  ggplot(aes(x = date, y = search, color = President)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(
    x = "",
    y = "Google Trends"
  ) +
  ggtitle("Welfare Trends")



# Report Table

reporting_snap$President <- relevel(factor(reporting_snap$President), ref = "Obama")

reporting_snap_regression <- lm(search ~ date + President, data = reporting_snap)
summary(reporting_snap_regression)




# Crime Table

crime_snap$President <- relevel(factor(crime_snap$President), ref = "Obama")

crime_snap_regression <- lm(search ~ date + President, data = crime_snap)
summary(crime_snap_regression)



# Welfare Table

welfare_snap$President <- relevel(factor(welfare_snap$President), ref = "Obama")

welfare_snap_regression <- lm(search ~ date + President, data = welfare_snap)
summary(welfare_snap_regression)

```



Replicating Table 3

Crime Table
```{r create regression lines}
# Crime Table

crime$President <- relevel(factor(crime$President), ref = "Obama")

crime_regression <- lm(Trend ~ Month + President, data = crime)
summary(crime_regression)

```


Welfare Table
```{r create regression lines}
# Welfare Table

welfare$President <- relevel(factor(welfare$President), ref = "Obama")

welfare_regression <- lm(Trend ~ Month + President, data = welfare)
summary(welfare_regression)

```


Report Table
```{r create regression lines}
# Report Table

reporting$President <- relevel(factor(reporting$President), ref = "Obama")

reporting_regression <- lm(Trend ~ Month + President, data = reporting)
summary(reporting_regression)

```


```{r load topic model and setup}
# Loading the topic model for the news transcripts
library(stm)
load("TopicModel.RData")
document_topics <- make.dt(immigrFit, meta = out$meta)
topic_terms <- t(exp(immigrFit$beta$logbeta[[1]]))
rownames(topic_terms) <- out$vocab
colnames(topic_terms) <- sprintf("Topic%d", 1:ncol(topic_terms))
```

Replication for Figure 2
```{r replicate figure 2}

# CNN
cnn_monthly_duration <- document_topics |> 
  filter(channel == "cnn") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(daily_duration = sum(duration)) |> 
  select(year_month, daily_duration, time) |> 
  distinct()

cnn_monthly_duration$year_month <- as.Date(paste0(cnn_monthly_duration$year_month, "-01"), format = "%Y-%m-%d")

# FOX
fox_monthly_duration <- document_topics |> 
  filter(channel == "fox") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(daily_duration = sum(duration)) |> 
  select(year_month, daily_duration, time) |> 
  distinct()
  

fox_monthly_duration$year_month <- as.Date(paste0(fox_monthly_duration$year_month, "-01"), format = "%Y-%m-%d")

# MSNBC
msnbc_monthly_duration <- document_topics |> 
  filter(channel == "msnbc") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(daily_duration = sum(duration)) |> 
  select(year_month, daily_duration, time) |> 
  distinct()

msnbc_monthly_duration$year_month <- as.Date(paste0(msnbc_monthly_duration$year_month, "-01"), format = "%Y-%m-%d")

combined_data <- rbind(
  transform(cnn_monthly_duration, channel = "cnn"),
  transform(fox_monthly_duration, channel = "fox"),
  transform(msnbc_monthly_duration, channel = "msnbc")
)

combined_data |> 
  ggplot(aes(x=year_month, y=daily_duration, color=channel, group = interaction(channel, time))) +
  geom_point() +
  geom_smooth(se = F) +
  geom_vline(xintercept = ymd("2015-06-01"), linetype="dashed") +
  geom_vline(xintercept = ymd("2017-01-01"), linetype="dashed") +
  labs(
    x = "",
    y = "Num Monthly Immigration Segs"
  )
```

Replication for Figure 3

Crime Topic Graph
```{r replicate figure 3 for crime data}

# Topic 1 and 3 - CRIME

# CNN Crime
cnn_crime_topic <- document_topics |> 
  filter(channel == "cnn") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(topic_one_sum = sum(Topic1), topic_three_sum = sum(Topic3)) |>
  mutate(crime_topic = topic_one_sum + topic_three_sum) |> 
  select(year_month, crime_topic, time) |> 
  distinct()
cnn_crime_topic$year_month <- as.Date(paste0(cnn_crime_topic$year_month, "-01"), format = "%Y-%m-%d")

# FOX Crime
fox_crime_topic <- document_topics |> 
  filter(channel == "fox") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(topic_one_sum = sum(Topic1), topic_three_sum = sum(Topic3)) |>
  mutate(crime_topic = topic_one_sum + topic_three_sum) |> 
  select(year_month, crime_topic, time) |> 
  distinct()
fox_crime_topic$year_month <- as.Date(paste0(fox_crime_topic$year_month, "-01"), format = "%Y-%m-%d")

# MSNBC Crime
msnbc_crime_topic <- document_topics |> 
  filter(channel == "msnbc") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(topic_one_sum = sum(Topic1), topic_three_sum = sum(Topic3)) |>
  mutate(crime_topic = topic_one_sum + topic_three_sum) |> 
  select(year_month, crime_topic, time) |> 
  distinct()
msnbc_crime_topic$year_month <- as.Date(paste0(msnbc_crime_topic$year_month, "-01"), format = "%Y-%m-%d")


combined_data <- rbind(
  transform(cnn_crime_topic, channel = "cnn"),
  transform(fox_crime_topic, channel = "fox"),
  transform(msnbc_crime_topic, channel = "msnbc")
)

combined_data |> 
  ggplot(aes(x=year_month, y=crime_topic, color=channel, group = interaction(channel, time))) +
  geom_point() +
  geom_smooth(se = F) +
  geom_vline(xintercept = ymd("2015-06-01"), linetype="dashed") +
  geom_vline(xintercept = ymd("2017-01-01"), linetype="dashed") +
  labs(
    x = "",
    y = "Immigrant Crime News Coverage"
  )

```

Welfare Topic Graph
```{r replicate figure 3 for welfare data}
# Topic 13 - WELFARE

# CNN Welfare
cnn_welfare_topic <- document_topics |> 
  filter(channel == "cnn") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(welfare_topic = sum(Topic13)) |>
  select(year_month, welfare_topic, time) |> 
  distinct()
cnn_welfare_topic$year_month <- as.Date(paste0(cnn_welfare_topic$year_month, "-01"), format = "%Y-%m-%d")

# FOX Welfare
fox_welfare_topic <- document_topics |> 
  filter(channel == "fox") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(welfare_topic = sum(Topic13)) |>
  select(year_month, welfare_topic, time) |> 
  distinct()
fox_welfare_topic$year_month <- as.Date(paste0(fox_welfare_topic$year_month, "-01"), format = "%Y-%m-%d")

# MSNBC Welfare
msnbc_welfare_topic <- document_topics |> 
  filter(channel == "msnbc") |> 
  mutate(date = as.Date(date)) |> 
  mutate(year_month = format(date, format = "%Y-%m")) |> 
  group_by(year_month) |> 
  mutate(welfare_topic = sum(Topic13)) |>
  select(year_month, welfare_topic, time) |> 
  distinct()
msnbc_welfare_topic$year_month <- as.Date(paste0(msnbc_welfare_topic$year_month, "-01"), format = "%Y-%m-%d")


combined_data <- rbind(
  transform(cnn_welfare_topic, channel = "cnn"),
  transform(fox_welfare_topic, channel = "fox"),
  transform(msnbc_welfare_topic, channel = "msnbc")
)

combined_data |> 
  ggplot(aes(x=year_month, y=welfare_topic, color=channel, group = interaction(channel, time))) +
  geom_point() +
  geom_smooth(se = F) +
  geom_vline(xintercept = ymd("2015-06-01"), linetype="dashed") +
  geom_vline(xintercept = ymd("2017-01-01"), linetype="dashed") +
  labs(
    x = "",
    y = "Immigrant Welfare News Coverage"
  )

```

Replicate Table 4
```{r Replicate Table 4}

# Immigration segs
combined_data <- rbind(
  transform(cnn_monthly_duration, channel = "cnn"),
  transform(fox_monthly_duration, channel = "fox"),
  transform(msnbc_monthly_duration, channel = "msnbc")
)

# Immigration + Crime Coverage
combined_data <- rbind(
  transform(cnn_crime_topic, channel = "cnn"),
  transform(fox_crime_topic, channel = "fox"),
  transform(msnbc_crime_topic, channel = "msnbc")
)

# Immigration + Welfare Coverage
combined_data <- rbind(
  transform(cnn_welfare_topic, channel = "cnn"),
  transform(fox_welfare_topic, channel = "fox"),
  transform(msnbc_welfare_topic, channel = "msnbc")
)

# Trump admin
crime

# 
```

