---
title: "citibike predictions 2015 and 2020"
output: html_document
date: '2022-06-20'
---

The main challenges I faced where downloading the 2020 trips data, building the model and trying to predict the 2015 and 2020 number of trips. 

For 2015 trips I found that the rmse of my model is around 7878 number of trips 

Here is the code that predicts it

```{r}
#2015 sesction
load("trips_2015.RData")
load("best.RData")


trip_predictions_2015 <- trips_test %>% add_predictions(model)
rmse_2015 <- sqrt(mean((predict(model, trips_test_final) - trips_test_final$num_trips)^2))
rmse_2015 # 7878

```


For 2020 trips I found that the rmse of my model is around 36154

Here is the code that predicts it

```{r}
#2020 section 
load("trips_2020.RData")


trip_predictions_2020 <- trips_test_2020 %>% add_predictions(model)
rmse_2020 <- sqrt(mean((predict(model, trip_predictions_2020) - trip_predictions_2020$num_trips_2020)^2))

rmse_2020 #36154
```

What I can concluded is that my model overfit but at the same time I think that the rmse shows that since 2020 was such a different year, my model was not able to predict that (i.e. COVID) and therefore the predictions of number of trips were very off. 