---
title: "day1_week3_exercises.Rmd"
output: html_document
date: '2022-06-13'
---

Reproduce the table in ISRS 5.29 using the original dataset in body.dat.txt
```{r}
library(tidyverse)
library(scales)

library(modelr)
options(na.action = na.warn)

library(broom)

theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)
body_dat <- read.table("body.dat.txt", sep='')

height_weight <- select(body_dat, "weight" = V23, "height" = V24)
form <- as.formula(weight ~ height)
model <- lm(form, height_weight)

tidy(model)
glance(model)

```

Labs 3.6.3 through 3.6.6 of Intro to Statistical Learning
```{r}
#3.6.3
library(ISLR2)
lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)

lm.fit <- lm(medv ~ ., data = Boston)
summary(lm.fit)

library(car)
vif(lm.fit)

lm.fit1 <- lm(medv ~ . - age, data = Boston)
summary(lm.fit1)

#3.6.4
summary(lm(medv ~ lstat*age, data = Boston))

#3.6.5
lm.fit2 <- lm(medv ~ lstat + I(lstat^2), data = Boston)
summary(lm.fit2)

lm.fit <- lm(medv ~ lstat, data = Boston)
anova(lm.fit, lm.fit2)

par(mfrow = c(2,2))
plot(lm.fit2)

lm.fit5 <- lm(medv ~ poly(lstat, 5), data = Boston)
summary(lm.fit5)

summary(lm(medv ~ log(rm), data = Boston))

#3.6.6
head(Carseats)

lm.fit <- lm(Sales ~ . + Income:Advertising + Price:Age, data = Carseats)
summary(lm.fit)

attach(Carseats)
contrasts(ShelveLoc)
```

ISRS 
6.1 
a) weight = 123.05 - 8.94*smoke 

b) If the mother is a smoker, the baby is 8.94 units of weight (ounces) lighter
non-smoker mother, baby's birth weight --> weight = 123.05
smoker mother, baby's birth weight --> weight = 123.05 - 8.94 = 114.11

c) the p-value in the table is close to zero so we reject the null and we state that there is a statistically significant relationship between the average birth weight and smoking

6.2
a) weight = 120.07 - 1.93*parity 

b) According to the slope, if the child is not a first born, his/her weight will decrease by 1.93 ounces, otherwise it will stay the same.
firstborn birth weight --> weight = 120.07 
not a firstborn birth weight --> weight = 120.07 - 1.93 = 118.14

c) Since the p-value is close to 10% (as shown in the table), we fail to reject the null, 
therefore there is no relationship between the average birth weight and parity

6.3
a) bwt = -80.41 + 0.44xgestation - 3.33xparity - 0.01xage + 1.15xheight + 0.05xweight - 8.40xsmoke
b) For an increase of one unit (a day) in gestation, the total btw of the baby increases by 0.44 units
   For an increase of one unit in the mother's age, the total btw of the baby decreases by 0.01 units
c) Parity is probably correlated to another variable (for example the mother's age), so before, when we were only considering that and not other variable, there was probably a bias in estimating it.
d) 
```{r}
actual_value <- 120
predicted_value <-  -80.41 + 0.44*284 - 3.33*0 - 0.01*27 + 1.15*62 + 0.05*100 - 8.40*0 #120.58
```

e)
R**2 = amount of variability in the response that was explained by the model
R**2adj = 1 - (Var(ei)/(n-k-1))/(Var(yi)/(n-1))
```{r}
r_sq <- 1 - 249.28/332.57
r_sq_Adj <- 1 - ((249.28/332.57) * ((1236 - 1)/(1236 - 6 -1)))
r_sq
r_sq_Adj
```
```{r}
library(tidyverse)
library(scales)
library(modelr)
options(na.action = na.warn)
library(broom)
bw <- read.table("babyweights.txt", sep="")

form1 <- as.formula(bwt ~ smoke)
model1 <- lm(form1, bw)
tidy(model1)

form2 <- as.formula(bwt ~ parity)
model2 <- lm(form2, bw)
tidy(model2)

form3 <- as.formula(bwt ~ .)
model3 <- lm(form3, bw)
tidy(model3)

```

ISL

```{r}
#5.3.1
library(ISLR2)
set.seed(1)
train <- sample(392, 196)
lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = train)

attach(Auto)
mean((mpg - predict(lm.fit, Auto))[-train]^2)

lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)


lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)

set.seed(2)
train <- sample(392, 196)
lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = train)

attach(Auto)
mean((mpg - predict(lm.fit, Auto))[-train]^2)

lm.fit2 <- lm(mpg ~ poly(horsepower, 2), data = Auto, subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)


lm.fit3 <- lm(mpg ~ poly(horsepower, 3), data = Auto, subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)

#5.3.2
glm.fit <- glm(mpg ~ horsepower, data = Auto)
coef(glm.fit)

lm.fit <- lm(mpg ~ horsepower, data = Auto)
coef(lm.fit)

library(boot)
glm.fit <- glm(mpg ~ horsepower, data = Auto)
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta

cv.error <- rep(0, 10)
for(i in 1:10){
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}
cv.error

#5.3.3
set.seed(17)
cv.error.10 <- rep(0, 10)
for(i in 1:10){
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K=10)$delta[1]
}
cv.error.10
```

