---
title: "lr_exercise"
output:
  html_document: default
  pdf_document: default
date: "2023-06-13"
---

```{r setup, include=FALSE}
library(scales)
library(broom)
library(modelr)
library(tidyverse)

options(na.action = na.warn)

theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)
```

## Linear Regression 5.29
ISRS 5.29
```{r 5.29}
data <- read.table(file = "body.dat.txt")
#head(data)


#find heights: 
heights=data["V24"]

#find weights: 
weights=data["V23"]

df_data=data.frame(heights,weights)
names(df_data)=c("heights","weights")


ggplot(df_data,mapping=aes(x=heights,y=weights))+
  geom_point(color="blue",alpha=0.5)+
  xlab("Heights(in cm)")+
  ylab("Weights(in kg)")

```

## ISRS 6.1,6.2,6.3

#6.1 baby weight, partI: avg birth weight vs smoke
(a) bwt=123.05-8.94*smoke, where smoke takes the values of 0 and 1
(b) bwt for smoker mom: 123.05-8.94=114.11
bwt for nonsmoker mom: 123.05
(c) there is a significant relationship because the p values is really small
```{r exercises6.1}
data1=read.table(file = "babyweights.txt",header=TRUE)
#head(data1)

#duplicate the table 
model=lm(bwt ~ smoke,data1)
summary(model)

#compute the avg birth weight for smoker and nonsmoker mom
#check whether there is a statiscally significant relationship


data1  %>% group_by(smoke) %>% 
  summarize(avg_wbt=mean(bwt)) %>% 
  filter(smoke %in% c(1,0))  %>%
  ggplot(aes(x=smoke,y=avg_wbt))+
  geom_col()+
  xlab("smoke")+
  ylab("bwt")


data1 %>% filter(smoke %in% c(1,0))  %>%
  ggplot(aes(x=smoke,y=bwt))+
  geom_hline(yintercept=123.05)+
  geom_hline(yintercept=114.11)+
  geom_point()
```
#6.2

parity(0 if the child is the first born and 1 otherwise) vs bwt
(a) bwt=120.07-1.93*parity
(b)first born: 120.7
others: 120.7-1.93
(c) there is not a significant relationship between the avg birth weight and parity b/c p value is big
```{r exercises6.2}
model1=lm(bwt~parity,data1)
summary(model1)
```
#6.3 baby weight3
gestation, parity, age, height, weight, smoke
(1) bwt=-80.41+0.44*gestation-3.33*parity-0.01*age+1.15*height+0.05weight-8.4smoke
(2)as the length of preganancy in days(gestation) increases, bwt increases (significant relationship b/c small pvalue);
as the age increases, bwt decreases (not very strong relationship bc large pvalue)
(3)diff variables may interfere

```{r exercises6.3}
model2=lm(bwt~gestation+parity+age+height+weight+smoke,data1)
summary(model2)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
