<<<<<<< HEAD
---
title: "Movielens"
date: '`r Sys.time()`'
output:
  html_document:
    #code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(scales)
library(tidyverse)
library(knitr)

# set plot theme
theme_set(theme_bw())
```

# Load and preview data

Read data from the `ratings.csv` file
```{r load-data}
ratings <- read_csv('ratings.csv',
                    col_names = c('user_id','movie_id','rating','timestamp'))
```

Loaded `r format(object.size(ratings), units="Mb")` of ratings data, containing `r format(nrow(ratings), big.mark = ",")` ratings. Here's a preview:
```{r preview-data}
head(ratings) %>% kable()
```

# Summary statistics

```{r dist-ratings}
# plot the distribution of rating values https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=26
#okay, so num of ratings along with avg rating
#first I probably want timestamp gone
#reference https://www.listendata.com/2015/06/r-keep-drop-columns-from-data-frame.html
ratings = subset(ratings, select = -timestamp )
#okay, i group by movies, i can then grab the average of each one
#AND the num of rows. which ought to be ratings
group_by(ratings, rating) %>% ggplot(aes(x = rating)) + geom_histogram(bins = 10) + xlab('Rating') + ylab('Number of ratings')
```

## Per-movie stats

```{r aggregate-by-movie}
# aggregate ratings by movie, computing mean rating and number of ratings
# hint: use the n() function for easy counting within a group
group_by(ratings, movie_id) %>% summarise(avg = mean(rating), count = n())
```

```{r dist-movie-popularity}
# plot distribution of movie popularity (= number of ratings the movie received)
# hint: try scale_x_log10() for a logarithmic x axis
group_by(ratings, movie_id) %>% summarise(popularity = n()) %>% ggplot(aes(x = popularity)) + geom_histogram() + scale_x_log10()

```

```{r dist-mean-ratings-by-movie}
# plot distribution of mean ratings by movie https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=28
# hint: try geom_histogram and geom_density
group_by(ratings, movie_id) %>% summarise(avg = mean(rating)) %>% 
ggplot(aes(x = avg)) + geom_density(fill = "black") +xlab("Mean Rating by Movie")
```

```{r cdf-movie-pop}
# rank movies by popularity and compute the cdf, or fraction of movies covered by the top-k movies https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=30
# hint: use dplyr's rank and arrange functions, and the base R sum and cumsum functions
# store the result in a new data frame so you can use it in creating figure 2 from the paper below

# reference https://cmdlinetips.com/2020/08/dplyr-arrange-sort-reorder-by-variable/
library(dplyr)
for_cdf <-group_by(ratings, movie_id) %>% summarise(count = n()) %>% mutate(ranking = rank(desc(count))) %>% arrange(ranking) %>% mutate(c_sum = cumsum(count)/sum(count) )

for_cdf

for_cdf %>% ggplot(aes(x = ranking, y = c_sum)) + geom_line()
# plot the CDF of movie popularity
```


# Per-user stats

```{r aggregate-by-user}
# aggregate ratings by user, computing mean and number of ratings
group_by(ratings, user_id) %>% summarise(count = n(), avg = mean(rating))
```

```{r dist-user-activity}
# plot distribution of user activity (= number of ratings the user made)
# hint: try a log scale here

group_by(ratings, user_id) %>% summarise(num_ratings = n()) %>% ggplot(aes(x = num_ratings)) + geom_histogram() + scale_x_log10()


```

# Anatomy of the long tail

```{r long-tail}
# generate the equivalent of figure 2 of this paper:
# https://5harad.com/papers/long_tail.pdf

# Specifically, for the subset of users who rated at least 10 movies,
# produce a plot that shows the fraction of users satisfied (vertical
# axis) as a function of inventory size (horizontal axis). We will
# define "satisfied" as follows: an individual user is satisfied p% of
# the time at inventory of size k if at least p% of the movies they
# rated are contained in the top k most popular movies. As in the
# paper, produce one curve for the 100% user satisfaction level and
# another for 90%---do not, however, bother implementing the null
# model (shown in the dashed lines).



users_movies <- group_by(ratings, user_id) #%>% mutate(num_reviews = n())

View(users_movies)

 hundred_satisfied <- group_by(ratings, movie_id) %>%
  summarise(count = n()) %>%
  mutate(ranking = rank(desc(count))) %>%
  arrange(ranking) %>%
  inner_join(users_movies, by=c('movie_id'='movie_id')) %>%
  subset( select = -c(count, rating) ) %>%
  ungroup() %>%
  group_by(user_id) %>%
  summarise(max_weird = max(ranking)) %>%
  arrange(max_weird) %>%
  group_by(max_weird) %>%
  summarise(num_users = n()) %>%
  mutate(amount_pleased = cumsum(num_users)/sum(num_users))
  
ninety_satisfied <- group_by(ratings, movie_id) %>%
  summarise(count = n()) %>%
  mutate(ranking = rank(desc(count))) %>%
  arrange(ranking) %>%
  inner_join(users_movies, by=c('movie_id'='movie_id')) %>%
  subset( select = -c(count, rating) ) %>%
  ungroup() %>%
  group_by(user_id) %>%
  filter(ranking >= quantile(ranking, probs = .9)) %>%
  summarise(ninety_weird = min(ranking)) %>%
  arrange(ninety_weird) %>%
  group_by(ninety_weird) %>%
  summarise(num_users = n()) %>%
  mutate(amount_pleased = cumsum(num_users)/sum(num_users))
 

ggplot() + geom_line(data = hundred_satisfied, aes(x = max_weird, y = amount_pleased))+ geom_line(data = ninety_satisfied, aes(x = ninety_weird, y = amount_pleased))


```
