<<<<<<< HEAD
---
title: "Movielens"
date: '`r Sys.time()`'
output:
  html_document:
    #code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(scales)
library(tidyverse)
library(knitr)
library(dplyr)
# set plot theme
theme_set(theme_bw())
```

# Load and preview data

Read data from the `ratings.csv` file
```{r load-data}
ratings <- read_csv('ratings.csv',
                    col_names = c('user_id','movie_id','rating','timestamp'))
```

Loaded `r format(object.size(ratings), units="Mb")` of ratings data, containing `r format(nrow(ratings), big.mark = ",")` ratings. Here's a preview:
```{r preview-data}
head(ratings) %>% kable()
```

# Summary statistics

```{r dist-ratings}
# plot the distribution of rating values https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=26
ratings %>% ggplot() + stat_count(mapping = aes( x = rating), alpha = 1/3) +
  xlab("Rating") + ylab("Number of ratings") 
```

## Per-movie stats

```{r aggregate-by-movie}
# aggregate ratings by movie, computing mean rating and number of ratings
# hint: use the n() function for easy counting within a group
m_id <- group_by(ratings, movie_id)
mean_count_ratings_movie <- summarize(m_id, mean_ratings = mean(rating), num_ratings = n())
```

```{r dist-movie-popularity} 
# plot distribution of movie popularity (= number of ratings the movie received)
# hint: try scale_x_log10() for a logarithmic x axis
mean_count_ratings_movie %>% ggplot(mapping = aes(x = num_ratings)) + geom_histogram() + scale_x_log10() + xlab("number of ratings the movie received") + ylab("number of movies with x ratings") 
```

```{r dist-mean-ratings-by-movie}
# plot distribution of mean ratings by movie https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=28
# hint: try geom_histogram and geom_density
mean_count_ratings_movie %>% ggplot(mapping = aes(x = mean_ratings)) + geom_density(color = "gray", fill = "gray")  + xlab("Mean Rating by Movie")
```

```{r cdf-movie-pop}
# rank movies by popularity and compute the cdf, or fraction of movies covered by the top-k movies https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=30
# hint: use dplyr's rank and arrange functions, and the base R sum and cumsum functions
# store the result in a new data frame so you can use it in creating figure 2 from the paper below
movies_ranks <- mean_count_ratings_movie %>% arrange(desc(num_ratings)) %>% mutate(rank = rank(desc(num_ratings), ties.method = "random"), cdf = cumsum(num_ratings)/sum(num_ratings))

# plot the CDF of movie popularity
ggplot(movies_ranks, mapping = aes(x = rank, y = cdf)) + geom_smooth() + geom_vline(xintercept=c(3000), linetype="dotted") + xlab("Movie Rank") + ylab("CDF")
```


# Per-user stats

```{r aggregate-by-user}
# aggregate ratings by user, computing mean and number of ratings
u_id <- group_by(ratings, user_id)
mean_count_ratings_user <- summarize(u_id, mean_ratings = mean(rating), num_ratings = n())
```

```{r dist-user-activity}
# plot distribution of user activity (= number of ratings the user made)
# hint: try a log scale here
mean_count_ratings_user %>% ggplot(mapping = aes(x = num_ratings)) + geom_histogram() + scale_x_log10() + xlab("Number of ratings the user made") + ylab("Number of users that made x ratings")
```

# Anatomy of the long tail

```{r long-tail}
# generate the equivalent of figure 2 of this paper:
# https://5harad.com/papers/long_tail.pdf

# Specifically, for the subset of users who rated at least 10 movies,

user_rank <- inner_join(u_id, movies_ranks, "movie_id")
user_sat <- user_rank %>% summarize(sat_100 = max(rank), sat_90 = quantile(rank, probs = 0.9))

# number of users satisfied
num_users_sat_100 <- user_sat %>% group_by(sat_100) %>% summarize(tot_users_sat_100 = n())
num_users_sat_90 <- user_sat %>% group_by(sat_90) %>% summarize(tot_users_sat_90 = n())

final_100 <- arrange(num_users_sat_100, sat_100) %>% mutate(perc_user_sat_100 = cumsum(tot_users_sat_100)/sum(tot_users_sat_100))
final_90 <- arrange(num_users_sat_90, sat_90) %>% mutate(perc_user_sat_90 = cumsum(tot_users_sat_90)/sum(tot_users_sat_90))
  
ggplot() +
  geom_smooth(final_100, mapping = aes(x = sat_100, y = perc_user_sat_100), color = 'blue') + 
  geom_smooth(final_90, mapping = aes(x = sat_90, y = perc_user_sat_90), color = 'red') + 
  geom_vline(xintercept=c(3000), linetype="dotted") +
  xlab("Inventory Size") +
  ylab("Percent of Users Satisfied")
  
  

# produce a plot that shows the fraction of users satisfied (vertical
# axis) as a function of inventory size (horizontal axis). We will
# define "satisfied" as follows: an individual user is satisfied p% of
# the time at inventory of size k if at least p% of the movies they
# rated are contained in the top k most popular movies. As in the
# paper, produce one curve for the 100% user satisfaction level and
# another for 90%---do not, however, bother implementing the null
# model (shown in the dashed lines).



```
=======
---
title: "Movielens"
date: '`r Sys.time()`'
output:
  html_document:
    #code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(scales)
library(tidyverse)
library(knitr)

# set plot theme
theme_set(theme_bw())
```

# Load and preview data

Read data from the `ratings.csv` file
```{r load-data}
ratings <- read_csv('ratings.csv',
                    col_names = c('user_id','movie_id','rating','timestamp'))
```

Loaded `r format(object.size(ratings), units="Mb")` of ratings data, containing `r format(nrow(ratings), big.mark = ",")` ratings. Here's a preview:
```{r preview-data}
head(ratings) %>% kable()
```

# Summary statistics

```{r dist-ratings}
# plot the distribution of rating values https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=26
```

## Per-movie stats

```{r aggregate-by-movie}
# aggregate ratings by movie, computing mean rating and number of ratings
# hint: use the n() function for easy counting within a group
```

```{r dist-movie-popularity}
# plot distribution of movie popularity (= number of ratings the movie received)
# hint: try scale_x_log10() for a logarithmic x axis
```

```{r dist-mean-ratings-by-movie}
# plot distribution of mean ratings by movie https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=28
# hint: try geom_histogram and geom_density
```

```{r cdf-movie-pop}
# rank movies by popularity (number of ratings) and compute the cdf, or fraction of movies covered by the top-k movies https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=30
# hint: use dplyr's rank and arrange functions, and the base R sum and cumsum functions
# store the result in a new data frame so you can use it in creating figure 2 from the paper below

# plot the CDF of movie popularity
```


# Per-user stats

```{r aggregate-by-user}
# aggregate ratings by user, computing mean and number of ratings
```

```{r dist-user-activity}
# plot distribution of user activity (= number of ratings the user made)
# hint: try a log scale here
```

# Anatomy of the long tail

```{r long-tail}
# generate the equivalent of figure 2a of this paper:
# note: don't worry about the "null model" lines
# just do the solid lines and dotted line (optional)
# https://5harad.com/papers/long_tail.pdf

# Specifically, for the subset of users who rated at least 10 movies,
# produce a plot that shows the fraction of users satisfied (vertical
# axis) as a function of inventory size (horizontal axis). We will
# define "satisfied" as follows: an individual user is satisfied p% of
# the time at inventory of size k if at least p% of the movies they
# rated are contained in the top k most popular movies. As in the
# paper, produce one curve for the 100% user satisfaction level and
# another for 90%---do not, however, bother implementing the null
# model (shown in the dashed lines).
```
>>>>>>> 2974764705b517c7bcc4663da3b65080179b158a
