---
title: "Movielens"
date: '`r Sys.time()`'
output:
  html_document:
    #code_folding: hide
    number_sections: yes
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
library(scales)
library(tidyverse)
library(knitr)

# set plot theme
theme_set(theme_bw())
```

# Load and preview data

Read data from the `ratings.csv` file
```{r load-data}
ratings <- read_csv('ratings.csv',
                    col_names = c('user_id','movie_id','rating','timestamp'))
```

Loaded `r format(object.size(ratings), units="Mb")` of ratings data, containing `r format(nrow(ratings), big.mark = ",")` ratings. Here's a preview:
```{r preview-data}
head(ratings) %>% kable()
```

# Summary statistics

```{r dist-ratings}

library(ggplot2)
library(dplyr)

ratings %>% 
group_by(rating) %>% 
summarize(count = n())  

ratings %>%
ggplot(mapping = aes(x = rating)) +
geom_bar(fill = 'red', color = 'darkred') +
scale_y_continuous(labels = scales::label_comma()) +
labs(y = "Number of ratings")

# plot the distribution of rating values https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=26
```

## Per-movie stats

```{r aggregate-by-movie}
# aggregate ratings by movie, computing mean rating and number of ratings
# hint: use the n() function for easy counting within a grouph
df <- ratings %>% 
group_by(movie_id) %>% 
summarize(num_of_ratings = n(), mean_rating = mean(rating))

```{r dist-movie-popularity}
# plot distribution of movie popularity (= number of ratings the movie received)
# hint: try scale_x_log10() for a logarithmic x axis
df %>%
ggplot(mapping = aes(x = num_of_ratings)) + 
geom_histogram(fill = 'lightblue', color = 'darkblue') +
scale_x_log10() +
labs(title = 'Movie Popularity', 
x = 'Number of Ratings',
y = 'Number of Movies'
)
```

```{r dist-mean-ratings-by-movie}
# plot distribution of mean ratings by movie https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=28
# hint: try geom_histogram and geom_density
df %>%
ggplot(mapping = aes(x = mean_rating)) + 
geom_density(fill = 'lightgreen', color = 'darkgreen') +
labs(
  title = 'Distribution of Mean Ratings by Movie',
  x = 'Average Rating Score',
  y = 'Number of Movies'
)

```

```{r cdf-movie-pop}
# rank movies by popularity (number of ratings) and compute the cdf, or fraction of all views covered by the top-k movies https://speakerdeck.com/jhofman/modeling-social-data-lecture-2-introduction-to-counting?slide=30
# hint: use dplyr's rank and arrange functions, and the base R sum and cumsum functions
# store the result in a new data frame so you can use it in creating figure 2 from the paper below
cdf_df <- ratings %>%
group_by(movie_id) %>%
summarize(count = n()) %>%
arrange(desc(count)) %>%
mutate(rank = row_number()) %>%
mutate(cdf = cumsum(count)/sum(count))


group_by(movie_id) %>%
summarize(count = n()) %>%
arrange(desc(count)) %>%
mutate(popularity = row_number()) %>%
mutate(cdf = cumsum(count)/sum(count)) %>%
ggplot(map
# plot the CDF of movie popularity
```


# Per-user stats

```{r aggregate-by-user}
# aggregate ratings by user, computing mean and number of ratings
```

```{r dist-user-activity}
# plot distribution of user activity (= number of ratings the user made)
# hint: try a log scale here
```
# plot the CDF of movie popularity
```
cdf_df %>%
ggplot(mapping = aes(x = rank, y = cdf * 100)) +
geom_point() +
scale_y_continuous(labels = \(x) paste(x, '%')) +
labs(
  title = 'What Fractions of Ratings are Given to the Most Popular Movies?',
  x = 'Movie Rank',
  y = 'CDF'
)


# Per-user stats

```{r aggregate-by-user}
# aggregate ratings by user, computing mean and number of ratings
user_activity <- ratings %>%
group_by(user_id) %>%
summarize(count = n(), mean_ratings = mean(rating))
```

```{r dist-user-activity}
# plot distribution of user activity (= number of ratings the user made)
# hint: try a log scale here
user_activity %>%
ggplot(mapping = aes(x = count)) + 
geom_histogram(fill = 'lightblue', color = 'darkblue') +
scale_x_log10() +
labs(title = 'User Activity', 
x = 'Number of Ratings',
y = 'Number of Users'
) 
```

# Anatomy of the long tail

```{r long-tail}
# generate the equivalent of figure 2a of this paper:
# note: don't worry about the "null model" lines
# just do the solid lines and dotted line (optional)
# https://5harad.com/papers/long_tail.pdf

# Specifically, for the subset of users who rated at least 10 movies,
# produce a plot that shows the fraction of users satisfied (vertical
# axis) as a function of inventory size (horizontal axis). We will
# define "satisfied" as follows: an individual user is satisfied p% of
# the time at inventory of size k if at least p% of the movies they
# rated are contained in the top k most popular movies. As in the
# paper, produce one curve for the 100% user satisfaction level and
# another for 90%---do not, however, bother implementing the null
# model (shown in the dashed lines).
```
df_rank <- ratings %>%
group_by(movie_id) %>%
summarize(count = n()) %>%
arrange(desc(count)) %>%
mutate(rank = row_number()) %>%
select(movie_id, rank)

df_new <- ratings %>% left_join(df_rank, by = 'movie_id')

df_100 <- df_new %>%
group_by(user_id) %>%
summarize(inventory_size = max(rank)) %>%
arrange(inventory_size) %>%
group_by(inventory_size) %>%
summarize(users_satisfied = n()) %>%
mutate(cdf = cumsum(users_satisfied)/sum(users_satisfied)) 

df_90 <- df_new %>%
group_by(user_id) %>%
summarize(inventory_size = quantile(rank, probs = 0.9, type = 1)) %>%
arrange(inventory_size) %>%
group_by(inventory_size) %>%
summarize(users_satisfied = n()) %>%
mutate(cdf = cumsum(users_satisfied)/sum(users_satisfied)) 

df_plot <- rbind(
  df_100 %>% mutate(curve = '100% satisfied'),
  df_90 %>% mutate(curve = '90% satisfied')
  )

ggplot(df_plot, aes(x = inventory_size, y = cdf, color = curve)) +
geom_line() + 
labs(
  color = "Curve",
  title = "User Satisfication",
  x = "Inventory Size",
  y = "Percent of Users Satisfied"
)


