---
title: "Predicting Daily Citibike Trips"
author: "Chana Rapp"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(lubridate)
```

## Load data

```{r}
trips_per_day <-
    read_tsv("trips_per_day.tsv")
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42)

num_days <- nrow(trips_per_day)
frac_train <- 0.9
num_train <- floor(num_days * frac_train)

# randomly sample rows for the training set 
ndx <- sample(1:num_days, num_train, replace=F)

# used to fit the model
trips_per_day_train <- trips_per_day[ndx, ]

# used to evaluate the fit
trips_per_day_validate <- trips_per_day[-ndx, ]
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data
K <- 1:8
train_err_1 <- c()
validate_err_1 <- c()
model1 <- lm(num_trips ~ poly(tmin, degree = 1), data=trips_per_day_train)

for (k in K) {
  
    # fit on the training data
    model <- lm(num_trips ~ poly(tmin, k, raw = T), data=trips_per_day_train)
    
    # evaluate on the training data
    train_err_1[k] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

    # evaluate on the validate data
    validate_err_1[k] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
# Taking the effects of snow into account
model2 <- lm(num_trips ~ poly(tmin, degree = 1) + snow, data=trips_per_day_train)
validate_err_2 <- sqrt(mean((predict(model2, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

# Taking the effects of snow and precipitation into account
model3 <- lm(num_trips ~ poly(tmin, degree = 1) + snow + prcp, data=trips_per_day_train)
validate_err_3 <- sqrt(mean((predict(model3, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

# Taking the effects of snow, snow depth, and precipitation into account
model4 <- lm(num_trips ~ poly(tmin, degree = 1) + snow + snwd + prcp, data=trips_per_day_train)
validate_err_4 <- sqrt(mean((predict(model4, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

# Taking the effects of snow depth, precipitation, and whether it's a weekend into account
trips_per_day_train <- 
    trips_per_day_train %>%
    mutate(weekend = (wday(ymd, week_start = 7) == 1 | wday(ymd) == 6 | wday(ymd) == 7))

trips_per_day_validate <- 
    trips_per_day_validate %>%
    mutate(weekend = (wday(ymd, week_start = 7) == 1 | wday(ymd) == 6 | wday(ymd) == 7))

model5 <- lm(num_trips ~ poly(tmin, degree = 1) + snow + prcp + weekend, data=trips_per_day_train)
validate_err_5 <- sqrt(mean((predict(model5, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

# Taking the effects of snow depth, precipitation, temp min and max, and whether it's a weekend into account
model6 <- lm(num_trips ~ poly(tmin, degree = 1) + snow + prcp + weekend + tmax, data=trips_per_day_train)
validate_err_6 <- sqrt(mean((predict(model6, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

model7 <- lm(num_trips ~ poly(tmin, degree = 1) + (snow * prcp) + weekend + tmax, data=trips_per_day_train)
validate_err_7 <- sqrt(mean((predict(model7, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

trips_per_day_train <- 
    trips_per_day_train %>%
    mutate(month = mday(ymd))

trips_per_day_validate <- 
    trips_per_day_validate %>%
    mutate(month = mday(ymd))

model8 <- lm(num_trips ~ poly(tmin, degree = 1) + snow + prcp + weekend + tmax + month, data=trips_per_day_train)
validate_err_8 <- sqrt(mean((predict(model8, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))

```

## Model comparison

Try different models and compare their validation performance.

```{r}
K <- 1:8
models <- c("num_trips ~ poly(tmin, degree = ",
    "num_trips ~ snow + poly(tmin, degree = ",
    "num_trips ~ snow + prcp + poly(tmin, degree = ",
    "num_trips ~ snow + snwd + prcp + poly(tmin, degree = ",
    "num_trips ~ snow + prcp + weekend + poly(tmin, degree = ",
    "num_trips ~ snow + prcp + weekend + tmax + poly(tmin, degree = ",
    "num_trips ~ (snow * prcp) + weekend + tmax + poly(tmin, degree = ",
    "num_trips ~ snow + prcp + month + weekend + tmax + poly(tmin, degree = ")

train_err <- matrix(nrow = 8, ncol = 8)
validate_err <- matrix(nrow = 8, ncol = 8)

for (m in 1:8) {
    for (k in K) {
    
        # fit on the training data
        model <- lm(as.formula(paste(models[m], k, ")")), trips_per_day_train)
        
        # evaluate on the training data
        train_err[m, k] <- sqrt(mean((predict(model, trips_per_day_train) - trips_per_day_train$num_trips)^2))

        # evaluate on the validate data
        validate_err[m, k] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }
}
```

## Final model

Pick your best model and plot the results.

```{r}
final_model <- lm(num_trips ~ poly(tmin, degree = 4) + snow + prcp + weekend + tmax + month, data=trips_per_day_train)

results <- ggplot(trips_per_day_validate, aes(x = num_trips, y = predict(final_model, trips_per_day_validate))) +
    geom_point() +
    geom_abline(na.rm =TRUE) +
    xlab("Number of Trips (actual value)") +
    ylab("Predicted Number of Trips")
```

## Save model

```{r}
save(final_model, file = ".RData")
ggsave("C:\\Users\\chana\\Documents\\coursework\\week3\\predict_citibike_results.png", results)
```

## Test set evaluatest_citibike_predictions.Rmdtion

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
