---
title: "Predicting Daily Citibike Trips"
author: "Mane Diouf"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(lubridate)
```

## Load data

```{r}
trips_per_day <- read_tsv('trips_per_day.tsv') 

```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
# Reproducibility --> getting the same result no matter whoever runs the code
# or whenever you run it again
set.seed(123)

# Number of observation
obs <- nrow(trips_per_day)

# Create the 10% final test set
ndx <- sample(
    1:obs,
    size = floor(0.10 * obs) # floor - round it down instead of up
)
test_data <- trips_per_day[ndx, ]
train_valid_data <- trips_per_day[-ndx, ]

#Create validation set from the remaining 90%
train_fract <- nrow(train_valid_data)
validation_ndx <- sample(
    1:train_fract,
    size = floor(0.20 * train_fract)
)
valid_data <- train_valid_data[validation_ndx, ]
train_data <- train_valid_data[-validation_ndx, ]

```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data

# fit a model for each polynomial degree
K <- 1:6
train_err <- c()
validate_err <- c()
for (k in K) {
    # fit on the training data
    model <- lm(num_trips ~ poly(tmin, k, raw = TRUE),
    data=train_data
    )
    # evaluate on the training data
    train_err[k] <- sqrt(
        mean((predict(model, train_data) - train_data$num_trips)^2)
    )
    # evaluate on the validate data --> RMSE
    validate_err[k] <- sqrt(
        mean((predict(model, valid_data) - valid_data$num_trips)^2)
    )
}

```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
train_data = train_data %>% mutate(day = wday(ymd))
valid_data = valid_data %>% mutate(day = wday(ymd))

```

## Model comparison

Try different models and compare their validation performance.

```{r}
model1 <- lm(num_trips ~ ., data=train_data)
summary(model1)
rmse(model1, valid_data)

model2 <- lm(num_trips ~ tmin, data=train_data)
summary(model2)
rmse(model2, valid_data)

model3 <- lm(num_trips ~.+day+ poly(tmax, 3, raw = TRUE), data=train_data)
summary(model3)
rmse(model3, valid_data)

model4 <- lm(num_trips ~. -prcp, data=train_data)
summary(model4)
rmse(model4, valid_data)

model5 <- lm(num_trips ~.+day+ poly(tmin, 5, raw = TRUE), data=train_data)
summary(model5)
rmse(model5, valid_data)

```

## Final model

Pick your best model and plot the results.

```{r}
add_predictions(valid_data, model3)%>%
ggplot(aes(x = ymd, y = pred)) +
  geom_line()+
  geom_point(aes(x = ymd, y=num_trips, color= "pink")) +
  xlab('Date') +
  ylab('Number of trips')

```

## Save model

```{r}
save(model3, file = 'C:/Users/mane/Desktop/coursework/week3/model.Rdata')
```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
load("model.Rdata")
pred <- predict(model3, test_data)
test_data <- test_data %>% mutate(day = wday(ymd))
rmse(model3, test_data)

# I think this model will perform good in the future based on the 
# pretty acceptable rmse but we never know how the future can change the model.
```
