---
title: "Predicting Daily Citibike Trips"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(here)
library(lubridate)
library(modelr)
```

## Load data

```{r data-loading}
Trip_per_day <- read_tsv(here('week3','trips_per_day.tsv'))
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r Test, Train, Validation set}
set.seed(42)
frac_test<-0.1
frac_train<-0.8

Days <- nrow(Trip_per_day)
num_test <- floor(Days * frac_test)
num_train <- floor((Days - num_test) * frac_train)

ndx = sample(1:Days,num_test,replace = FALSE)
trips_per_day_test <- Trip_per_day[ndx,]
train_val <- Trip_per_day[-ndx,]

tdx <- sample(1:(Days-num_test),num_train,replace = FALSE)
trips_per_day_train <- train_val[tdx,]
trips_per_day_val <- train_val[-tdx,]
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data
num_folds = 5
train_val_fold <- train_val%>%
mutate(fold = (row_number()%%num_folds)+1)
K <- 1:8
avg_val_err <- c()
se_val_err <- c()
for (k in K) {

    val_err <-c()
    for (f in 1:num_folds){
        trips_per_day_fold_train <- filter(train_val_fold,fold != f)
        model <- lm(num_trips ~ poly(tmin,k,raw = TRUE),data = trips_per_day_fold_train)

        trips_per_day_fold_val <- filter(train_val_fold, fold == f)
        val_err[f]<- sqrt(mean((predict(model,trips_per_day_fold_val) - trips_per_day_fold_val$num_trips)^2))
    }
    avg_val_err[k] <- mean(val_err)
    se_val_err[k] <- sd(val_err)/sqrt(num_folds)
}
data.frame(K,avg_val_err,se_val_err)%>%
ggplot(aes(x=K,y=avg_val_err))+
geom_pointrange(aes(ymin = avg_val_err - se_val_err,
                    ymax = avg_val_err + se_val_err,
                    color = avg_val_err == min(avg_val_err)))+
geom_line(color = "red")+
scale_x_continuous(breaks= 1:12)+
theme(legend.position = "none")
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
train_val<-train_val%>%
mutate(day = wday(ymd,week_start = 1))%>%
mutate(day_type = ifelse(day %in% c(6,7),"weekend","weekdays" ))%>%
mutate(sigprcp = (prcp>=0.10))%>%
mutate(avgtemp =log((tmax+tmin)/2))%>%
mutate(month = month(ymd))%>%
mutate(seasonality = ifelse(month %in%c(1,2,3,11,12),"winter","summer"))

train_val%>%ggplot(aes(x=ymd,y=avgtemp,color = day))+geom_point()
train_val%>%print(n=10,Inf)

```

## Model comparison

Try different models and compare their validation performance.

```{r}

num_folds = 5
train_val_fold <- train_val%>%
mutate(fold = (row_number()%%num_folds)+1)
avg_val_err <- c()
se_val_err <- c()
val_err <-matrix(NA,7,num_folds)
rsquared <-matrix(NA,7,num_folds)
for (f in 1:num_folds){
    trips_per_day_fold_train <- filter(train_val_fold,fold != f)
    baseline_model <- lm(num_trips ~ poly(tmin,5,raw = TRUE),data = trips_per_day_fold_train)
    model_1 <-lm(num_trips ~ .-ymd -date-fold , data = trips_per_day_fold_train)
    model_2 <-lm(num_trips~.+month*avgtemp-avgtemp-day-tmin-fold-date-ymd-prcp-snow,data = trips_per_day_fold_train)
    model_3 <-lm(num_trips~.+seasonality*avgtemp-avgtemp-day-tmin-fold-date-ymd-prcp-snow,data = trips_per_day_fold_train)
    model_4 <-lm(num_trips~.+poly(month*avgtemp,3)+seasonality*avgtemp+poly(tmin,4,raw =T)-avgtemp-day-tmin-date-ymd-prcp-snow-fold,data = trips_per_day_fold_train)
    model_5 <-lm(num_trips~.+seasonality*poly(tmin,3,raw=T)+poly(avgtemp,2,raw = T)-day-ymd-fold-date-prcp-month-tmin-seasonality,data = trips_per_day_fold_train)
    model_6 <-lm(num_trips ~ .-day-date-month-avgtemp-tmin-ymd-snow-fold,data=trips_per_day_fold_train)# I will go for simplicity this appears to be more robust than others
    model_7 <-lm(num_trips~.+seasonality*poly(tmin,2,raw=T)+exp(prcp)-avgtemp-day-ymd-date-month-seasonality-fold,data = trips_per_day_fold_train)
    trips_per_day_fold_val <- filter(train_val_fold, fold == f)
    val_err[1,f]<-rmse(baseline_model, trips_per_day_fold_val)
    val_err[2,f]<-rmse(model_1, trips_per_day_fold_val)
    val_err[3,f]<-rmse(model_2, trips_per_day_fold_val)
    val_err[4,f]<-rmse(model_3, trips_per_day_fold_val)
    val_err[5,f]<-rmse(model_4, trips_per_day_fold_val)
    val_err[6,f]<-rmse(model_5, trips_per_day_fold_val)
    val_err[7,f]<-rmse(model_6, trips_per_day_fold_val)
    rsquared[1,f]<-rsquare(baseline_model, trips_per_day_fold_val)
    rsquared[2,f]<-rsquare(model_1, trips_per_day_fold_val)
    rsquared[3,f]<-rsquare(model_2, trips_per_day_fold_val)
    rsquared[4,f]<-rsquare(model_3, trips_per_day_fold_val)
    rsquared[5,f]<-rsquare(model_4, trips_per_day_fold_val)
    rsquared[6,f]<-rsquare(model_5, trips_per_day_fold_val)
    rsquared[7,f]<-rsquare(model_6, trips_per_day_fold_val)
}

rowMeans(val_err)
val_err%>%apply(1,sd)/sqrt(num_folds)
rsquared

```

## Final model

Pick your best model and plot the results.

```{r}
add_predictions(trips_per_day_fold_val,model_6)%>%
ggplot(aes(x=ymd,y=pred,color ='red'))+
geom_line()+
geom_point(aes(x=ymd,num_trips,color='blue'))
```

## Save model

```{r}
final_model <- model_6
summary(final_model)
save(final_model,file = "C:/Users/nabin/DS3/coursework/week3/model.Rdata")
```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
# Since I ma opting for simple model with basic feature extration I believe it will perform with rmse around 3500- 4000
model = load(here("week3","model.Rdata"))
summary(final_model)
# feature extration for model 
trips_per_day_test<-trips_per_day_test%>%
mutate(day = wday(ymd,week_start = 1))%>%
mutate(day_type = ifelse(day %in% c(6,7),"weekend","weekdays" ))%>%
mutate(sigprcp = (prcp>=0.10))%>%
mutate(avgtemp =log((tmax+tmin)/2))%>%
mutate(month = month(ymd))%>%
mutate(seasonality = ifelse(month %in%c(1,2,3,11,12),"winter","summer"))%>%
mutate(fold = 1)
# final rmse score
rmse = rmse(final_model,trips_per_day_test)
rmse
```
