---
title: "Predict Citibike"
output: html_notebook
---

Use the trips_per_day.tsv file that has one row for each day, the number of trips taken on that day, and the minimum temperature on that day.

```{r setup}

library(tidyverse)
library(scales)
library(modelr)
library(lubridate)

theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)

trips_per_day <- read_tsv('trips_per_day.tsv')

View(trips_per_day)

```


Split the data into randomly selected training, validation, and test sets, with 80% of the data for training the model, 10% for validation, and 10% for a final test set (to be used once and only once towards the end of this exercise). You can adapt the code from last week's complexity control notebook to do this. You can use a single validation fold, don't worry about k-fold cross-validation here.

```{r split data 80-10-10}

num_days <- nrow(trips_per_day)
frac_train <- 0.8
num_train <- floor(num_days * frac_train)

# randomly sample rows for the training set
training_data <- sample(1:num_days, num_train, replace=F)

# used to fit the model
trips_per_day_train <- trips_per_day[training_data, ]

# used to evaluate the fit and test
trips_per_day_validate_and_test <- trips_per_day[-training_data, ]

num_days_in_validate_and_test <-nrow(trips_per_day_validate_and_test)
frac_validate <- 0.5
num_validate <- floor(num_days_in_validate_and_test * frac_validate)

# randomly sample rows for the validation set
validation_data <- sample(1:num_days_in_validate_and_test, num_validate, replace=F) 

# used to validate the model
trips_per_day_validate <- trips_per_day_validate_and_test[validation_data, ]

# used to test the model
trips_per_day_test <- trips_per_day_validate_and_test[-validation_data, ]

```


Start out with the model in that notebook, which uses only the minimum temperature on each day to predict the number of trips taken that day. Try different polynomial degrees in the minimum temperature and check that you get results similar to what's in that notebook, although they likely won't be identical due to shuffling of which days end up in the train, validation, and test splits. Quantify your performance using root mean-squared error.

```{r model for each polynomial degree with min temp}

K <- 1:8

train_err <- c()
validate_err <- c()

for (k in K) {
  
  # fit on the training data
  model_tmin <- lm(num_trips ~ poly(tmin, k, raw=T), data=trips_per_day_train)
  
  # evaluate on the training data
  train_err[k] <- sqrt(mean((predict(model_tmin, trips_per_day_train) - trips_per_day_train$num_trips)^2))
  
  # evaluate on the validate data
  validate_err[k] <- sqrt(mean((predict(model_tmin, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}

train_err
validate_err

# plot the training and validation error as a function of polynomial degree

plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')

```

```{r model for a degree 5 polynomial with tmin}

# model for degree 5 polynomial
model_tmin5 <- lm(num_trips ~ poly(tmin, 5, raw=T), data=trips_per_day_train)

# train model
trips_per_day_train <- trips_per_day_train %>%
  add_predictions(model_tmin5) %>%
  mutate(split="train")

# validate model
trips_per_day_validate <- trips_per_day_validate %>%
  add_predictions(model_tmin5) %>%
  mutate(split="validate")

# plot the training data and validation data with the model
plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

ggplot(plot_data, aes(x=tmin, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_line(aes(y=pred)) +
  xlab("Minimum temperature") +
  ylab('Daily trips') +
  scale_y_continuous()

# test the model
test_err <- sqrt(mean((predict(model_tmin5, trips_per_day_test) - trips_per_day_test$num_trips)^2))
test_err

rsquare(model_tmin5, trips_per_day_train)
rsquare(model_tmin5, trips_per_day_validate)

```



RMSE with min temp:

          RMSE for training error       RMSE for validation error
degree 1:         5623.200                        6086.703
degree 2:         5622.696                        6080.974
degree 3:         5405.370                        6164.553
degree 4:         5317.392                        6216.061
degree 5:         5311.401                        6218.859
degree 6:         5307.016                        6241.981
degree 7:         5306.796                        6236.515
degree 8:         5303.677                        6248.001

R^2 for deg 5     0.7009357                       0.7208459

---------------------------------------------------

## Model with precipitation

```{r look at precipitation data}

# plot the number of trips as a function of the precipitation

# find the mean, median, and standard deviation of the preciptation
trips_per_day %>%
  summarize(prcp_mean = mean(prcp), prcp_median = median(prcp), prcp_sd = sd(prcp))

trips_per_day %>%
  ggplot(aes(x = prcp, y = num_trips)) +
  geom_point() +
  geom_vline(xintercept = mean(trips_per_day$prcp))



trips_per_day_with_rain <- trips_per_day %>% 
  mutate(too_rainy = prcp > mean(prcp))

View(trips_per_day_with_rain)

```

```{r split data with rain 80-10-10}

num_days <- nrow(trips_per_day_with_rain)
frac_train <- 0.8
num_train <- floor(num_days * frac_train)

# randomly sample rows for the training set
training_data <- sample(1:num_days, num_train, replace=F)

# used to fit the model
trips_per_day_with_rain_train <- trips_per_day_with_rain[training_data, ]

# used to evaluate the fit and test
trips_per_day_with_rain_validate_and_test <- trips_per_day_with_rain[-training_data, ]

num_days_in_validate_and_test <-nrow(trips_per_day_with_rain_validate_and_test)
frac_validate <- 0.5
num_validate <- floor(num_days_in_validate_and_test * frac_validate)

# randomly sample rows for the validation set
validation_data <- sample(1:num_days_in_validate_and_test, num_validate, replace=F) 

# used to validate the model
trips_per_day_with_rain_validate <- trips_per_day_with_rain_validate_and_test[validation_data, ]

# used to test the model
trips_per_day_with_rain_test <- trips_per_day_with_rain_validate_and_test[-validation_data, ]

```

```{r model for each polynomial degree with min temp and rain}

K <- 1:8

train_err <- c()
validate_err <- c()

for (k in K) {
  
  # fit on the training data
  model_prcp <- lm(num_trips ~ poly(tmin, k, raw=T) + too_rainy, data=trips_per_day_with_rain_train)
  
  # evaluate on the training data
  train_err[k] <- sqrt(mean((predict(model_prcp, trips_per_day_with_rain_train) - trips_per_day_with_rain_train$num_trips)^2))
  
  # evaluate on the validate data
  validate_err[k] <- sqrt(mean((predict(model_prcp, trips_per_day_with_rain_validate) - trips_per_day_with_rain_validate$num_trips)^2))
}

train_err
validate_err

# plot the training and validation error as a function of polynomial degree

plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')


```

```{r model for a degree 5 polynomial with min temp and rain}

# model for degree 5 polynomial
model_prcp5 <- lm(num_trips ~ poly(tmin, 5, raw=T) + too_rainy, data=trips_per_day_with_rain_train)

# train model
trips_per_day_with_rain_train <- trips_per_day_with_rain_train %>%
  add_predictions(model_prcp5) %>%
  mutate(split="train")

# validate model
trips_per_day_with_rain_validate <- trips_per_day_with_rain_validate %>%
  add_predictions(model_prcp5) %>%
  mutate(split="validate")

# plot the training data and validation data with the model
plot_data <- bind_rows(trips_per_day_with_rain_train, trips_per_day_with_rain_validate)

ggplot(plot_data, aes(x=ymd, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_line(aes(y=pred)) +
  xlab("YMD") +
  ylab('Daily trips') +
  scale_y_continuous()

# plot with actual vs. predicted
ggplot(plot_data, aes(x=pred, y=num_trips, color=split)) +
  geom_point() +
  geom_abline() +
  scale_x_continuous("Predicted Number of Trips", label=comma) +
  scale_y_continuous("Actual Number of Trips", label=comma)
  

# test the model
test_err <- sqrt(mean((predict(model_prcp5, trips_per_day_with_rain_test) - trips_per_day_with_rain_test$num_trips)^2))
test_err

rsquare(model_prcp5, trips_per_day_with_rain_train)
rsquare(model_prcp5, trips_per_day_with_rain_validate)

```


RMSE with min temp and rain:

          RMSE for training error       RMSE for validation error
degree 1:         4903.669                        5745.404
degree 2:         4897.954                        5782.235
degree 3:         4684.809                        5617.024
degree 4:         4634.298                        5609.840
degree 5:         4632.417                        5574.024
degree 6:         4632.408                        5574.608
degree 7:         4630.670                        5609.968
degree 8:         4623.799                        5554.882

R^2 for deg 5     0.7798285                       0.7704893

--------------------------------------------------

--------------------------------------------------
## model with weekend

```{r look at whether a day is a weekend}

# plot the number of trips as a function of whether or not a day is a weekend
trips_per_day %>%
  ggplot(aes(x = wday(ymd), y = num_trips)) +
  geom_point()

# add a column to the dataframe to indicate if a day is a weekend
trips_per_day_with_weekends <- trips_per_day %>% 
  mutate(is_weekend = wday(ymd) %in% c(1,7))

View(trips_per_day_with_weekends)

```

```{r split data with weekends 80-10-10}

num_days <- nrow(trips_per_day_with_weekends)
frac_train <- 0.8
num_train <- floor(num_days * frac_train)

# randomly sample rows for the training set
training_data <- sample(1:num_days, num_train, replace=F)

# used to fit the model
trips_per_day_with_weekends_train <- trips_per_day_with_weekends[training_data, ]

# used to evaluate the fit and test
trips_per_day_with_weekends_validate_and_test <- trips_per_day_with_weekends[-training_data, ]

num_days_in_validate_and_test <-nrow(trips_per_day_with_weekends_validate_and_test)
frac_validate <- 0.5
num_validate <- floor(num_days_in_validate_and_test * frac_validate)

# randomly sample rows for the validation set
validation_data <- sample(1:num_days_in_validate_and_test, num_validate, replace=F) 

# used to validate the model
trips_per_day_with_weekends_validate <- trips_per_day_with_weekends_validate_and_test[validation_data, ]

# used to test the model
trips_per_day_with_weekends_test <- trips_per_day_with_weekends_validate_and_test[-validation_data, ]

```

```{r model for each polynomial degree with min temp and weekends}

K <- 1:8

train_err <- c()
validate_err <- c()

for (k in K) {
  
  # fit on the training data
  model_wknd <- lm(num_trips ~ poly(tmin, k, raw=T) + is_weekend, data=trips_per_day_with_weekends_train)
  
  # evaluate on the training data
  train_err[k] <- sqrt(mean((predict(model_wknd, trips_per_day_with_weekends_train) - trips_per_day_with_weekends_train$num_trips)^2))
  
  # evaluate on the validate data
  validate_err[k] <- sqrt(mean((predict(model_wknd, trips_per_day_with_weekends_validate) - trips_per_day_with_weekends_validate$num_trips)^2))
}

train_err
validate_err

# plot the training and validation error as a function of polynomial degree

plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')


```

```{r model for a degree 5 polynomial with min temp and weekends}

# model for degree 5 polynomial
model_wknd5 <- lm(num_trips ~ poly(tmin, 5, raw=T) + is_weekend, data=trips_per_day_with_weekends_train)

# train model
trips_per_day_with_weekends_train <- trips_per_day_with_weekends_train %>%
  add_predictions(model_wknd5) %>%
  mutate(split="train")

# validate model
trips_per_day_with_weekends_validate <- trips_per_day_with_weekends_validate %>%
  add_predictions(model_wknd5) %>%
  mutate(split="validate")

# plot the training data and validation data with the model
plot_data <- bind_rows(trips_per_day_with_weekends_train, trips_per_day_with_weekends_validate)

ggplot(plot_data, aes(x=ymd, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_line(aes(y=pred)) +
  xlab("YMD") +
  ylab('Daily trips') +
  scale_y_continuous()

# plot with actual vs. predicted
ggplot(plot_data, aes(x=pred, y=num_trips, color=split)) +
  geom_point() +
  geom_abline() +
  scale_x_continuous("Predicted Number of Trips", label=comma) +
  scale_y_continuous("Actual Number of Trips", label=comma)

# test the model
test_err <- sqrt(mean((predict(model_wknd5, trips_per_day_with_weekends_test) - trips_per_day_with_weekends_test$num_trips)^2))
test_err

rsquare(model_wknd5, trips_per_day_with_weekends_train)
rsquare(model_wknd5, trips_per_day_with_weekends_validate)

```

RMSE with min temp and weekends:

          RMSE for training error       RMSE for validation error
degree 1:         5490.795                        5610.241
degree 2:         5489.336                        5596.840
degree 3:         5289.581                        5124.005
degree 4:         5188.729                        5050.044
degree 5:         5188.176                        5042.083
degree 6:         5188.124                        5042.184
degree 7:         5164.229                        5039.975
degree 8:         5136.939                        5045.129

R^2               0.7448963                       0.7783582

--------------------------------------------------


--------------------------------------------------
## Max temp

```{r view distribution of max temp}

# plot the number of trips as a function of the precipitation
trips_per_day %>%
  ggplot(aes(x = tmax, y = num_trips)) +
  geom_point()

```

```{r split data for max temp 80-10-10}

num_days <- nrow(trips_per_day)
frac_train <- 0.8
num_train <- floor(num_days * frac_train)

# randomly sample rows for the training set
training_data <- sample(1:num_days, num_train, replace=F)

# used to fit the model
trips_per_day_train <- trips_per_day[training_data, ]

# used to evaluate the fit and test
trips_per_day_validate_and_test <- trips_per_day[-training_data, ]

num_days_in_validate_and_test <-nrow(trips_per_day_validate_and_test)
frac_validate <- 0.5
num_validate <- floor(num_days_in_validate_and_test * frac_validate)

# randomly sample rows for the validation set
validation_data <- sample(1:num_days_in_validate_and_test, num_validate, replace=F) 

# used to validate the model
trips_per_day_validate <- trips_per_day_validate_and_test[validation_data, ]

# used to test the model
trips_per_day_test <- trips_per_day_validate_and_test[-validation_data, ]

```

```{r model for each polynomial degree with max temp}

K <- 1:8

train_err <- c()
validate_err <- c()

for (k in K) {
  
  # fit on the training data
  model_tmax <- lm(num_trips ~ poly(tmax, k, raw=T), data=trips_per_day_train)
  
  # evaluate on the training data
  train_err[k] <- sqrt(mean((predict(model_tmax, trips_per_day_train) - trips_per_day_train$num_trips)^2))
  
  # evaluate on the validate data
  validate_err[k] <- sqrt(mean((predict(model_tmax, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
}

train_err
validate_err

# plot the training and validation error as a function of polynomial degree

plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')


```

```{r model for a degree 5 polynomial with tmax}

# model for degree 5 polynomial
model_tmax5 <- lm(num_trips ~ poly(tmax, 5, raw=T), data=trips_per_day_train)

# train model
trips_per_day_train <- trips_per_day_train %>%
  add_predictions(model_tmax5) %>%
  mutate(split="train")

# validate model
trips_per_day_validate <- trips_per_day_validate %>%
  add_predictions(model_tmax5) %>%
  mutate(split="validate")

# plot the training data and validation data with the model
plot_data <- bind_rows(trips_per_day_train, trips_per_day_validate)

ggplot(plot_data, aes(x=tmax, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_line(aes(y=pred)) +
  xlab("Maximum temperature") +
  ylab('Daily trips') +
  scale_y_continuous()

# test the model
test_err <- sqrt(mean((predict(model_tmax5, trips_per_day_test) - trips_per_day_test$num_trips)^2))
test_err

rsquare(model_tmax5, trips_per_day_train)
rsquare(model_tmax5, trips_per_day_validate)

```

RMSE with max temp:

          RMSE for training error       RMSE for validation error
degree 1:         5496.660                        5339.519
degree 2:         5488.154                        5457.691
degree 3:         5276.818                        5229.254
degree 4:         5153.824                        5110.619
degree 5:         5152.232                        5123.421
degree 6:         5109.768                        5145.244
degree 7:         5105.340                        5105.967
degree 8:         5014.325                        5111.434

R^2               0.7497496                       0.7202377

--------------------------------------------------


--------------------------------------------------
# min temp and precipitation and weekends and max temp

```{r create dataframe with all necessary info}

trips_per_day_with_all_info <- trips_per_day %>% 
  mutate(too_rainy = prcp > mean(prcp)) %>% 
  mutate(is_weekend = wday(ymd) %in% c(1,7))

View(trips_per_day_with_all_info)

```

```{r split data with all info 80-10-10}

num_days <- nrow(trips_per_day_with_all_info)
frac_train <- 0.8
num_train <- floor(num_days * frac_train)

# randomly sample rows for the training set
training_data <- sample(1:num_days, num_train, replace=F)

# used to fit the model
trips_per_day_with_all_info_train <- trips_per_day_with_all_info[training_data, ]

# used to evaluate the fit and test
trips_per_day_with_all_info_validate_and_test <- trips_per_day_with_all_info[-training_data, ]

num_days_in_validate_and_test <-nrow(trips_per_day_with_all_info_validate_and_test)
frac_validate <- 0.5
num_validate <- floor(num_days_in_validate_and_test * frac_validate)

# randomly sample rows for the validation set
validation_data <- sample(1:num_days_in_validate_and_test, num_validate, replace=F) 

# used to validate the model
trips_per_day_with_all_info_validate <- trips_per_day_with_all_info_validate_and_test[validation_data, ]

# used to test the model
trips_per_day_with_all_info_test <- trips_per_day_with_all_info_validate_and_test[-validation_data, ]

```

```{r model for each polynomial degree with all info}

K <- 1:8

train_err <- c()
validate_err <- c()

for (k in K) {
  
# fit on the training data
model_all <- lm(num_trips ~ poly(tmin, k, raw=T) + too_rainy + is_weekend + poly(tmax, k, raw=T), data=trips_per_day_with_all_info_train)
  
# evaluate on the training data
train_err[k] <- sqrt(mean((predict(model_all, trips_per_day_with_all_info_train) - trips_per_day_with_all_info_train$num_trips)^2))
  
  # evaluate on the validate data
  validate_err[k] <- sqrt(mean((predict(model_all, trips_per_day_with_all_info_validate) - trips_per_day_with_all_info_validate$num_trips)^2))
}

train_err
validate_err

# plot the training and validation error as a function of polynomial degree

plot_data <- data.frame(K, train_err, validate_err) %>%
  gather("split", "error", -K)

ggplot(plot_data, aes(x=K, y=error, color=split)) +
  geom_line() +
  scale_x_continuous(breaks=K) +
  xlab('Polynomial Degree') +
  ylab('RMSE')

```
RMSE with all info:

          RMSE for training error       RMSE for validation error
degree 1:         4142.098                        3106.123
degree 2:         4070.378                        2988.029
degree 3:         3827.690                        2784.540
degree 4:         3804.649                        2688.986
degree 5:         3804.151                        2694.059
degree 6:         3772.687                        2586.137
degree 7:         3769.344                        2571.798
degree 8:         3741.975                        2615.304

```{r model for a degree 5 polynomial}

# model for degree 3 polynomial
model_all5 <- lm(num_trips ~ poly(tmin, 5, raw=T) + too_rainy + is_weekend + tmax, data=trips_per_day_with_all_info_train)

# train model
trips_per_day_with_all_info_train <- trips_per_day_with_all_info_train %>%
  add_predictions(model_all5) %>%
  mutate(split="train")

# validate model
trips_per_day_with_all_info_validate <- trips_per_day_with_all_info_validate %>%
  add_predictions(model_all5) %>%
  mutate(split="validate")

# plot the training data and validation data with the model
plot_data <- bind_rows(trips_per_day_with_all_info_train, trips_per_day_with_all_info_validate)

ggplot(plot_data, aes(x=ymd, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_line(aes(y=pred)) +
  xlab("YMD") +
  ylab('Daily trips') +
  scale_y_continuous()

# find R^2 values

rsquare(model_all5, trips_per_day_with_all_info_train)
rsquare(model_all5, trips_per_day_with_all_info_validate)


```


```{r model for a degree 4 polynomial}

# model for degree 3 polynomial
model_all4 <- lm(num_trips ~ poly(tmin, 4, raw=T) + too_rainy + is_weekend + tmax, data=trips_per_day_with_all_info_train)

# train model
trips_per_day_with_all_info_train <- trips_per_day_with_all_info_train %>%
  add_predictions(model_all4) %>%
  mutate(split="train")

# validate model
trips_per_day_with_all_info_validate <- trips_per_day_with_all_info_validate %>%
  add_predictions(model_all4) %>%
  mutate(split="validate")

# plot the training data and validation data with the model
plot_data <- bind_rows(trips_per_day_with_all_info_train, trips_per_day_with_all_info_validate)

ggplot(plot_data, aes(x=ymd, y=num_trips)) +
  geom_point(aes(color=split)) +
  geom_line(aes(y=pred)) +
  xlab("YMD") +
  ylab('Daily trips') +
  scale_y_continuous()

# find R^2 values

rsquare(model_all4, trips_per_day_with_all_info_train)
rsquare(model_all4, trips_per_day_with_all_info_validate)


```








--------------------------------------------------
















































