---
title: "Predicting Daily Citibike Trips"
author: "Jacqueline Louie (Stony Brook University)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
```

## Load data

```{r}
trips_per_day <- read_tsv('trips_per_day.tsv')
head(trips_per_day)
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42)
num_days <- nrow(trips_per_day)
num_folds <- 5

# Holding out 10% of data for final test set
frac_test <- 0.1
num_test <- floor(num_days * frac_test)

# Randomly sampling rows for the test set
ndx <- sample(1:num_days, num_test, replace=F)

# DO NOT TOUCH THIS
trips_per_day_test <- trips_per_day[ndx, ]

# using k-fold validation
trips_per_day_kfold <- trips_per_day[-ndx, ] %>%
    mutate(fold = (row_number() %% num_folds) + 1)

head(trips_per_day_kfold) # confirming that it looks right
range <- max(trips_per_day_kfold$num_trips) - min(trips_per_day_kfold$num_trips)

## Test set size: 36
## Train/Validate set size: 329
mean(trips_per_day_kfold$num_trips)
```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data
K <- 1:8
avg_validate_err <- c()
se_validate_err <- c()
test_avg_val <- c()
for (k in K) {
    # doing 5 fold cross-validation within each value of k
    validate_err <- c()
    for (f in 1:num_folds) {
        # fitting on training data
        trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
        model <- lm(num_trips ~ poly(tmin, k, raw = T), data = trips_per_day_train)

        # evaluate on the validation data
        trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    }

    # average validation error/standard error 
    avg_validate_err[k] <- mean(validate_err)
    se_validate_err[k] <- sd(validate_err) / sqrt(num_folds)
    test_avg_val[k] <- avg_validate_err[k] / mean(trips_per_day_kfold$num_trips)
}
avg_validate_err
se_validate_err
test_avg_val
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
# Possible model features to include
# Current features: ymd, num_trips, date, precipitation, snow depth, snow, max temp, min temp
# Possibly interesting features that contribute to trips
# tmin (colder temps)
# tmax (hotter temps)
# day of the week (weekends vs weekdays) and month of the year
trips_per_day_kfold <- mutate(trips_per_day_kfold, day_of_week = wday(ymd), month = month(ymd))
# snow depth (substantial snow depth i decided over 5) and precipitation (substantial precipitation decided for over 0.5)
trips_per_day_kfold <- mutate(trips_per_day_kfold, snow_sub = (snwd > 5), prcp_sub = (prcp > 0.5))

## Which models will i try (if the previous works)
# Model One: min temp and max temp (tried it and it made a little difference)
# Model Two: min/max temp and day of the week (tried it and it was better!)
# Model Three: min/max temp, day of week, and month (tried it and it was better)
# Model Four: min/max temp, day of week, month, snowd significance (got worse)
# Model Fourb: Four but not significance the snowd value (got slightly better but not that significant)
# Model Five: min/max temp, day of week, month, precipitation significance (got much better! RMSE dropped around .03)
# Model Fiveb: Five but the prcp value (doing precipitation as its own value is better)
# Model Fivec(best model before trying interactions): Five but the prcp value transformed with log (doing precipitation as 
# its value but logged did better than non logged)
# Model Six: min/max temp, dayofweek, month, precipitation, prcp interacted with time of year/day of week (was not any better)

# Chose Model FiveC
```

## Model comparison

Try different models and compare their validation performance.

```{r}
m1_validate_err <- c()
m1_test_val <- c()
m1_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m1_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m1_test_val[f] <- m1_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m1_rel_val[f] <- m1_validate_err[f] / range
}
m1_avg_validate <- mean(m1_validate_err)
m1_test_avg_val <- mean(m1_test_val)
m1_rel_avg_val <- mean(m1_rel_val)

m1_validate_err
m1_test_val
m1_rel_val

m1_avg_validate
m1_test_avg_val
m1_rel_avg_val
```

```{r}
m2_validate_err <- c()
m2_test_val <- c()
m2_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ factor(day_of_week) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m2_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m2_test_val[f] <- m2_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m2_rel_val[f] <- m2_validate_err[f] / range
}
m2_avg_validate <- mean(m2_validate_err)
m2_test_avg_val <- mean(m2_test_val)
m2_rel_avg_val <- mean(m2_rel_val)

m2_validate_err
m2_test_val
m2_rel_val

m2_avg_validate
m2_test_avg_val
m2_rel_avg_val
```

```{r}
m3_validate_err <- c()
m3_test_val <- c()
m3_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m3_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m3_test_val[f] <- m3_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m3_rel_val[f] <- m3_validate_err[f] / range
}
m3_avg_validate <- mean(m3_validate_err)
m3_test_avg_val <- mean(m3_test_val)
m3_rel_avg_val <- mean(m3_rel_val)

m3_validate_err
m3_test_val
m3_rel_val

m3_avg_validate
m3_test_avg_val
m3_rel_avg_val
```

```{r}
m4_validate_err <- c()
m4_test_val <- c()
m4_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ snow_sub + factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m4_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m4_test_val[f] <- m4_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m4_rel_val[f] <- m4_validate_err[f] / range
}
m4_avg_validate <- mean(m4_validate_err)
m4_test_avg_val <- mean(m4_test_val)
m4_rel_avg_val <- mean(m4_rel_val)

m4_validate_err
m4_test_val
m4_rel_val

m4_avg_validate
m4_test_avg_val
m4_rel_avg_val
# model actually got worse
```

```{r}
# model with snow got worse so im trying snow depth value itself
m4b_validate_err <- c()
m4b_test_val <- c()
m4b_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ snwd + factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m4b_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m4b_test_val[f] <- m4b_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m4b_rel_val[f] <- m4b_validate_err[f] / range
}
m4b_avg_validate <- mean(m4b_validate_err)
m4b_test_avg_val <- mean(m4b_test_val)
m4b_rel_avg_val <- mean(m4b_rel_val)

m4b_validate_err
m4b_test_val
m4b_rel_val

m4b_avg_validate
m4b_test_avg_val
m4b_rel_avg_val
# adding snow depth doesn't contribute that much
```

```{r}
m5_validate_err <- c()
m5_test_val <- c()
m5_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ prcp_sub + factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m5_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m5_test_val[f] <- m5_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m5_rel_val[f] <- m5_validate_err[f] / range
}
m5_avg_validate <- mean(m5_validate_err)
m5_test_avg_val <- mean(m5_test_val)
m5_rel_avg_val <- mean(m5_rel_val)

m5_validate_err
m5_test_val
m5_rel_val

m5_avg_validate
m5_test_avg_val
m5_rel_avg_val
# model got better by alot!
```

```{r}
m5b_validate_err <- c()
m5b_test_val <- c()
m5b_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ prcp + factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m5b_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m5b_test_val[f] <- m5b_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m5b_rel_val[f] <- m5b_validate_err[f] / range
}
m5b_avg_validate <- mean(m5b_validate_err)
m5b_test_avg_val <- mean(m5b_test_val)
m5b_rel_avg_val <- mean(m5b_rel_val)

m5b_validate_err
m5b_test_val
m5b_rel_val 

m5b_avg_validate
m5b_test_avg_val
m5b_rel_avg_val
```

```{r}
m5c_validate_err <- c()
m5c_test_val <- c()
m5c_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ log(prcp + 1) + factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m5c_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m5c_test_val[f] <- m5c_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m5c_rel_val[f] <- m5c_validate_err[f] / range
}
m5c_avg_validate <- mean(m5c_validate_err)
m5c_test_avg_val <- mean(m5c_test_val)
m5c_rel_avg_val <- mean(m5c_rel_val)

m5c_validate_err
m5c_test_val
m5c_rel_val

m5c_avg_validate
m5c_test_avg_val
m5c_rel_avg_val
```

```{r}
m6_validate_err <- c()
m6_test_val <- c()
m6_rel_val <- c()
for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    model <- lm(num_trips ~ factor(day_of_week):log(prcp + 1) + log(prcp + 1) + factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)
    m6_validate_err[f] <- sqrt(mean((predict(model, trips_per_day_validate) - trips_per_day_validate$num_trips)^2))
    m6_test_val[f] <- m6_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    m6_rel_val[f] <- m6_validate_err[f] / range
}
m6_avg_validate <- mean(m6_validate_err)
m6_test_avg_val <- mean(m6_test_val)
m6_rel_avg_val <- mean(m6_rel_val)

m6_validate_err
m6_test_val
m6_avg_validate

m6_test_avg_val
m6_rel_val
m6_rel_avg_val
```

## Final model

Pick your best model and plot the results.

```{r}
# Final Model
final_validate_err <- c()
final_test_val <- c()
final_rel_val <- c()

# storing prediction values
final_preds <- mutate(trips_per_day_kfold, predicted_value = 0)

for (f in 1:num_folds) {
    # doing 5 fold cross-validation within each value of k
    trips_per_day_train <- filter(trips_per_day_kfold, fold != f) 
    final_model <- lm(num_trips ~ log(prcp + 1) + factor(day_of_week) + factor(month) + poly(tmin, 4, raw = T) + poly(tmax, 4, raw = T), data = trips_per_day_train)

    # evaluate on the validation data
    trips_per_day_validate <- filter(trips_per_day_kfold, fold == f)

    # store the predicted values
    final_preds$predicted_value[final_preds$fold == f] <- predict(final_model, trips_per_day_validate)
    
    final_validate_err[f] <- rmse(final_model, trips_per_day_validate)
    final_test_val[f] <- final_validate_err[f] / mean(trips_per_day_kfold$num_trips)
    final_rel_val[f] <- final_validate_err[f] / range
}
final_avg_validate <- mean(final_validate_err)
final_test_avg_val <- mean(final_test_val)
final_rel_avg_val <- mean(final_rel_val)

final_validate_err
final_test_val
final_avg_validate

final_test_avg_val
final_rel_val
final_rel_avg_val

trips_per_day_kfold
final_preds
select(final_preds, ymd, num_trips, predicted_value)

## Plotting results of the model
ggplot(final_preds, aes(x = ymd, y = predicted_value)) +
    geom_line(aes(y = predicted_value)) +
    geom_point(aes(y = num_trips))
    labs(x = "Date", y = "Predicted")

ggplot(final_preds, aes(x = predicted_value, y = num_trips)) +
    geom_point() +
    geom_abline() +
    labs(x = "Predicted", y = "Actual")

```

## Save model

```{r}
save(final_model, file = 'final_model.RData')
```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
trips_per_day_test <- mutate(trips_per_day_kfold, day_of_week = wday(ymd), month = month(ymd))

load('final_model.RData')
rmse(final_model, trips_per_day_test)
rsquare(final_model, trips_per_day_test)
# My RMSE for the test set was 3208.211
# I think the model will perform okay? I'm a little concerned about if I overfit the training/validation data
```
