---
title: "Predicting Daily Citibike Trips"
author: "Addina Rahaman"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(here)
```

## Load data

```{r}
trips <- read_tsv(here('week3', 'trips_per_day.tsv'), col_names=TRUE)

# check which cols are NA
for (col in names(trips)) {
    cat(col, '\n')
    print(sum(is.na(trips[[col]])))
}
# No cols are NA

```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42)

# test split
num_days <- nrow(trips)
frac_test <- 0.1
num_test <- floor(num_days * frac_test)

test <- trips[test_indices,] # 36 rows
remaining <- trips[-test_indices,] # 329 rows

```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data

rmse <- function(y, y_hat) {
    sqrt(mean((y_hat - y)^2))
}

num_folds <- 5
train_val <- remaining %>% 
    mutate(fold = (row_number() %% num_folds) + 1)

# cross fold validation loop
val_err = c()
for (f in 1:num_folds) {
    # split train and validation
    train <- filter(train_val, fold != f)
    val <- filter(train_val, fold == f)

    # create linear model
    linear_model = lm(num_trips ~ poly(tmin, 1, raw=T), data=train)

    # get validation error
    val_err[f] = rmse(val$num_trips, predict(linear_model, val))
}
# print avg RMSE
avg_rmse_tmin_deg1 <- mean(val_err)
cat('Average RMSE when using poly(tmin, 1): ', avg_rmse_tmin_deg1) # got 5847.476
se_rmse_tmin_deg1 <- sd(val_err) / sqrt(num_folds)
cat('Standard error when using poly(tmin, 1): ', se_rmse_tmin_deg1) # got 232.2059

# compare this with trips$num_trips
summary(trips$num_trips)
#   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
#     876   13119   23908   22138   31204   38619 
# The average RMSE makes sense when looking at the summary statistics of num_trips. It could be improved, however. 

# -------------------------------------
# TODO: try different polynomial degrees and evaluate RMSE on validation data

# Try degrees 2-10
K <- 2:10
avg_val_error <- c(avg_rmse_tmin_deg1) # Initialize with deg 1
se_val_error <- c(se_rmse_tmin_deg1)

for (k in K) {

    val_err = c()
    for (f in 1:num_folds) {
        # split train and validation
        train <- filter(train_val, fold != f)
        val <- filter(train_val, fold == f)

        # create linear model
        model = lm(num_trips ~ poly(tmin, k, raw=T), data=train)

        # get validation error
        val_err[f] = rmse(val$num_trips, predict(model, val))
    }

    avg_val_error[k] <- mean(val_err)
    se_val_error[k] <- sd(val_err) / sqrt(num_folds)

}

degrees <- 1:10
# turn this into a dataframe to plot
rmse_df <- data.frame(degrees, avg_val_error, se_val_error)

# use pointrange
rmse_df %>% 
    ggplot(aes(x=degrees, y=avg_val_error)) +
    geom_pointrange(
        aes(
            ymin = avg_val_error - se_val_error,
            ymax = avg_val_error + se_val_error,
            color = avg_val_error == min(avg_val_error)
        )
    ) + 
    geom_line(color='lightgreen') + 
    scale_x_continuous(breaks=1:10) +
    labs(
        x='Degree',
        y='Validation RMSE',
    ) +
    theme(legend.position = 'none') # remove legend

# The model with the lowest error is the one with degree 6
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}

# using prcp, snwd, snow, tmax, and tmin (all weather data)

# also using weekday(1-7), if it is a holiday, and average daily wind speed 

# (1) clean holiday data for only 2014 data and only pull out the dates 
holidays <- read_csv(here('week3', 'holidays.csv'), col_names=c('index', 'date', 'holiday'))
holidays <- holidays %>% 
    filter(grepl('^2014-', date)) %>% 
    pull(date)

# (2) clean wind data by averaging the data between stations
winds_raw <- read_csv(here('week3', 'winds.csv'))
winds <- winds_raw %>% 
    group_by(DATE) %>% 
    summarize(
        average_winds=mean(AWND)
    ) %>% 
    rename(date=DATE)


# (3) for training and validation data, add holiday data and wind data
train_val_aug <- train_val %>% 
    mutate(date=ymd(date)) %>% 
    mutate(day_of_week=weekdays(date)) %>% 
    mutate(is_holiday=(date %in% holidays)) %>%
    left_join(winds, by='date')

# this tells us which date average_winds is NA (a problem i found later)
train_val_aug %>% 
    filter(is.na(average_winds)) %>% 
    select(date)

# fill NA with average of nearby days
library(zoo)
train_val_aug$average_winds <- na.approx(train_val_aug$average_winds, na.rm = FALSE)

# (4) Do same for testing data
test <- test %>% 
    mutate(date=ymd(date)) %>% 
    mutate(day_of_week=weekdays(date)) %>% 
    mutate(is_holiday=(date %in% holidays)) %>%
    left_join(winds, by='date')

# plot every predictor against num_trips to see how they are correlated

train_val_aug %>% 
    pivot_longer(
        cols=c(prcp, snwd, snow, tmax, tmin, average_winds),
        names_to='predictor',
        values_to='stat'
    ) %>% 
    ggplot(aes(x=stat, y=num_trips)) +
    geom_point() +
    geom_smooth() +
    scale_y_log10(label=comma) +
    labs(
        x = "Stat",
        y = "Number of Trips"
    ) +
    facet_wrap(~ predictor, scales="free_x")

# average_winds appears to be negative (maybe quadratic)
# for prcp, snow, and snwd, it seems more people are likely to ride a bike if it is 0
# tmax and tmin are quadratic and positive - but highly correlated so maybe only add one of them

train_val_aug %>% 
    ggplot(aes(x=is_holiday)) +
    geom_bar() +
    scale_y_log10(label=comma) +
    labs(
        x = "Is Holiday",
        y = "Number of Trips"
    ) 

train_val_aug %>% 
    ggplot(aes(x=day_of_week)) +
    geom_bar() +
    scale_y_log10(label=comma) +
    labs(
        x = "Day of Week",
        y = "Number of Trips"
    ) 

# Holiday and Day of Week are not dependent on each other

# create a second train dataset
train_val_aug2 <- train_val_aug %>% 
    mutate(
        did_snow=(snow != 0),
        sig_snwd=(snwd != 0),
        did_rain=(prcp != 0)
    )

test2 <- test %>% 
    mutate(
        did_snow=(snow != 0),
        sig_snwd=(snwd != 0),
        did_rain=(prcp != 0)
    )

```

## Model comparison

Try different models and compare their validation performance.

```{r}
# Here I test different types of models
avg_r2 <- c()
se_r2 <- c()
avg_rmse <- c()
se_rmse <- c()

mse <- function(y, y_hat) {
    (rmse(y, y_hat))^2
}

r.squared <- function(y, y_hat) {
    y.bar = mean(y)
    1 - (mse(y, y_hat)/mse(y, y.bar))
}

# Model (1): adding all features -> this assumes that each feature is independent and that the effect of one feature on ridership is not affected by other features

r2_1 <- c()
rmse_1 <- c()
for (f in 1:num_folds) {
    train <- filter(train_val_aug, fold != f)
    val <- filter(train_val_aug, fold == f)

    model = lm(num_trips ~ prcp + snwd + snow + tmax + tmin + day_of_week + is_holiday + average_winds, data=train)

    r2_1[f] <- r.squared(val$num_trips, predict(model, val))
    rmse_1[f] <- rmse(val$num_trips, predict(model, val))
}

avg_r2[1] <- mean(r2_1)
se_r2[1] <- sd(r2_1) / sqrt(num_folds)
avg_rmse[1] <- mean(rmse_1)
se_rmse[1] <- sd(rmse_1) / sqrt(num_folds)

# Model (2): only have particular related features interact (i.e. snwd & snow, tmin & tmax etc)

r2_2 <- c()
rmse_2 <- c()
for (f in 1:num_folds) {
    train <- filter(train_val_aug, fold != f)
    val <- filter(train_val_aug, fold == f)

    model = lm(num_trips ~ prcp + (snwd * snow) + (tmax * tmin) + average_winds + day_of_week + is_holiday, data=train)

    r2_2[f] <- r.squared(val$num_trips, predict(model, val))
    rmse_2[f] <- rmse(val$num_trips, predict(model, val))
}

avg_r2[2] <- mean(r2_2)
se_r2[2] <- sd(r2_2) / sqrt(num_folds)
avg_rmse[2] <- mean(rmse_2)
se_rmse[2] <- sd(rmse_2) / sqrt(num_folds)

# Model (3): 

r2_3 <- c()
rmse_3 <- c()
for (f in 1:num_folds) {
    train <- filter(train_val_aug, fold != f)
    val <- filter(train_val_aug, fold == f)

    model = lm(num_trips ~ prcp + average_winds + (snwd * snow) + (tmax * tmin) + (day_of_week * is_holiday), data=train)

    r2_3[f] <- r.squared(val$num_trips, predict(model, val))
    rmse_3[f] <- rmse(val$num_trips, predict(model, val))
}

avg_r2[3] <- mean(r2_3)
se_r2[3] <- sd(r2_3) / sqrt(num_folds)
avg_rmse[3] <- mean(rmse_3)
se_rmse[3] <- sd(rmse_3) / sqrt(num_folds)

# Model (4): applying nonlinear relationships as observed before

r2_4 <- c()
rmse_4 <- c()
for (f in 1:num_folds) {
    train <- filter(train_val_aug, fold != f)
    val <- filter(train_val_aug, fold == f)

    model = lm(num_trips ~ poly(tmax, 2) + prcp + snwd + snow + day_of_week + is_holiday + average_winds, data=train)

    r2_4[f] <- r.squared(val$num_trips, predict(model, val))
    rmse_4[f] <- rmse(val$num_trips, predict(model, val))
}

avg_r2[4] <- mean(r2_4)
se_r2[4] <- sd(r2_4) / sqrt(num_folds)
avg_rmse[4] <- mean(rmse_4)
se_rmse[4] <- sd(rmse_4) / sqrt(num_folds)

# Model 5: applying the nonlinear relationships and the new binary data

r2_5 <- c()
rmse_5 <- c()
for (f in 1:num_folds) {
    train <- filter(train_val_aug2, fold != f)
    val <- filter(train_val_aug2, fold == f)

    model = lm(num_trips ~ poly(tmax, 2) + did_rain + did_snow + sig_snwd + day_of_week + is_holiday + average_winds, data=train)
        
    r2_5[f] <- r.squared(val$num_trips, predict(model, val))
    rmse_5[f] <- rmse(val$num_trips, predict(model, val))
}

avg_r2[5] <- mean(r2_5)
se_r2[5] <- sd(r2_5) / sqrt(num_folds)
avg_rmse[5] <- mean(rmse_5)
se_rmse[5] <- sd(rmse_5) / sqrt(num_folds)

# Model 6: include both binary data and continuous data

r2_6 <- c()
rmse_6 <- c()
for (f in 1:num_folds) {
    train <- filter(train_val_aug2, fold != f)
    val <- filter(train_val_aug2, fold == f)

    model = lm(num_trips ~ poly(tmax, 2) + did_rain + did_snow + sig_snwd + day_of_week + is_holiday + average_winds + prcp + snwd + snow, data=train)
        
    r2_6[f] <- r.squared(val$num_trips, predict(model, val))
    rmse_6[f] <- rmse(val$num_trips, predict(model, val))
}

avg_r2[6] <- mean(r2_6)
se_r2[6] <- sd(r2_6) / sqrt(num_folds)
avg_rmse[6] <- mean(rmse_6)
se_rmse[6] <- sd(rmse_6) / sqrt(num_folds)

# Plot the R2
versions <- 1:6 
avg_r2_df <- data.frame(versions, avg_r2, se_r2)
avg_r2_df %>% 
    ggplot(aes(x=versions, y=avg_r2)) +
    geom_pointrange(
        aes(
            ymin = avg_r2 - se_r2,
            ymax = avg_r2 + se_r2,
            color = avg_r2 == max(avg_r2)
        )
    ) + 
    geom_line(color='lightgreen') + 
    scale_x_continuous(breaks=1:6) +
    labs(
        x='Model Version',
        y='Validation R-squared',
    ) +
    theme(legend.position = 'none')

# Model 6 is the best with 0.88 R-squared

# Plot the RMSE
avg_rmse_df <- data.frame(versions, avg_rmse, se_rmse)
avg_rmse_df %>% 
    ggplot(aes(x=versions, y=avg_rmse)) +
    geom_pointrange(
        aes(
            ymin = avg_rmse - se_rmse,
            ymax = avg_rmse + se_rmse,
            color = avg_rmse == min(avg_rmse)
        )
    ) + 
    geom_line(color='lightgreen') + 
    scale_x_continuous(breaks=1:16) +
    labs(
        x='Model Version',
        y='Validation RMSE',
    ) +
    theme(legend.position = 'none')
# Model 6 also has a significantly lower RMSE

```

## Final model

Pick your best model and plot the results.

```{r}
final_model =  lm(num_trips ~ poly(tmax, 2) + did_rain + did_snow + sig_snwd + day_of_week + is_holiday + average_winds + prcp + snwd + snow, data=train_val_aug2)

plot_data <- train_val_aug2 %>% 
    add_predictions(final_model)

# Plot 1
plot_data %>% 
    ggplot(aes(x=date)) +
    geom_point(aes(y=num_trips)) +
    geom_smooth(aes(y=pred)) + 
    labs(
        x='Date',
        y='Number of Trips'
    )

# Plot 2
plot_data %>% 
    ggplot(aes(x=pred, y=num_trips)) +
    geom_point() + 
    geom_abline(linetype='dashed') + 
    labs(
        x='Predicted',
        y='Actual',
    )

```

## Save model

```{r}

save(final_model, file='citibike_model.RData')

```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
