---
title: "Predicting Daily Citibike Trips"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(jsonlite)
library(lubridate)
library(timeDate)
```

## Load data

```{r}
# Load Citibike Data
trips_per_day <- read_tsv('trips_per_day.tsv')

# Load MTA Turnstile Data; Feb - Dec 2014
mta <- fromJSON("mta_turnstile_2014.json")

write.csv(mta, "mta_turnstile_2014.csv", row.names = FALSE)
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
cor_matrix <- cor(trips_per_day[sapply(trips_per_day, is.numeric)])

# input parameters: rows, columns
# check which features are highly correlated with the label
cor_matrix[, "num_trips"]

# Trips are mostly driven by temperature, reduced by snow/precipitation, with mild seasonal trends.

us_holidays <- as.Date(holidayNYSE(year = 2014:2026))

trips_per_day <- trips_per_day %>%
  mutate(

    # Temperature Feature
    tavg = (tmin + tmax) / 2,
    
    # Rain Features
    heavy_rain = prcp > 0.9,
    
    # Snow Features
    deep_snow = snwd > 4,
    snow_event = snow > 0,
    
    # Calendar Features
    day_of_week = wday(ymd, label = TRUE),

    is_weekend = wday(ymd) %in% c(1, 7),
    is_monday = wday(ymd) == 2,
    is_friday = wday(ymd) == 6,

    season = case_when(
      month(ymd) %in% c(12, 1, 2) ~ "winter",
      month(ymd) %in% c(3, 4, 5)  ~ "spring",
      month(ymd) %in% c(6, 7, 8)  ~ "summer",
      month(ymd) %in% c(9, 10, 11) ~ "fall"
    ),
    
    is_holiday = ymd %in% us_holidays
  )

# Hot-Encode Calendar Features
trips_per_day$day_of_week <- factor(trips_per_day$day_of_week)
trips_per_day$season <- factor(trips_per_day$season)

# Create a new column that has both the date and time saved
mta <- mta %>%
  mutate(
    datetime = ymd_hms(paste(substr(date, 1, 10), time)) # substr(x, start, stop)
  ) %>%
  arrange(c_a, unit, scp, station, datetime)

# Convert str to numeric
mta <- mta %>%
  mutate(
    entries = as.numeric(entries),
    exits = as.numeric(exits)
  )

mta <- mta %>%
  mutate(
    entry_diff = entries - lag(entries), # how many people entered during a time window
    exit_diff = exits - lag(exits) # how many people exited during a time window
  )

# Turnstile counters sometimes reset
# This creates negative jumps
# Let's replace them with 0 to avoid corrupting trip counts
mta <- mta %>%
  mutate(
    entry_diff = pmax(entry_diff, 0, na.rm = TRUE),
    exit_diff = pmax(exit_diff, 0, na.rm = TRUE)
  )

# Get the total MTA ridership by date
mta_daily <- mta %>%
  mutate(date_only = as.Date(datetime)) %>%
  group_by(date_only) %>%
  summarize(mta_ridership = sum(entry_diff + exit_diff, na.rm = TRUE))

# Replace extreme outlier days with the median daily ridership
mta_daily <- mta_daily %>%
  mutate(
    mta_ridership = if_else(
      mta_ridership > 2e7, # 20 million
      median(mta_ridership, na.rm = TRUE),
      mta_ridership
    )
  )

trips_per_day <- trips_per_day %>%
  left_join(mta_daily, by = c('ymd' = 'date_only'))


# Replace missing MTA ridership in January with the median values
trips_per_day <- trips_per_day %>%
  mutate(
    mta_ridership = ifelse(
      is.na(mta_ridership),
      median(mta_ridership, na.rm = TRUE),
      mta_ridership
    )
  )
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.
```{r}
# Set seed for reproducibility
set.seed(42)

# Create unique IDs for each trip
trips_per_day <- trips_per_day %>%
mutate("id" = row_number())

# Number of Rows to Sample From
n_rows_total <- nrow(trips_per_day)

train_fraction <- 0.8 # 80% Training Data + 10% Validation Data
test_fraction <- 0.1 # 10% Test Data

# Compute 80% of the Total Rows
n_train_rows <- floor(n_rows_total * train_fraction)

# Compute 10% of the Total Rows
n_test_rows <- floor(n_rows_total * test_fraction)

# Randomly Sample Rows for the Test Set
test_idx <- sample(1:n_rows_total, n_test_rows, replace=F)

# Used to Test Model
test_data <- trips_per_day[test_idx, ]

# Exclude Test Data from Training and Validation Sets
train_valid_pool <- trips_per_day[-test_idx, ]

# Number of Rows from the Pooled Data to Sample From
n_pool_rows <- train_valid_pool %>% nrow()

# Randomly Sample Rows for the Training Set 
train_idx <- sample(1:n_pool_rows, n_train_rows, replace=F)

# Used to Fit the Model
train_data <- train_valid_pool[train_idx, ]

# Used to Evaluate the Fit
valid_data <- train_valid_pool[-train_idx, ]

# Ensure Testing Set is Not Leaked into Training and Validation Data
# Return should be 0
intersect(train_data$id, test_data$id)
intersect(valid_data$id, test_data$id)

```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data

library(caret)

# Fix random seed so that fold creation is identical every run
set.seed(42)

# Create 5 cross-validation folds (data splits)
folds <- createFolds(train_valid_pool$num_trips, k = 5)


ctrl <- trainControl(
  method = "cv", # k-fold cross validation method
  index = folds # use the manually created splits above
)

# storage for CV results across polynomial degrees
rmse <- numeric(8)
se <- numeric(8)

# loop over polynomial degrees (model complexity)
for (k in 1:8) {

  # Build regression formula dynamically for current degree k
  formula <- as.formula(
    paste0("num_trips ~ poly(tmin, ", k, ", raw = TRUE)")
  )

  # fit linear model with polynomial term using 5-fold CV
  fit <- train(
    formula,
    data = train_valid_pool,
    method = "lm",
    trControl = ctrl
  )

  # Extract average cross-validated RMSE for this degree
  rmse[k] <- fit$results$RMSE

  # The standard error on this estimate
  se[k] <- fit$results$RMSESD
}

# combine results into a table for comparison across degrees
results <- data.frame(
  degree = 1:8,
  RMSE = rmse,
  SE = se
)

results

# select polynomial degree with lowest cross-validated RMSE
k <- results$degree[which.min(results$RMSE)]
k 
```

## Model comparison

Try different models and compare their validation performance.

```{r}
fml1 <- as.formula(paste0("num_trips ~ poly(tmin, ", k, ", raw = TRUE) + deep_snow + heavy_rain "))

mod1 <- train(
  fml1,
  data = train_valid_pool,
  method = "lm",
  trControl = ctrl
)

mod1$results # R-Squared = 0.72; RMSE = 5469.749

fml2 <- as.formula(paste0("num_trips ~ poly(tmax, ", k, ", raw = TRUE) + is_weekend + season "))

mod2 <- train(
  fml2,
  data = train_valid_pool,
  method = "lm",
  trControl = ctrl
)

mod2$results # R-Squared = 0.78658; RMSE = 4816.351

fml3 <- as.formula(paste0("num_trips ~ poly(tmax, ", k, ", raw = TRUE) + deep_snow + prcp + is_weekend"))

mod3 <- train(
  fml3,
  data = train_valid_pool,
  method = "lm",
  trControl = ctrl
)

mod3$results # R-Squared = 0.8397; RMSE = 4216.982

fml4 <- as.formula(paste0("num_trips ~ poly(tmax, ", k, ", raw = TRUE) + snow_event + prcp + is_weekend:is_holiday"))

mod4 <- train(
  fml4,
  data = train_valid_pool,
  method = "lm",
  trControl = ctrl
)

mod4$results # R-Squared = 0.8579; RMSE = 3979.996

fml5 <- as.formula(paste0("num_trips ~ poly(tmax, ", k, ", raw = TRUE) + snow_event + prcp:heavy_rain + is_weekend:is_holiday"))

mod5 <- train(
  fml5,
  data = train_valid_pool,
  method = "lm",
  trControl = ctrl
)

mod5$results # R-Squared = 0.8806; RMSE 3790.417

```

## Final model

Pick your best model and plot the results.

```{r}
train_data <- train_data %>%
  add_predictions(mod5) %>%
  mutate(split = "train")

valid_data <- valid_data %>%
  add_predictions(mod5) %>%
  mutate(split = "validate")

plot_data <- bind_rows(train_data, valid_data)

plot_data %>%
ggplot(mapping = aes(x = ymd, y = num_trips)) +
geom_point(aes(color = split)) +
  xlab('Date') +
  ylab('Daily trips')

plot_data %>%
ggplot(mapping = aes(x = pred, y = num_trips)) +
geom_point() + 
geom_abline(intercept = 0, slope = 1, size = 1, color = 'darkblue')
  xlab('Predicted Values') +
  ylab('Actual Values')
```

## Save model

```{r}
save(mod5, file = "Citibike_Prediction_Model.RData")
```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
