---
title: "Predicting Daily Citibike Trips"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
library(tidyverse)
library(scales)
options(na.action = na.warn)

library(broom)

theme_set(theme_bw())
```

## Load data

```{r}
trips <- read_tsv(here('week3','trips_per_day.tsv'))
head(trips)
```

## Train/test split

Hold out 10% of the data as a final test set. Use the remaining 90% for
training and validation.

```{r}
set.seed(42)
total_trips <- nrow(trips)
ninety <- 0.9*total_trips
ninety
sample <-  sample(1:total_trips ,ninety, replace = F)
num_fold <- 10
trips <- trips [sample, ]  %>% mutate(fold = (row_number() %% num_fold)+1 )
head(trips)


```

## Baseline model: temperature only

Start with the polynomial model from the complexity control notebook.

```{r}
# TODO: fit a linear model using poly(tmin, degree) on training data
# TODO: try different polynomial degrees and evaluate RMSE on validation data
K <- 1:7
average_validate_err <- c()

se_validate_err <- c()
for (k in K) {
    validate_err <- c()
    for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ poly(tmin,k,raw = T), data=trips_train )

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
    }
    average_validate_err[k] <- mean(validate_err)
    
    se_validate_err[k] <- sd(validate_err)/sqrt(num_fold)
}
min(average_validate_err)
# 5735.943

plot_data <- data.frame(K, average_validate_err, se_validate_err)
ggplot(plot_data, aes(x=K, y=average_validate_err)) +
  geom_pointrange(aes(ymin=average_validate_err - se_validate_err,
                      ymax=average_validate_err + se_validate_err,
                      color=average_validate_err == min(average_validate_err))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

## Feature engineering

Add features that might help: day of week, month, precipitation, etc.
Remember: only use information available *before* the day in question.

```{r}
trips <-trips %>% mutate(month = format(ymd, "%b"))
 validate_err <- c()
for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ tmin + month , data=trips_train )

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
}
 average_validate_err <- mean(validate_err)
 average_validate_err
#5436.708

```
```{r add-tmax}
validate_err <- c()
for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ tmin + prcp + month , data=trips_train )
#tmin * prcp + month
# tmin:prcp + month

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
}
average_validate_err <- mean(validate_err)
average_validate_err
# 4555.602

```
```{r add-prcp-feature}
validate_err <- c()
for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ tmin + prcp , data=trips_train )

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
}
average_validate_err <- mean(validate_err)
average_validate_err
# 5006.405

```

``` {r add-snow-feature}
validate_err <- c()
for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ tmin:snow + month , data=trips_train )

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
}
average_validate_err <- mean(validate_err)
average_validate_err
#  5835.191
```

## Model comparison

Try different models and compare their validation performance.

```{r}
#Cross- Validation 
K <- 1:7
average_validate_err <- c()

se_validate_err <- c()
for (k in K) {
    validate_err <- c()
    for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ poly(prcp,k,raw = T), data=trips_train )

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
    }
    average_validate_err[k] <- mean(validate_err)
    
    se_validate_err[k] <- sd(validate_err)/sqrt(num_fold)
}
min(average_validate_err)
#9881.623

plot_data <- data.frame(K, average_validate_err, se_validate_err)
ggplot(plot_data, aes(x=K, y=average_validate_err)) +
  geom_pointrange(aes(ymin=average_validate_err - se_validate_err,
                      ymax=average_validate_err + se_validate_err,
                      color=average_validate_err == min(average_validate_err))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

```{r month-model}
trips <-trips %>% mutate(month = format(ymd, "%b"))
 validate_err <- c()
for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~  month , data=trips_train )

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
}
 average_validate_err <- mean(validate_err)
 average_validate_err

 # 5556.315
```

```{r tmax model  }
K <- 1:7
average_validate_err <- c()

se_validate_err <- c()
for (k in K) {
    validate_err <- c()
    for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ poly(snow,k,raw = T), data=trips_train )

        trips_validate <- filter(trips,  fold == f)
        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
    }
    average_validate_err[k] <- mean(validate_err)
    
    se_validate_err[k] <- sd(validate_err)/sqrt(num_fold)
}
min(average_validate_err)
#5300

plot_data <- data.frame(K, average_validate_err, se_validate_err)
ggplot(plot_data, aes(x=K, y=average_validate_err)) +
  geom_pointrange(aes(ymin=average_validate_err - se_validate_err,
                      ymax=average_validate_err + se_validate_err,
                      color=average_validate_err == min(average_validate_err))) +
  geom_line(color = "red") +
  scale_x_continuous(breaks=1:12) +
  theme(legend.position="none") +
  xlab('Polynomial Degree') +
  ylab('RMSE on validation data')
```

## Final model

Pick your best model and plot the results.

```{r}
validate_err <- c()
prediction <- c()
actual <- c()
for (f in 1:num_fold) {
        trips_train <-   filter(trips,   fold != f)
        model <- lm(num_trips ~ tmin + prcp + month , data=trips_train )

         

        trips_validate <- filter(trips,  fold == f)
        predicted <- predict(model, trips_validate)

        prediction <- c(prediction,predicted )
        actual <- c(actual,trips_validate$num_trips)

        validate_err[f] <- sqrt(mean((predict(model, trips_validate) - trips_validate$num_trips)^2))
    
}
average_validate_err <- mean(validate_err)
average_validate_err

plot_data <- data.frame(predicted_value = prediction, actual_value = actual)

options(repr.plot.width=4, repr.plot.height=3)

ggplot(plot_data, aes(x = predicted_value, y = actual_value)) +
  geom_point() +
  geom_abline(linetype = "dashed") +
  xlab('Predicted') +
  ylab('Actual')

```

## Save model

```{r}

```

## Test set evaluation

**Do this only once, at the very end!**

```{r}
# TODO: load your saved model and predict on trips_test
# TODO: compute RMSE on test set
# TODO: how do you think this model will perform on future data?
```
