---
title: "predict_citibike"
author: "Christopher Esquivel"
date: '2022-06-16'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Get libraries and set up

```{r}
library(tidyverse)
library(dplyr)
library(scales)
library(lubridate)
library(modelr)

theme_set(theme_bw())
```

Get data

```{r}

holidays <- read_csv("US Bank holidays")

trips_per_day <- read_tsv("trips_per_day.tsv")

weather <- read_csv("weather.csv")

weather <- weather %>%
mutate(ymd = DATE)

trips_per_day <- inner_join(trips_per_day, weather, on = "ymd")

trips_per_day <- trips_per_day %>%
transmute(ymd, num_trips, prcp = PRCP, snwd = SNWD, snow = SNOW, tmax = TMAX, tmin = TMIN)


trips_per_day <- trips_per_day %>% left_join(holidays, by="ymd") %>% mutate(is_holiday = as.numeric(!is.na(holiday))) %>% mutate(is_weekend = as.numeric(wday(ymd) == c(1, 7))) %>% select(-c(n, holiday))

```

Plot number of trips and minimum temperature

```{r}
ggplot(trips_per_day, aes(x = tmin, y = num_trips)) +
  geom_point() +
  xlab('Minimum temperature') +
  ylab('Daily trips') +
  scale_y_continuous()
```
Get training and validation. => 90% of the data
Testing => 10%
```{r}
set.seed(6)

num_days <- nrow(trips_per_day)

frac_train <- 0.8
frac_validate <- 0.1
frac_test <- 0.1

num_train <- floor(num_days * frac_train)
num_validate <- floor(num_days * frac_validate)
num_test <- floor(num_days * frac_test)

index_train <- sort(sample(seq_len(num_days), size = num_train))
index_not_train <- setdiff(seq_len(num_days), index_train)
index_validate <- sort(sample(index_not_train, size = num_validate))
index_test <- setdiff(index_not_train, index_validate)

df_train <- trips_per_day[index_train, ]
df_validate <- trips_per_day[index_validate, ]
df_test <- trips_per_day[index_test, ]
```

Create models for tmin

```{r}
K <- 1:8
train_err <- c()
validate_err <- c()

for (k in K) {
  
  model <- lm(num_trips ~ poly(tmin, k, raw = T), data = df_train)
  
  train_err[k] <- sqrt(mean((predict(model, df_train) - df_train$num_trips)^2))
  
  validate_err[k] <- sqrt(mean((predict(model, df_validate) - df_validate$num_trips)^2))
}
```

Plot the training and validation error

```{r}
plot_data <- data.frame(K, train_err, validate_err) %>% gather("split", "error", -K)

ggplot(plot_data, aes(x = K, y = error, color = split)) + geom_line() + scale_x_continuous(breaks=K) + xlab('Polynomial Degree') + ylab('RMSE') 
```

Do the same thing for tmax

```{r}
K <- 1:8
train_err <- c()
validate_err <- c()

for (k in K) {
  
  model <- lm(num_trips ~ poly(tmax, k, raw = T), data = df_train)
  
  train_err[k] <- sqrt(mean((predict(model, df_train) - df_train$num_trips)^2))
  
  validate_err[k] <- sqrt(mean((predict(model, df_validate) - df_validate$num_trips)^2))
}
```

```{r}
plot_data <- data.frame(K, train_err, validate_err) %>% gather("split", "error", -K)

ggplot(plot_data, aes(x = K, y = error, color = split)) + geom_line() + scale_x_continuous(breaks=K) + xlab('Polynomial Degree') + ylab('RMSE') 
```


```{r}
K <- 1:8
train_err <- c()
validate_err <- c()

for (k in K) {
  
  model <- lm(num_trips ~ poly(tmin, k) + tmax + prcp + snwd + is_holiday + is_weekend, data = df_train)
  
  train_err[k] <- sqrt(mean((predict(model, df_train) - df_train$num_trips)^2))
  
  validate_err[k] <- sqrt(mean((predict(model, df_validate) - df_validate$num_trips)^2))
}


plot_data <- data.frame(K, train_err, validate_err) %>% gather("split", "error", -K)

ggplot(plot_data, aes(x = K, y = error, color = split)) + geom_line() + scale_x_continuous(breaks=K) + xlab('Polynomial Degree') + ylab('RMSE') 
```

```{r}
K <- 1:8
train_err <- c()
validate_err <- c()

for (k in K) {
  
  model <- lm(num_trips ~ tmin + poly(tmax, k) + prcp + snwd + is_holiday + is_weekend, data = df_train)
  
  train_err[k] <- sqrt(mean((predict(model, df_train) - df_train$num_trips)^2))
  
  validate_err[k] <- sqrt(mean((predict(model, df_validate) - df_validate$num_trips)^2))
}


plot_data <- data.frame(K, train_err, validate_err) %>% gather("split", "error", -K)

ggplot(plot_data, aes(x = K, y = error, color = split)) + geom_line() + scale_x_continuous(breaks=K) + xlab('Polynomial Degree') + ylab('RMSE') 
```

```{r}
K <- 1:8
train_err <- c()
validate_err <- c()

for (k in K) {
  
  model <- lm(num_trips ~ tmin + tmax + poly(prcp, k) + snwd + is_holiday + is_weekend, data = df_train)
  
  train_err[k] <- sqrt(mean((predict(model, df_train) - df_train$num_trips)^2))
  
  validate_err[k] <- sqrt(mean((predict(model, df_validate) - df_validate$num_trips)^2))
}


plot_data <- data.frame(K, train_err, validate_err) %>% gather("split", "error", -K)

ggplot(plot_data, aes(x = K, y = error, color = split)) + geom_line() + scale_x_continuous(breaks=K) + xlab('Polynomial Degree') + ylab('RMSE') 
```

```{r}
K <- 1:8
train_err <- c()
validate_err <- c()

for (k in K) {
  
  model <- lm(num_trips ~ tmin + tmax + prcp + poly(snwd, k) + is_holiday + is_weekend, data = df_train)
  
  train_err[k] <- sqrt(mean((predict(model, df_train) - df_train$num_trips)^2))
  
  validate_err[k] <- sqrt(mean((predict(model, df_validate) - df_validate$num_trips)^2))
}


plot_data <- data.frame(K, train_err, validate_err) %>% gather("split", "error", -K)

ggplot(plot_data, aes(x = K, y = error, color = split)) + geom_line() + scale_x_continuous(breaks=K) + xlab('Polynomial Degree') + ylab('RMSE') 
```

Final model

```{r}
fm_op <- lm(num_trips ~ tmin + tmax + prcp + snwd + is_holiday + is_weekend, data = df_train)

summary(fm_op)
```

Get RMSE when using the training data

```{r}

train_err <- sqrt(mean((predict(fm_op, df_train) - df_train$num_trips)^2))
  
validate_err <- sqrt(mean((predict(fm_op, df_validate) - df_validate$num_trips)^2))

train_err
validate_err
```

```{r}
fm_op <- lm(num_trips ~ tmin + tmax + prcp + snwd + is_holiday + is_weekend, data = df_train)

#After being trained, check how it does when using validation data
fm_op_validated <- df_validate %>% add_predictions(fm_op)

#Plot to see how good it follows the trend seen in the validation data
ggplot(data = fm_op_validated, mapping = aes(x = tmin, y = num_trips)) + geom_point(aes(y = num_trips)) + geom_line(aes(y = pred)) + xlab("Min Temperature") + ylab("Num trips per day")
```

```{r}
ggplot(data = fm_op_validated) + geom_point(mapping = aes(x = pred, y = num_trips)) + geom_abline(linetype = "dashed")
```

```{r}
date_num_trips <- ggplot(data = fm_op_validated) + geom_line(aes(x = ymd, y = pred)) + geom_point(aes(x = ymd, y = num_trips)) + xlab("date") + ylab("num trips")

pred_and_actual <- ggplot(data = fm_op_validated) + geom_point(aes(x = pred, y = num_trips)) + geom_abline(linetype = "dashed")

date_num_trips
pred_and_actual
```

Time to save the model

```{r}
save(trips_per_day, df_train, df_validate, df_test, fm_op, date_num_trips, pred_and_actual, file = "predict_citibike.RData")
```

Testing time!

```{r}
fm_op_tested <- df_test %>% add_predictions(fm_op)

rmse_test <- sqrt(mean((predict(fm_op, df_test) - df_test$num_trips)^2))

rmse_test
```
Plot the model with the testing data

```{r}
ggplot(data = fm_op_tested) + geom_line(aes(x = ymd, y = pred)) + geom_point(aes(x = ymd, y = num_trips)) + xlab("date") + ylab("num trips")

ggplot(data = fm_op_tested) + geom_point(aes(x = pred, y = num_trips)) + geom_abline(linetype = "dashed")
```

When passing in the data for testing instead of the training data.
adjusted R-squared is 0.87

```{r}
fm_tested <- lm(num_trips ~ tmin + tmax + prcp + snwd + is_holiday + is_weekend, data = df_test)

summary(fm_tested)
```

I think my model will do good when testing on new data. It is a simple model. Perhaps it will perform the same as when using the testing data. 
