---
title: "Testing Daily Citibike Trips for 2015 with LM trained on 2014"
author: "Jacqueline Louie (Stony Brook University)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(modelr)
```

```{r}
trips_per_day_2015 <- read_tsv('../week4/trips_per_day_2015.tsv')
weather_2015 <- read_csv('../week4/weather_2015.csv')

head(trips_per_day_2015)
head(weather_2015)
```

Write a new file called test_citibike_predictions.Rmd that loads in the new trips_per_day_2015.csv 
and weather_2015.csv files along with your saved model (from yesterday's .Rdata file), and predicts 
the number of trips for each day. If you used any other data for your model, make sure to include code
 that downloads and incorporates that data as well.

Compute the RMSE between the actual and predicted trips for 2015 and compare the results to what you 
found with cross-validation on the 2014 data.

```{r}
weather_2015_adj <- mutate(weather_2015, ymd = DATE, prcp = PRCP, tmin = TMIN, tmax = TMAX, snwd = SNWD, snow = SNOW) %>%
    select(ymd, prcp, snwd, snow, tmin, tmax) %>% 
    mutate(tmin = tmin * 0.10, tmax = tmax * 0.10) # mutating tmin and tmax to match the scale for 2014 tmin tmax

# joining on DATE
test_set <- inner_join(trips_per_day_2015, weather_2015_adj, by = "ymd") %>% 
    mutate(day_of_week = wday(ymd), month = month(ymd))

head(test_set)
```

```{r}
load("final_model.RData")
final_model

rmse(final_model, test_set)
rsquare(final_model, test_set)
```