#2014 Results and Test Data
Results from Validation Data
RMSE = 3447.48	
R^2 = 0.8330824	

Results from Test Data
RMSE = 3574.282	
R^2 = 0.8654721		

#Testing the 2015 data
```{r}
#download all necessary packages
library(tidyverse)
library(scales)
library(modelr)

#load data files
load("trips_2015.Rdata")
load("model.RData")
holiday <- read_csv("US Bank holidays", col_names = c("row_number", "ymd", "title"))

theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)

#join the datasets and summarize number of counts
trips_summary <-trips_2015 %>% group_by(ymd) %>%
  summarize(count = n())
trips_2015 <- left_join(trips_summary, weather_2015, by = 'ymd')
trips_2015 <- left_join(trips_2015, holiday, by = "ymd") %>% mutate(holiday = !is.na(title)) %>% mutate(holiday = as.integer(holiday))
trips_2015 <- trips_2015 %>% rename(num_trips = count)

#convert fahrenheit to celsius
#library(weathermetrics)
#trips_2015$tmax <- fahrenheit.to.celsius(trips_2015$tmax, round = 2)

#adding other variables used in prediction model
library(lubridate)
trips_2015$day_of_week <- wday(trips_2015$ymd)
#1 is Sunday
trips_2015 <- trips_2015 %>% mutate(weekday = day_of_week > 1 & day_of_week < 7) %>%
  mutate(substantial_prec = as.integer(prcp > 0.7))

#test dataset is all of 2015
trips_2015 <- trips_2015 %>%
  add_predictions(model_c) %>%
  mutate(split = "test")

#Predictions
pred_actual <- trips_test %>%
  add_predictions(model_c) %>%
  mutate(actual = num_trips)

sd(trips_test$num_trips)

pred_actual %>%
  summarize(rmse = sqrt(mean((pred - actual)^2)),
             r_sq = rsquare(model_c,trips_2015))
```

When running the 2015 data the results for RMSE were 3574.282 and the R^2 was 0.702379 which told us that our model was pretty good in predicting number of trips as it was for the year 2014. 

#Testing the 2020 data
```{r}
#download all necessary packages
library(tidyverse)
library(scales)
library(modelr)

#load data
load("trips_2020.Rdata")
load("model.RData")
holiday <- read_csv("US Bank holidays", col_names = c("row_number", "ymd", "title"))

theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)

#join the datasets and summarize number of counts
trips_summary <-trips_2020 %>% group_by(ymd) %>%
  summarize(count = n())
trips_2020 <- left_join(trips_summary, weather_2020, by = 'ymd')
trips_2020 <- left_join(trips_2020, holiday, by = "ymd") %>% mutate(holiday = !is.na(title)) %>% mutate(holiday = as.integer(holiday))
trips_2020 <- trips_2020 %>% rename(num_trips = count)

#adding other variables used in prediction model
library(lubridate)
trips_2020$day_of_week <- wday(trips_2020$ymd)
#1 is Sunday
trips_2020 <- trips_2020 %>% mutate(weekday = day_of_week > 1 & day_of_week < 7) %>%
  mutate(substantial_prec = as.integer(prcp > 0.7))

#test dataset is all of 2015
trips_2020 <- trips_2020 %>%
  add_predictions(model_c) %>%
  mutate(split = "test")


#Predictions
pred_actual <- trips_2020 %>%
  add_predictions(model_c) %>%
  mutate(actual = num_trips)

sd(trips_2020$num_trips)

pred_actual %>%
  summarize(rmse = sqrt(mean((pred - actual)^2)),
            r_2 = rsquare(model_c,trips_2020))
```

The 2020 data produced a RMSE of 35142.22 and R^2 of 0.3558857 which shows our model was not suited for predicting the number of trips like it was for the years 2014 and 2015. Factors like Covid and laws would need to be implemented into the model for 2020 to produce better results. 

The model I originally had from yesterday's data needed to be tweeked to produce better results for 2014's test data and 2015, especially after changing the temperature units.