---
title: "Week3_Day1_Classwork"
output: html_document
date: '2022-06-13'
---

```{r}
library(tidyverse)
library(scales)
library(modelr)
library(scales)
library(broom)
library(MASS)
library(ISLR2)

options(na.action = na.warn)

theme_set(theme_bw())
options(repr.plot.width=4, repr.plot.height=3)

```

```{r}

body <- read.table('body.dat.txt')

model <- lm(body$V23 ~ V24, data = body)

tidy(model)


```

Lab 3.6.3

```{r}

library(car)

lm.fit <- lm(medv ~ lstat + age, data = Boston)
summary(lm.fit)

lm.fit2 <- lm(medv ~ ., data = Boston)
summary(lm.fit2)

vif(lm.fit2)

lmfit3 <- lm(medv ~ . - age, data = Boston)
summary(lmfit3)

lm.fit1 <- update(lm.fit, ~ . - age)

```
Lab 3.6.4

```{r}
summary(lm(medv ~ lstat * age, data = Boston))
```

Lab 3.6.5

```{r}

lm.fit2 <- lm(medv ~ lstat + I(lstat^2), data = Boston)
summary(lm.fit2)

lm.fit <- lm(medv ~ lstat, data = Boston)
anova(lm.fit, lm.fit2)

par(mfrow = c(2,3))
plot(lm.fit2)

lm.fit5 <- lm(medv ~ poly(lstat, 5), data = Boston)
summary(lm.fit5)

summary(lm(medv ~ log(rm), data = Boston))

```


Lab 3.6.6

```{r}

head(Carseats)

lm.fit <- lm(Sales ~ . + Income: Advertising + Price: Age, data = Carseats)
summary(lm.fit)

attach(Carseats)
contrasts(ShelveLoc)

```

6.1

(a) y = -8.94x + 123.05
(b)  The estimated body weight of babies born to smoking mothers is 8.94 ounces lower than babies born to non-smoking mothers.
  Non-smoker: y = -8.94(0) + 123.05 = 123.05
  Smoker: y = -8.94(1) + 123.05 = 114.11
(c) The p-value = 0 < 0.05, thus there is a statistically significant relationship between the average birth weight and smoking

```{r}
babyweights <- read.table('babyweights.txt')

model <- lm(bwt ~ smoke, data = babyweights)
tidy(model)


```

6.2

(a) y = -1.93x + 120.07
(b) The estimated body weight of non-firstborn babies is 1.93 ounces less than firstborn babies.
  Firstborn: 120.07
  Not firstborn: 118.14
(c) Since p-value = 0.1052 > 0.05, we fail to reject the null hypothesis. There isn't a statistically significant relationship between the average birth weight and parity. .

```{r}
babyweights <- read.table('babyweights.txt')

model <- lm(bwt ~ parity, data = babyweights)
tidy(model)
```
6.3

(a) weight = -80.41 + 0.44(gestation) - 3.33(parity) -0.01(age) + 1.15(height) + 0.05(weight) - 8.4(smoke)
(b) As the gestation increases by 1 day, the average baby weight increases by about 0.44 ounces.
  As the mother's age increases, the average baby weight decreases by about 0.01 ounces.
  (If all else is held constant)
(c) Parity might be correlated with another predictor in the model which complicates model estimation.
(d) 120 - 120.58 = -0.58

```{r}

-80.41 + 0.44 * 284 - 3.33 * 0 -0.01*27 + 1.15*62 + 0.05*100 - 8.4*0

```
(e) R^2 = 0.2504435
  Adj R^2 = 0.2467842
```{r}
#R^2
1 - (249.28/332.57)

#Adj R^2
1 - ((249.28/(1236-6-1))/(332.57/(1236-1)))
```
```{r}
babyweights <- read.table('babyweights.txt')

model <- lm(bwt ~ ., data = babyweights)
tidy(model)


```
Lab 5.3.1

```{r}
library(ISLR2)
set.seed(1)
train <- sample(392, 196)

lm.fit <- lm(mpg ~ horsepower, data = Auto, subset = train)

attach(Auto)
mean((mpg - predict(lm.fit, Auto))[-train]^2)

lm.fit2 <- lm(mpg ~ poly(horsepower,2), data = Auto, subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)

lm.fit3 <- lm(mpg ~ poly(horsepower,3), data = Auto, subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)

set.seed(2)
train <- sample(392, 196)
lm.fit <- lm(mpg ~ horsepower, subset = train)
mean((mpg - predict(lm.fit, Auto))[-train]^2)
lm.fit2 <- lm(mpg ~ poly(horsepower,2), data = Auto, subset = train)
mean((mpg - predict(lm.fit2, Auto))[-train]^2)
lm.fit3 <- lm(mpg ~ poly(horsepower,3), data = Auto, subset = train)
mean((mpg - predict(lm.fit3, Auto))[-train]^2)

```
Lab 5.3.2

```{r}
glm.fit <- glm(mpg ~ horsepower, data = Auto)
coef(glm.fit)

lm.fit <- lm(mpg ~ horsepower, data = Auto)
coef(lm.fit)

library(boot)
glm.fit <- glm(mpg ~ horsepower, data = Auto)
cv.err <- cv.glm(Auto, glm.fit)
cv.err$delta

cv.error <- rep(0,10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error[i] <- cv.glm(Auto, glm.fit)$delta[1]
}

cv.error

```

Lab 5.3.3

```{r}
set.seed(17)
cv.error.10 <- rep(0,10)
for (i in 1:10) {
  glm.fit <- glm(mpg ~ poly(horsepower, i), data = Auto)
  cv.error.10[i] <- cv.glm(Auto, glm.fit, K = 10)$delta[1]
}
cv.error.10

```


