6.1 
a. Regression line equation 
 y = 123.05-8.94x 
 slope = -8.94
 intercept = 123.05
 
 b. plugging in 1 for smoker and 0 for non-smoker. 
 For smokers 
 y = 123.05 - 8.94(1) = 114.11 
 
 For non smokers
 y = 123.05 - 8.94(0) = 123.05 
 
 c. It is significant because p-value is < 0.05. 
 
 6.2 
 
 a. Regression line equation 
 
 y = 120.07 - 1.93X 
 
 b. plugging in x =0 for first born. y = 120.07
    plugging in x =1 for others. y = 118.14 
    
    c. 0.1052 >= 0.05. i.e., p-value is not <= 0.05. Hence, it is not significant. 
    

6.3 

a. y = -80.41 + 0.44(gestation) - 3.33 (parity) - 0.01 (age) + 1.15 (height) + 0.05 (weight) - 8.40 (smoke)

b. slope of gestation = 0.44
slope of age = -0.01 

c. There might be difference because the measurement is included when holding the other variables constant. 

d. residual 
ei = yi - y^i 

```{r}
ei = 120 - (-80.41 + 124.96 - 0.27 + 71.3 +5)
ei
```
e. r^2 = 1 - var(ei)/ var(yi) 
```{r}
1 - 249.28/332.57
```

f. adjusted R^2 = r^2 * n-1/ n-k-1
```{r}
0.25044 * (1236-1)/(1236-6-1)
```

